### 2015-06-22 main.tex
Initial commit

\documentclass{article}

\usepackage{graphicx} 
\usepackage{paralist}

\usepackage{hyperref}

\usepackage{url}
\usepackage{booktabs}

\usepackage{tikz}

\usepackage[draft,nomargin,footnote]{fixme}

\graphicspath{{figs/}}

\usepackage{xspace}
\newcommand{\eg}{\textit{e.g.}\xspace}
\newcommand{\etal}{\textit{et al.}\xspace}
\newcommand{\ie}{\textit{i.e.}\xspace}
\newcommand{\etc}{\textit{etc.}\xspace}
\newcommand{\vs}{\textit{vs.}\xspace}

\title{Learning by Teaching a Robot: the Case of Handwriting}

\author{S\'everin Lemaignan, Alexis Jacq, Fernando Garcia, Pierre
Dillenbourg \\ CHILI Lab, \'Ecole Polytechnique F\'ed\'erale de Lausanne}

\begin{document}
\maketitle
\end{document}

### 2015-06-23 main.tex
First structure, placeholder pictures

@@
 \usepackage{graphicx} 
+\usepackage{subfigure}
 \usepackage{paralist}
@@
 
+\usepackage[usenames,dvipsnames]{xcolor}
 \usepackage{tikz}
+\usetikzlibrary{positioning, calc}
 
@@
 
-\author{S\'everin Lemaignan, Alexis Jacq, Fernando Garcia, Pierre
+\author{S\'everin Lemaignan, Alexis Jacq, Fernando Garcia, Ana Paiva, Pierre
 Dillenbourg \\ CHILI Lab, \'Ecole Polytechnique F\'ed\'erale de Lausanne}
@@
 \maketitle
+
+\begin{abstract}
+    TDB
+\end{abstract}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{A different paradigm for educative robots}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{henry}
+    \caption{Henry teaching Nao how to write numbers, with the help of an
+    occupational therapist.}
+    \label{fig:henry}
+\end{figure}
+
+The \emph{learning by teaching} paradigm, which engages the target student in
+the act of teaching another, has been shown to produce motivational,
+meta-cognitive, and educational benefits in a range of disciplines
+\cite{Rohrbeck2003}. To our best knowledge, the application of the paradigm to
+handwriting intervention remains, however, unexplored. One reason for this may
+due to the requirement of an appropriately unskilled peer for the target child
+to tutor, as this may present a logistical constraint if the target child is the
+lowest performer in their class.  In some cases, it may be appropriate for a
+peer or teacher to simulate a na\"ive learner for the target child to teach.
+For handwriting, where one's skill level is visually evident, however, this
+acting is likely to be eventually detected. As such, there is motivation for the
+use of a teachable agent which can be configured for a variety of skill levels,
+and for which children do not have preconceptions about its handwriting ability.
+
+Also notable, the robot is not used in the usual context of robotics or computer
+education, but instead in an activity -- handwriting -- which requires fine
+physical skills. In such activities, the embodied nature of the robot is
+appropriate as in interventions where motor mimicry is elicited
+\cite{Berninger1997} the arm motion for instance is, \emph{by itself}, part of
+the teaching. Furthermore, when facing a child with school difficulties, robots
+can play the role of a na\"ive learner which neither adults nor peers -- because
+of the social effects it would induce -- can convincingly play. Along these
+lines, we hope to see more research on non-STEM educational applications of
+robotics.
+
+Beyond handwriting, we believe that this work provides a novel perspective on
+the role for robots in the field of education. \emph{Learning by teaching} is a
+powerful paradigm because of not only its pedagogical efficacy, but its
+potential to positively impact the child's motivation and self-esteem. We hope that 
+this article shows that this is a very relevant context of use for robots. Indeed,
+when facing a child with school difficulties, robots can play the role of a na\"ive 
+learner which neither adults nor peers -- because of the social effects it would 
+induce -- can convincingly play.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{The robotic system}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Algorithmic Approach}
+
+\subsubsection{Path Manipulation in the Letters' Eigenspace}
+
+\begin{figure}[ht!]
+    \centering
+    \subfigure[Nine samples of the cursive ``h'', next to the reference]{
+        \includegraphics[height=6cm]{h}
+    }
+    \subfigure[Raw samples projected in the Eigenspace of the dataset]{
+        \includegraphics[height=4cm]{eigenspace-base}
+    }
+    \subfigure[Normalized samples, projected in the Eigenspace. The letters are
+    clustered according to their styles.]{
+        \includegraphics[height=4cm]{eigenspace-uniformization}
+    }
+
+    \caption{\small Projecting demonstrated letters onto the Eigenspace
+    generated from the reference dataset effectively clusters the samples
+according to their topological similarity.}
+    \label{fig:stimuli}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Robotic Embodiement}
+\begin{figure}[ht]
+\centering
+
+\resizebox{0.9\linewidth}{!}{%
+
+\begin{tikzpicture}[
+    >=latex,
+    node distance=2cm,
+    every edge/.style={draw, very thick},
+    redarrow/.style={fill=Red, draw=Red},
+    greenarrow/.style={fill=GreenYellow, draw=GreenYellow},
+    yellowarrow/.style={fill=BurntOrange, draw=BurntOrange},
+    cmpt/.style={draw, align=center, rounded corners, inner sep=5pt, font=\sf, fill=black!20},
+    label/.style={midway, align=left, font=\scriptsize\sf, fill=white, above,opacity=0,text opacity=1}]
+
+    \node at (0,0)[cmpt, fill=Cyan!50, text width=2.7cm] (tactile) {tactile interaction \\ manager};
+    \node[cmpt, fill=Cyan!50, left=of tactile] (display) {display \\ manager};
+
+    \node [cmpt, fill=Green!50, below=of tactile, text width=2.7cm] (shape) {shape position \\ manager};
+    \node [cmpt, fill=Green!50, left=of shape] (learner) {letter \\ learner};
+    \node [cmpt, fill=Green!50, left=of learner] (card) {card \\ detector};
+
+    \node [cmpt, fill=YellowOrange!50, below=1.5cm of shape] (writer) {nao writer};
+    \node [cmpt, fill=YellowOrange!50, left=of writer] (tablet) {tablet detector};
+    \node [cmpt, fill=YellowOrange!50, left=of tablet] (camera) {nao camera};
+    \node [cmpt, fill=YellowOrange!50, below=0.6cm of tablet] (pose) {robot state publisher};
+
+    %%% Relations between components
+    \path (tactile) edge [->] node[label] {clear \\ screen} (display);
+
+    \path ($(shape.north west)!0.33!(shape.north east)$) edge [<-] node[label,left] {clear \\ screen} ($(tactile.south west)!0.33!(tactile.south east)$);
+    \path ($(shape.north west)!0.66!(shape.north east)$) edge [<-, greenarrow] node[label,right] {feedback \\ + location} ($(tactile.south west)!0.66!(tactile.south east)$);
+
+    \path (shape) edge [->, redarrow] node[label, left, align=left] {letter \\to write \\ + time \\ + position} (display);
+    \path (shape) edge [->,redarrow] node[label, right] {letter to write \\ + time \\ + position in\\tablet frame} (writer);
+
+    \path ($(shape.north west)!0.66!(shape.south west)$) edge [->, greenarrow] node[label,below] {feedback \\ + letter} ($(learner.north east)!0.66!(learner.south east)$);
+    \path ($(shape.north west)!0.33!(shape.south west)$) edge [<-, redarrow] node[label] {letter to write \\ + time} ($(learner.north east)!0.33!(learner.south east)$);
+
+    \path (card) edge [->, yellowarrow] node[label] {word to learn} (learner);
+    \path (camera) edge [->, yellowarrow] node[label, right] {camera image} (card);
+    \path (camera) edge [->, redarrow] node[label] {camera image} (tablet);
+    \path (tablet) edge [->, redarrow] node[label] {tablet pose} (writer);
+    \path (pose) edge [->, redarrow] node[label, right] {head pose} (tablet);
+
+    \path (-8, 0) edge [->, redarrow] node[label] {Writing letters} (-7, 0);
+    \path (-8, -0.6) edge [->, greenarrow] node[label] {Responding to feedback} (-7, -0.6);
+    \path (-8, -1.2) edge [->, yellowarrow] node[label] {Detecting card input} (-7, -1.2);
+
+\end{tikzpicture}
+}
+
+\caption{Overview of the system. Components in the top row run on the tablet, 
+those in the middle row on the central controller, and those in the bottom row 
+on the robot.}
+
+    \label{fig:archi}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Field Applications}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Studies at School}
+
+\begin{figure}
+    \centering
+    \includegraphics[height=3.9cm]{schools}
+    \includegraphics[height=3.9cm]{schools2}
+    \caption{The first field studies were focused on technical validation, with
+    above 100 pupils interacting with the robot over short periods (around 20
+minutes), either alone or in small groups.}
+    \label{fig:schools}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Case Study 1: Diego}
+
+\subsection{Context, Study Design}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\linewidth]{diego}
+    \caption{Diego working with Nao}
+    \label{fig:diego}
+\end{figure}
+
+\subsubsection{Results}
+
+\begin{figure}
+    \centering
+    \subfigure[Initial letter, generated by the robot]{
+        \includegraphics[height=6cm]{diego-initial-letter}
+    }
+    \subfigure[Final letter, after training with Diego]{
+        \includegraphics[height=6cm]{diego-final-letter}
+    }
+
+    \caption{\small Text generated by the robot, before and after training with
+    the child. As an example, the red box emphasizes the changes on the word
+    ``envoyer''.}
+    \label{fig:stimuli}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Case Study 2: Henry}
+
+\subsubsection{Context, Study Design}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{abc-writer}
+    \caption{Screenshot of the Android application developed to be used as
+    pre-test and post-test: the child must follow with his finger the path of
+the letter or number. We count the number of times the finger goes outside of
+the red-bordered area.}
+    \label{}
+\end{figure}
+
+\subsubsection{Results}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{learning_6}
+    \caption{Progresses made by Henry over one 45-minutes interaction session
+    while writing ``6''.}
+    \label{}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Agency and Commitment}
+
+However, we believe that the strongest impact of this work is for the
+human-robot interaction community and relates to the very \emph{nature} of the
+interaction fostered by this research. The work presented here investigates a
+particular role for a robot in the education of handwriting: not only is the
+robot actively performing the activity by drawing letters, but it does so in a
+way that engages the child in a very specific social role. The child is the
+teacher in this relationship and the robot is the learner: the child must engage
+in a (meta-) cognitive relationship with the robot to try to understand why the
+robot fails and how to help it best.  Here, the robot is more than just an
+activity facilitator or orchestrator -- its physical presence and embodiment
+induce agency and anthropomorphising, and cognitively engage the child into the
+learning activity, which we predict will lead to higher learning efficacy.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Acknowledgments}
+
 \end{document}

### 2015-06-23 main.tex
Copy-pasted content from HRI publication

@@
 
+Henry is five and an half, and has been diagnosed with visuo-constructive
+deficits. He is under the care of an occupational therapist, and tries to improve
+his inability to draw letters in a consistent manner. Diego is six and struggles
+at school with his poor handwriting and even poorer self-confidence.
+
+While Henry is lively and always quick at shifting his attention from one
+activity to another, Diego is shy and poised. Two very different children,
+facing however the same difficulty to write in a legible manner.
+
 \begin{figure}
@@
 
+Remediations for handwriting difficulties
+
+\subsection{Learning by Teaching}
+
 The \emph{learning by teaching} paradigm, which engages the target student in
 the act of teaching another, has been shown to produce motivational,
-meta-cognitive, and educational benefits in a range of disciplines
-\cite{Rohrbeck2003}. To our best knowledge, the application of the paradigm to
+meta-cognitive, and educational benefits in a range of disciplines~\cite{Rohrbeck2003}. 
+To our best knowledge, the application of the paradigm to
 handwriting intervention remains, however, unexplored. One reason for this may
@@
 and for which children do not have preconceptions about its handwriting ability.
+
+Robots have been used as teachers or social partners to promote children's
+learning in a range of contexts, most commonly related to language
+skills~\cite{han2010robot}, and less often to physical skills (such as
+calligraphy~\cite{Matsui2013}).  Looking at the converse (humans \emph{teaching}
+robots), Werfel notes in~\cite{Werfel2014} that most of the work focuses on the
+robot's benefits (in terms of language~\cite{Saunders2010} or
+physical~\cite{Mulling2013} skills, for example) rather than the learning
+experienced by the human tutor themselves.  Our work concentrates on this latter
+aspect: by demonstrating handwriting to a robot, we aim at improving the
+\emph{child's} performance. Note that our work must be distinguished from
+``learning from demonstration'' approaches to robots learning physical skills,
+as the agent we present is only simulating fine motor skills for interaction
+purposes.
+
+A robotic learning agent which employs the learning by teaching paradigm has
+previously been developed by Tanaka and Matsuzoe~\cite{Tanaka2012}. In their
+system, children learn vocabulary by teaching the {\sc nao} robot to act out
+verbs. The robot is tele-operated and mimics the actions that the children teach
+it, but with no long-term memory or learning algorithm in place.  Our project
+significantly extends this line of work in two ways. First, by investigating the
+context of children's acquisition of a challenging physical skill (handwriting),
+and second by proposing a robotic partner which is fully autonomous in its
+learning.
+
+\subsection{Teaching beyond ICT}
+
+This research departs from the traditional approaches to educative robotics
+(\fixme{cite reviews})
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The robotic system}
+\section{The Robotic System}
 
@@
 \subsection{Algorithmic Approach}
+
+\begin{figure}[thpb]
+    \centering
+    \includegraphics[width=0.6\columnwidth]{learningSdemo}
+    \caption{\label{fig:demonstrationShapes2}Example of the learning algorithm
+    responding (top) to user demonstration of shapes (bottom) for the letter `s'
+    (demonstrations received from two 7-8 year-old children taking turns).}
+
+\end{figure}
 
@@
 
+We would like to thank Deanna Hood for her key contribution to the project.
+This research was partially supported by the Funda\c{c}\~{a}o para a Ci\^{e}ncia
+e a Tecnologia (FCT) with reference UID/CEC/50021/2013, and by the Swiss
+National Science Foundation through the National Centre of Competence in
+Research Robotics.
+
+\bibliographystyle{abbrv}
+\bibliography{biblio}
+
 \end{document}

### 2015-07-02 main.tex
General experimental setup

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Robotic System}
+\section{Implementation of the Interaction}
+
+Figure~\ref{experimental_setup} illustrates our general experimental setup: a
+face-to-face child-robot interaction with an (autonomous) Aldebran's Nao robot.
+
+A tactile tablet (with a custom application) is used for both the robot and the
+child to write: during a typical round, the child requests the robot to write
+something (a single letter, a number or a full word), and push the tablet
+towards the robot, the robot writes on the tablet by gesturing the writing, but
+\emph{without actually physically touching the tablet}), the child then pull
+back the tablet, correct the robot's attempt by writing him/herself on top or
+next to the robot's writing (see the picture in Figure~\ref{fig:diego}), and
+``send'' his/her demonstration to the robot by pressing a small button on the
+tablet. The robot ``learns'' from this demonstration and tries again.
+
+Since the child is assumed to take on the role of the teacher, we had to ensure
+he/she would be able to manage by him/herself the turn-taking and the overall
+progression of the activity (moving to the next letter or word). In our design,
+the turn-taking relies on the robot prompting for feedback once it is done with
+its writing (simple sentences like ``What do you think?''), and pressing on a
+small robot icon on the tablet once the child has finished correcting. In our
+experiments, both were easy to grasp for children.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{experimental_setup}
+    \caption{Our experimental setup: face-to-face interaction with a Nao robot.
+    The robot writes on the tactile tablet, the child then corrects the robot by
+    directly overwriting its letters on the tablet, with a stylus directly. An adult
+    (either a therapist or an experimenter, depending on the studies), remains next
+    to the child to guide the work (prompting, turn taking, etc.).}
+    \label{experimental_setup}
+\end{figure}
+
+Implementing such a system raises several challenges: first, acquiring,
+analyzing and learning from hand-written demonstration lays at the core of the
+our approach and we developed several algorithms for the robot to generate
+initial bad writing and to respond in an adequate manner, showing visible (but
+not too quick) writing improvements.
+
+Then, the actual implementation on the robot required the coordination of
+several modules (from performing gestures and acquiring the user's input to
+the high-level state machine), spread over several machines (the robot itself,
+one laptop and up to four tactile tablets for certain studies we conducted). We
+relied on ROS to ensure the synchronization and communication between these
+modules.
+
+Finally, the interaction itself, involving a robot, one or two children and, in
+one study, an occupational therapist, also raised by itself several challenges.
+
+We detail each of these points in the following sections.
 

### 2015-07-02 main.tex
Robotic implementation (for now copy pasted from HRI paper)

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Robotic Embodiement}
+\subsection{Robotic Implementation}
+
+In order to develop a teachable agent that is appropriate for engaging a child
+in the learning by teaching paradigm, we have established capabilities for the
+robot to engage in handwriting and interactive turn-taking.
+
+The {\sc nao} V4 humanoid robot, which has been purposely designed by Aldebaran
+Robotics to look approachable~\cite{Gouaillier2008}, is used for this work. It
+is a commercially affordable biped robot, 58cm tall, with 25 degrees of freedom,
+two cameras, speech capabilities and the ability to autonomously execute a range
+of tasks.
+
+Precise control over what the robot is writing is necessary in the proposed
+application of a handwriting. Because of the limited fine motor skills possible
+with such an affordable robot, in addition to the absence of force feedback and
+other technical necessities, we have configured the {\sc nao} to use simulated
+handwriting with a synchronised tablet to achieve this level of control.
+
+The development of the necessary components for embedding the handwriting
+learning algorithm presented in Section \ref{sec:learningAlgorithm} in the
+humanoid agent are presented in the sections which follow.
+
+\subsubsection{Robot Trajectory Following Movements}
+
+Using simulated handwriting provides an opportunity for the robot's writing to
+appear smoother than would be achievable with a writing instrument. However, the
+robot's motions must still sufficiently match the displayed trajectory in order
+capture the engagement of the child participant in the action. Aldebaran's NaoQi
+API is used for the inverse kinematics of the trajectory following. The Robot
+Operating System (ROS)\footnote{The ROS stack for {\sc nao} is available at:
+\url{http://wiki.ros.org/nao_robot}.} is used for integration of the {\sc nao}
+with external reference frames, such as the tablet's location, using the $tf$
+transformation library \cite{Foote2013}.
+
+When using simulated handwriting, it is no longer necessary that the robot
+engages in the typical style of handwriting of using a writing instrument at a
+desk.  Having the robot point at a vertical writing surface to cause the
+trajectory to appear (as in Figure \ref{fig:naoWriting}) has several advantages:
+
+\begin{itemize}
+
+    \item The working space of the robot increases, both in the technical sense
+        and the interaction sense: someone can, in theory, show the tablet to
+        the robot from across the room and have it still respond, without
+        needing the tablet to be within arm's reach.
+
+    \item Concerns about whether or not the child would start mimicking the
+        robot's incorrect writing form (\eg pen grip) are mitigated.
+
+    \item Perhaps most significantly, the accuracy of the matching of the
+        robot's motion with the trajectory displayed on the tablet is not as
+        critical. This is because a pen tip would be expected to touch the
+        tablet exactly at the trajectory point, while a fingertip may not.
+
+\end{itemize}
+
+\begin{figure}[thpb]
+     \begin{center}
+            \includegraphics[width=0.7\linewidth]{naoWriting2.png}
+    \end{center}
+    \caption{A demonstration of the robot simulating the writing of a word with
+    its finger. The motion of the robot is synchronised with the display of the
+    tablet, communicating over ROS.}%
+
+   \label{fig:naoWriting}
+\end{figure}
+
+We have therefore designed the system in such a way that the robot is simulating
+handwriting by pointing at the tablet\footnote{Teachers interviewed for their
+feedback on the system advised that children are asked to draw letters in
+the air in a similar manner as part of their handwriting education. The
+behaviour is hence not unfamiliar to children.}. As interacting with a
+tablet with one's finger is not uncommon, this may aid the acceptance of the
+writing style by users.
+
+Because motion planning is performed with respect to the hand of the robot,
+rather than its fingertip, one or two of the orientation degrees of freedom of
+the hand are fixed to keep the finger approximately perpendicular to the writing
+surface, depending on the desired accuracy. The remaining free orientation(s),
+coupled with the whole-body motion control available, allow for a sufficient
+working space for writing on the entire tablet.
+
 \begin{figure}[ht]

### 2015-07-02 main.tex
List of studies

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Interaction Design}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Field Applications}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Field Studies}
+
+To date, the presented system has been deployed and tested in three different
+schools (totalling child-robot interactions with more than 70 children, aged 5
+to 8) and we also conducted 2 case studies over several weeks.
+Table~\ref{studies} gives an overview of these studies.
+
+\begin{table}[ht!]
+\centering
+\caption{Field studies conducted within the project}
+\label{studies}
+\begin{tabular}{@{}lp{4.5cm}p{2.2cm}p{2.5cm}l@{}}
+\toprule
+{\bf Study}     & {\bf Type}                                    & {\bf Duration}              & {\bf Children \#}                     & {\bf Ages} \\ \midrule
+{\it ISG 1}     & Unstructured group interaction at school      & 16 min/group               & 4 $\times$ 8 children                 & 6-7        \\
+{\it Florimont} & Individual/Pair interaction at school         & 11 min/group               & 7 (individual) + 7 $\times$ 2 (pairs) & 7-8        \\
+{\it Diego}     & Case-study, spanning over 4 weeks              & 1.5 hours $\times$ 4 weeks & 1                                     & 5          \\
+{\it Haut-Lac}  & Pair (with turn-taking) interaction at school & 26 min/group                            & 7 $\times$ 2                          & 5-6        \\
+{\it ISG 2}     & Individual interaction at school              & 18 min                            & 6                                     & 5-6        \\
+{\it Henry}     & Case-study, spanning over 4 weeks              & 1h $\times$ 4 weeks        & 1                                     & 5          \\ \bottomrule
+\end{tabular}
+\end{table}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Studies at School}
+\subsection{Studies at School}
 
@@
     \caption{The first field studies were focused on technical validation, with
-    above 100 pupils interacting with the robot over short periods (around 20
+    above 70 pupils interacting with the robot over short periods (around 15
 minutes), either alone or in small groups.}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Case Study 1: Diego}
+\subsection{Case Study 1: Diego}
 
-\subsection{Context, Study Design}
+\subsubsection{Context, Study Design}
 

### 2015-07-06 main.tex
Updated figure of experimental setup with camera and second tablet

@@
     \centering
-    \includegraphics[width=0.9\linewidth]{experimental_setup}
+    \includegraphics[width=0.8\linewidth]{experimental_setup}
     \caption{Our experimental setup: face-to-face interaction with a Nao robot.
@@
     (either a therapist or an experimenter, depending on the studies), remains next
-    to the child to guide the work (prompting, turn taking, etc.).}
+    to the child to guide the work (prompting, turn taking, etc.). For some studies, a second tablet and an additional camera (dashed) are employed.}
     \label{experimental_setup}

### 2015-07-06 main.tex
Added better diagram for Henry's learning of the 6

@@
     \centering
-    \includegraphics[width=0.9\linewidth]{learning_6}
-    \caption{Progresses made by Henry over one 45-minutes interaction session
-    while writing ``6''.}
-    \label{}
+    \includegraphics[width=0.9\linewidth]{learning_6_demos}
+    \caption{Henry's demonstrations for the number ``6'' (top row) and
+        corresponding shapes generated by the robot. After eight demomstrations,
+        Henry decided that the robot's ``6'' was good enough, and went to
+    another character: in that respect, he was the one leading the learning
+process of the robot.}
+    \label{henry_demos}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{learning_6_distances}
+    \caption{Two metrics to assess the handwriting progresses: Euclidian
+    distance in the Eigenspace of the number dataset (top figure) or in
+cartesian space (bottom figure). Green lines represent the robot performance,
+blue lines Henry's performance. The round IDs correspond to the demonstrations
+pictured on Figure~\ref{henry_demos}.}
+    \label{henry_distances}
 \end{figure}

### 2015-07-07 main.tex
WIP on intro

@@
 Henry is five and an half, and has been diagnosed with visuo-constructive
-deficits. He is under the care of an occupational therapist, and tries to improve
-his inability to draw letters in a consistent manner. Diego is six and struggles
-at school with his poor handwriting and even poorer self-confidence.
+deficits. He is under the care of an occupational therapist, and tries to
+workaround his inability to draw letters in a consistent manner. Diego is six
+and struggles at school with his poor handwriting and even poorer
+self-confidence.
 
@@
 activity to another, Diego is shy and poised. Two very different children,
-facing however the same difficulty to write in a legible manner.
+facing however the same difficulty to write in a legible manner. And, hidden
+behind this impaired skill, psycho-social difficulties also arise: they
+underperform at school, Henry has to go for follow-up visits to a therapist,
+they require ``special care''. This is a source of anxiety, for the children,
+and for their parents alike.
 
@@
 
-Remediations for handwriting difficulties
+Remediations for handwriting difficulties traditionally involve
+long interventions (at least 10 weeks~\cite{Hoy2011}), essentially handwriting
+training with an occu therapist, and primarly address the motor deficits.
+Improvements in self-confidence, anxiety occur (at best) as a consequence of the
+child improving his/her handwriting skills and improving his/her performances at
+school (when not too late).
+
+We present in this article a new take on this educative challenge, as a
+remediation procedure that involve a bad-writer robot that is taught by the
+child: by building on the \emph{learning by teaching} paradigm, not only the
+child practises hand-writing but, as he/she takes on the role of the teacher,
+he/she also positively reinforce his/her self-esteem and motivation (he/she
+shifts from the ``underperformer'' to become ``the one who knows and teaches'').
 
@@
 
+\paragraph{Engagement levels}
+
+A conclusion drawn in a systematic review of classical handwriting
+interventions~\cite{Hoy2011} is that any of the studies considered which
+involved fewer than two practice sessions per week and fewer than a total of 20
+practice sessions, including homework, were not found to demonstrate effective
+results. This highlights the necessity to engage students in an interaction
+which will be sustainable over the long-term if we want to address research
+questions which involve the measurement of learning gains.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-07-07 main.tex
WIP on conclusion

@@
 induce -- can convincingly play.
+
+\subsection{Agency and Commitment}
+
+However, we believe that the strongest impact of this work is for the
+human-robot interaction community and relates to the very \emph{nature} of the
+interaction fostered by this research. The work presented here investigates a
+particular role for a robot in the education of handwriting: not only is the
+robot actively performing the activity by drawing letters, but it does so in a
+way that engages the child in a very specific social role. The child is the
+teacher in this relationship and the robot is the learner: the child must engage
+in a (meta-) cognitive relationship with the robot to try to understand why the
+robot fails and how to help it best.  Here, the robot is more than just an
+activity facilitator or orchestrator -- its physical presence and embodiment
+induce agency and anthropomorphising, and cognitively engage the child into the
+learning activity, which we predict will lead to higher learning efficacy.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Agency and Commitment}
+\section{Conclusion}
 
-However, we believe that the strongest impact of this work is for the
-human-robot interaction community and relates to the very \emph{nature} of the
-interaction fostered by this research. The work presented here investigates a
-particular role for a robot in the education of handwriting: not only is the
-robot actively performing the activity by drawing letters, but it does so in a
-way that engages the child in a very specific social role. The child is the
-teacher in this relationship and the robot is the learner: the child must engage
-in a (meta-) cognitive relationship with the robot to try to understand why the
-robot fails and how to help it best.  Here, the robot is more than just an
-activity facilitator or orchestrator -- its physical presence and embodiment
-induce agency and anthropomorphising, and cognitively engage the child into the
-learning activity, which we predict will lead to higher learning efficacy.
+The CoWriter project, as presented above, gives a novel perspective on educative
+robots, at several levels:
+
+\begin{itemize}
+    \item robots in an educative context are \textbf{relevant and effective beyond ICT
+        teaching},
+
+    \item we can successfully \textbf{transpose the well-established
+        \emph{learning by teaching} paradigm} from education sciences to
+        robotics, even in a \textbf{complex form}: handwriting is difficult
+        physical skill, the robot learns and interact autonomously, the child is
+        responsible not only for the teaching but also for the teaching
+        orchestration by managing the turn taking and the progression of the
+        activity,
+
+    \item blending machine-learning techniques with human-robot interaction
+        allows for building a \textbf{believable agent}, that induces
+        \textbf{social commitment},
+
+    \item this proves essential to \textbf{sustain a long-term interaction} (several
+        hours) around a fundamentally routine -- yet challenging -- educational
+        task (hand-writing learning).
+\end{itemize}
 

### 2015-07-07 main.tex
WIP conclusion: role of teachers

@@
 
+No surprinsingly, the role of the robot \textit{vis-à-vis} the teacher has also
+been questionned (including an angry email from one teacher: ``You want to
+replace us!''): as we see it, the role of the robot within the classroom (or at
+the therapist's surgery) does not infringe upon the role of the adult (teacher
+or therapist).  The core of the \emph{learning by teaching} paradigm relies on
+the child becoming the teacher of an underperforming pupil (the robot): from
+that perspective, the robot does not replace the teacher, on the contrary. It
+plays a different role in the classroom, which happen to be novel as well: the
+robot is the \emph{least} performing student, and however a very patient, always
+eager to improve, one.
+
+The importance of the adult is further supported by our experiments: even with
+an autonomous, nominally performing robot, to put the teacher's clothes on
+remains (expectedly) difficult for 5-6 years old children, and during the
+experiments we conducted, the adult always played a key role at prompting the
+child to give feedback to the robot or to go to the next letter or word.  At a
+higher orchestration level (and as reported in the two case studies with Diego
+and Henry), the educational scenarios were also always designed and monitored by
+the adults.
+
+We initially envisionned the CoWriter system to be run in the back of a
+classroom with one child: this would have allowed an individual, face-to-face
+remediation approach, not otherwise tractable for a teacher with 20 pupils.
+This is unlikely to happen. In our experience, the teacher keeps such an
+important role that the interaction and the learning would hardly occur if the
+child is left alone (or even semi-alone). At the end, we can reassure teachers:
+robots are not going to replace them any time soon.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-07-07 main.tex
Diego : context,study design

@@
 \subsubsection{Context, Study Design}
+It was the first time we where trying a long-terme interaction between a child
+and the CoWriter robot. So the first aim of this study was to see if such an
+interaction was simply possible. If we could create an environement to keep a child engaged
+during one hour just in writing words with the robot.
+The second question was focused on the content of the interaction. Our goal was
+to figure out what extent the child would actually improve the robot's
+writing.\\
+
+Diego is a 5 years hold child. 
+Before the experiment, we could create a dataset of letters based on his main
+mistakes, but amplifiyed by PCA. Our idea was to make the robot
+exaggerating the mistakes of Diego. Thus, by correcting the robot, 
+Diego was actually going to correct himself.\\
+
+The experiment took one month. It was divided in four sessions of one hour, one session per week.
+In order to justify to the child an activity where a robot wants to learn
+handwriting, we decided to adduce a scenario. There where two Nao robots: a
+blue one (Mimi) and an orange one (Clem). They where introduced to the child as
+two hold friends. Mimi was left, and the two robots were corresponding by
+handwrited mails. But Clem was begining in writing and it needed the help of
+Diego to write its responses to Mimi.\\
+
+During the first session, Diego was just teaching single letters to the robot.
+progressively we increased the difficulty : the second session he was teaching
+shirt words and in third session he was teaching long sentences. The last session was more like a test for Clem
+: it had to prove to Mimi that it actually improved its handwriting thanks to
+Diego.
 

### 2015-07-07 main.tex
Article overview

@@
 learning activity, which we predict will lead to higher learning efficacy.
+
+\subsection*{Overview of the article}
+
+We have sofar frame our intents: to design a robotic system that would support
+handwriting remediation in an original way: by getting children to teach a
+robotic agent how to write, those children would both practise without knowing
+it and recover self-confidence and self-esteem by supporting a
+worst-than-themselves robotic ``student''.
+
+The following sections of the paper go into details. The next section provides a
+brief overview of the robotic system and the interaction it induces. We then
+give the details of machine-learning techniques that allow the robot to learn
+from the children, followed by the implementation on the Nao robot.
+
+Next, we present and report on the four school experiments and the two case
+studies that we conducted over the last two years. While the focus of the school
+experiments was mostly technical validation and data acquisition, the two case
+studies that lasted each about 4 weeks, gave us initial insights on the
+relevance and effectiveness of our approach.
 

### 2015-07-07 main.tex
Henry context

@@
 
-Diego is a 5 years hold child. 
+Diego is a 6 years hold child. 
 Before the experiment, we could create a dataset of letters based on his main
@@
 \subsubsection{Context, Study Design}
+Henry is a five and half years hold child. It is followed by an occupational
+therapist. According to the therapist, he presents visuo-constructive deficits.
+From our perspective, the main effect of this trouble during handwriting
+activity was that Henry
+did not have any mental template to draw the shape of a letter. So he was 
+apparently performing random attempts and then was comparing with the provided template.
+What is more, Henry is strongly careless : he rarely payed attention to
+advices, even to what he was doing when he was currently drawing, and he was
+quickly shifting his attention from one
+activity to another.
+Hence, the goal of this study is to face those two chalenging issues
+(visuo-constructive deficits and inattention) in order to keep Henry focused
+on the activity during forty-minutes interactions, and to make the robot
+evidently learning from his demonstrations.\\
+
+This experiment took place in the therapist's office. It was divided in four
+sessions as well. This time, we assumed that a complex cenario like the one for Diego was
+no longer relevent with Henry. We just introduced the robot and quickly
+said that it was seeking help to train for a robot handwriting contest.\\
 

### 2015-07-07 main.tex
Added \small to captions

@@
     \includegraphics[width=0.9\linewidth]{henry}
-    \caption{Henry teaching Nao how to write numbers, with the help of an
+    \caption{\small Henry teaching Nao how to write numbers, with the help of an
     occupational therapist.}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{experimental_setup}
-    \caption{Our experimental setup: face-to-face interaction with a Nao robot.
-    The robot writes on the tactile tablet, the child then corrects the robot by
-    directly overwriting its letters on the tablet, with a stylus directly. An adult
-    (either a therapist or an experimenter, depending on the studies), remains next
-    to the child to guide the work (prompting, turn taking, etc.). For some studies, a second tablet and an additional camera (dashed) are employed.}
+    \includegraphics[width=0.6\columnwidth]{experimental_setup}
+    \caption{\small Our experimental setup: face-to-face interaction with a {\sc nao} robot.
+        The robot writes on the tactile tablet, the child then corrects the
+        robot by directly overwriting its letters on the tablet, with a stylus
+        directly. An adult (either a therapist or an experimenter, depending on
+        the studies), remains next to the child to guide the work (prompting,
+        turn taking, etc.). For some studies, a second tablet and an additional
+        camera (dashed) are employed.}
+
     \label{experimental_setup}
@@
     \end{center}
-    \caption{A demonstration of the robot simulating the writing of a word with
+    \caption{\small A demonstration of the robot simulating the writing of a word with
     its finger. The motion of the robot is synchronised with the display of the
@@
 \centering
-\caption{Field studies conducted within the project}
+\caption{\small Field studies conducted within the project}
 \label{studies}
@@
     \includegraphics[height=3.9cm]{schools2}
-    \caption{The first field studies were focused on technical validation, with
+    \caption{\small The first field studies were focused on technical validation, with
     above 70 pupils interacting with the robot over short periods (around 15
@@
     \includegraphics[width=0.5\linewidth]{diego}
-    \caption{Diego working with Nao}
+    \caption{\small Diego working with Nao}
     \label{fig:diego}
@@
     \includegraphics[width=0.9\linewidth]{abc-writer}
-    \caption{Screenshot of the Android application developed to be used as
+    \caption{\small Screenshot of the Android application developed to be used as
     pre-test and post-test: the child must follow with his finger the path of
@@
     \includegraphics[width=0.9\linewidth]{learning_6_demos}
-    \caption{Henry's demonstrations for the number ``6'' (top row) and
+    \caption{\small Henry's demonstrations for the number ``6'' (top row) and
         corresponding shapes generated by the robot. After eight demomstrations,
@@
     \includegraphics[width=0.9\linewidth]{learning_6_distances}
-    \caption{Two metrics to assess the handwriting progresses: Euclidian
+    \caption{\small Two metrics to assess the handwriting progresses: Euclidian
     distance in the Eigenspace of the number dataset (top figure) or in

### 2015-07-07 main.tex
Working on the technical implementation

@@
 
-In order to develop a teachable agent that is appropriate for engaging a child
-in the learning by teaching paradigm, we have established capabilities for the
-robot to engage in handwriting and interactive turn-taking.
+Our system is embodied through an Aldebaran's {\sc nao} (V4 or V5, depending on the studies)
+humanoid robot. This choice is motivated by its approachable
+design~\cite{Gouaillier2008}, its size (58cm) and inherently safe structure (lightweight
+plastic) (making it suitable for close interaction with children), its low price
+(making it closer to what school may afford in the coming years) and finally its
+ease of deployment on the field.
 
-The {\sc nao} V4 humanoid robot, which has been purposely designed by Aldebaran
-Robotics to look approachable~\cite{Gouaillier2008}, is used for this work. It
-is a commercially affordable biped robot, 58cm tall, with 25 degrees of freedom,
+{\sc nao} is a humanoid biped, has 25 degrees of freedom,
 two cameras, speech capabilities and the ability to autonomously execute a range
-of tasks.
+of tasks. While it can walk, this capability is not used in this project, and
+the robot remains in a crouching posture during the whole interaction.
 
-Precise control over what the robot is writing is necessary in the proposed
-application of a handwriting. Because of the limited fine motor skills possible
+Robotic handwriting requires precise closed-loop control of the arm and hand
+motion. Because of the limited fine motor skills possible
 with such an affordable robot, in addition to the absence of force feedback and
-other technical necessities, we have configured the {\sc nao} to use simulated
-handwriting with a synchronised tablet to achieve this level of control.
+other technical necessities, we have opted for \emph{simulated
+handwriting}: the robot draws letters in the air, and the actual writing is
+displayed on a synchronised tablet.
 
@@
 humanoid agent are presented in the sections which follow.
+
+\begin{figure}[ht]
+\centering
+
+\resizebox{0.6\linewidth}{!}{%
+
+\begin{tikzpicture}[
+    >=latex,
+    node distance=2cm,
+    every edge/.style={draw, very thick},
+    redarrow/.style={draw=Red},
+    greenarrow/.style={draw=GreenYellow},
+    yellowarrow/.style={draw=BurntOrange},
+    cmpt/.style={draw, align=center, rounded corners, inner sep=5pt, font=\sf, fill=black!20},
+    label/.style={midway, align=left, font=\scriptsize\sf, fill=white, above,opacity=0,text opacity=1}]
+
+    \node at (0,0) (laptop) {\includegraphics[width=2cm]{laptop}};
+    \node[below right=2 of laptop] (nao) {\includegraphics[width=2cm]{nao}};
+    \node[below left=2 of laptop] (tablet) {\includegraphics[width=2cm]{tablet+stylus}};
+    \node[above=2 of laptop] (selection) {\includegraphics[width=2cm]{selection_tablet}};
+
+    \path (nao) edge [->,redarrow, bend left] node[label, auto] {robot state} (laptop);
+    \path (laptop) edge [->,greenarrow, bend left] node[label, auto] {writing gesture} (nao);
+
+    \path (tablet) edge [->,redarrow, bend left] node[label, auto] {demonstrations,\\turn taking} (laptop);
+    \path (laptop) edge [->,redarrow, bend left] node[label, auto=right] {letter to write} (tablet);
+
+    \path (selection) edge [->,redarrow] node[label, auto] {letter/word to write} (laptop);
+
+    \path (-5, 2) edge [->, redarrow] node[label] {ROS} ++(1, 0);
+    \path (-5, 2.6) edge [->, greenarrow] node[label] {NaoQI} ++(1, 0);
+
+\end{tikzpicture}
+}
+
+\caption{\small Overview of the system. In total, the system runs about 10 ROS nodes,
+    distributed over the robot itself, a central laptop and Android tablets. Since
+    the system does not require any CPU-intensive process, the laptop could be
+    removed and the whole logic could run on the robot. Due to the relative
+    difficulty to deploy and debug ROS nodes directly on the robot, we however
+    relied on the laptop during our experiments.}
+
+    \label{fig:archi}
+\end{figure}
+
+%
+%\begin{figure}[ht]
+%\centering
+%
+%\resizebox{0.9\linewidth}{!}{%
+%
+%\begin{tikzpicture}[
+%    >=latex,
+%    node distance=2cm,
+%    every edge/.style={draw, very thick},
+%    redarrow/.style={fill=Red, draw=Red},
+%    greenarrow/.style={fill=GreenYellow, draw=GreenYellow},
+%    yellowarrow/.style={fill=BurntOrange, draw=BurntOrange},
+%    cmpt/.style={draw, align=center, rounded corners, inner sep=5pt, font=\sf, fill=black!20},
+%    label/.style={midway, align=left, font=\scriptsize\sf, fill=white, above,opacity=0,text opacity=1}]
+%
+%    \node at (0,0)[cmpt, fill=Cyan!50, text width=2.7cm] (tactile) {tactile interaction \\ manager};
+%    \node[cmpt, fill=Cyan!50, left=of tactile] (display) {display \\ manager};
+%  
+%    \node [cmpt, fill=Green!50, below=of tactile, text width=2.7cm] (shape) {shape position \\ manager};
+%    \node [cmpt, fill=Green!50, left=of shape] (learner) {letter \\ learner};
+%    \node [cmpt, fill=Green!50, left=of learner] (card) {card \\ detector};
+%
+%    \node [cmpt, fill=YellowOrange!50, below=1.5cm of shape] (writer) {nao writer};
+%    \node [cmpt, fill=YellowOrange!50, left=of writer] (tablet) {tablet detector};
+%    \node [cmpt, fill=YellowOrange!50, left=of tablet] (camera) {nao camera};
+%    \node [cmpt, fill=YellowOrange!50, below=0.6cm of tablet] (pose) {robot state publisher};
+%
+%    %%% Relations between components
+%    \path (tactile) edge [->] node[label] {clear \\ screen} (display);
+%
+%    \path ($(shape.north west)!0.33!(shape.north east)$) edge [<-] node[label,left] {clear \\ screen} ($(tactile.south west)!0.33!(tactile.south east)$);
+%    \path ($(shape.north west)!0.66!(shape.north east)$) edge [<-, greenarrow] node[label,right] {feedback \\ + location} ($(tactile.south west)!0.66!(tactile.south east)$);
+%
+%    \path (shape) edge [->, redarrow] node[label, left, align=left] {letter \\to write \\ + time \\ + position} (display);
+%    \path (shape) edge [->,redarrow] node[label, right] {letter to write \\ + time \\ + position in\\tablet frame} (writer);
+%
+%    \path ($(shape.north west)!0.66!(shape.south west)$) edge [->, greenarrow] node[label,below] {feedback \\ + letter} ($(learner.north east)!0.66!(learner.south east)$);
+%    \path ($(shape.north west)!0.33!(shape.south west)$) edge [<-, redarrow] node[label] {letter to write \\ + time} ($(learner.north east)!0.33!(learner.south east)$);
+%
+%    \path (card) edge [->, yellowarrow] node[label] {word to learn} (learner);
+%    \path (camera) edge [->, yellowarrow] node[label, right] {camera image} (card);
+%    \path (camera) edge [->, redarrow] node[label] {camera image} (tablet);
+%    \path (tablet) edge [->, redarrow] node[label] {tablet pose} (writer);
+%    \path (pose) edge [->, redarrow] node[label, right] {head pose} (tablet);
+%
+%    \path (-8, 0) edge [->, redarrow] node[label] {Writing letters} (-7, 0);
+%    \path (-8, -0.6) edge [->, greenarrow] node[label] {Responding to feedback} (-7, -0.6);
+%    \path (-8, -1.2) edge [->, yellowarrow] node[label] {Detecting card input} (-7, -1.2);
+%    
+%\end{tikzpicture}
+%}
+%
+%\caption{\small Overview of the system. Components in the top row run on the tablet, 
+%those in the middle row on the central controller, and those in the bottom row 
+%on the robot.}
+%
+%    \label{fig:archi}
+%\end{figure}
+%
 
@@
 working space for writing on the entire tablet.
-
-\begin{figure}[ht]
-\centering
-
-\resizebox{0.9\linewidth}{!}{%
-
-\begin{tikzpicture}[
-    >=latex,
-    node distance=2cm,
-    every edge/.style={draw, very thick},
-    redarrow/.style={fill=Red, draw=Red},
-    greenarrow/.style={fill=GreenYellow, draw=GreenYellow},
-    yellowarrow/.style={fill=BurntOrange, draw=BurntOrange},
-    cmpt/.style={draw, align=center, rounded corners, inner sep=5pt, font=\sf, fill=black!20},
-    label/.style={midway, align=left, font=\scriptsize\sf, fill=white, above,opacity=0,text opacity=1}]
-
-    \node at (0,0)[cmpt, fill=Cyan!50, text width=2.7cm] (tactile) {tactile interaction \\ manager};
-    \node[cmpt, fill=Cyan!50, left=of tactile] (display) {display \\ manager};
-
-    \node [cmpt, fill=Green!50, below=of tactile, text width=2.7cm] (shape) {shape position \\ manager};
-    \node [cmpt, fill=Green!50, left=of shape] (learner) {letter \\ learner};
-    \node [cmpt, fill=Green!50, left=of learner] (card) {card \\ detector};
-
-    \node [cmpt, fill=YellowOrange!50, below=1.5cm of shape] (writer) {nao writer};
-    \node [cmpt, fill=YellowOrange!50, left=of writer] (tablet) {tablet detector};
-    \node [cmpt, fill=YellowOrange!50, left=of tablet] (camera) {nao camera};
-    \node [cmpt, fill=YellowOrange!50, below=0.6cm of tablet] (pose) {robot state publisher};
-
-    %%% Relations between components
-    \path (tactile) edge [->] node[label] {clear \\ screen} (display);
-
-    \path ($(shape.north west)!0.33!(shape.north east)$) edge [<-] node[label,left] {clear \\ screen} ($(tactile.south west)!0.33!(tactile.south east)$);
-    \path ($(shape.north west)!0.66!(shape.north east)$) edge [<-, greenarrow] node[label,right] {feedback \\ + location} ($(tactile.south west)!0.66!(tactile.south east)$);
-
-    \path (shape) edge [->, redarrow] node[label, left, align=left] {letter \\to write \\ + time \\ + position} (display);
-    \path (shape) edge [->,redarrow] node[label, right] {letter to write \\ + time \\ + position in\\tablet frame} (writer);
-
-    \path ($(shape.north west)!0.66!(shape.south west)$) edge [->, greenarrow] node[label,below] {feedback \\ + letter} ($(learner.north east)!0.66!(learner.south east)$);
-    \path ($(shape.north west)!0.33!(shape.south west)$) edge [<-, redarrow] node[label] {letter to write \\ + time} ($(learner.north east)!0.33!(learner.south east)$);
-
-    \path (card) edge [->, yellowarrow] node[label] {word to learn} (learner);
-    \path (camera) edge [->, yellowarrow] node[label, right] {camera image} (card);
-    \path (camera) edge [->, redarrow] node[label] {camera image} (tablet);
-    \path (tablet) edge [->, redarrow] node[label] {tablet pose} (writer);
-    \path (pose) edge [->, redarrow] node[label, right] {head pose} (tablet);
-
-    \path (-8, 0) edge [->, redarrow] node[label] {Writing letters} (-7, 0);
-    \path (-8, -0.6) edge [->, greenarrow] node[label] {Responding to feedback} (-7, -0.6);
-    \path (-8, -1.2) edge [->, yellowarrow] node[label] {Detecting card input} (-7, -1.2);
-
-\end{tikzpicture}
-}
-
-\caption{Overview of the system. Components in the top row run on the tablet, 
-those in the middle row on the central controller, and those in the bottom row 
-on the robot.}
-
-    \label{fig:archi}
-\end{figure}
 

### 2015-07-07 main.tex
Henry study design

@@
 
-Diego is a 6 years hold child. 
+Diego is a six years old child. 
 Before the experiment, we could create a dataset of letters based on his main
@@
 \subsubsection{Context, Study Design}
-Henry is a five and half years hold child. It is followed by an occupational
+Henry is a five and half years old child. It is followed by an occupational
 therapist. According to the therapist, he presents visuo-constructive deficits.
@@
 
+Before the experiment, Henry was working on writing numbers with the therapist.
+Hence we decided to turn the CoWriter activity to teach numbers to the robot. Build on
+existing iPad application actually used by occupational therapists (Dawson Toth's \emph{ABC's
+Writer}), we designed an Android application for pre-test and post-test. It consisted in drawing
+numbers following an helping pattern, that becomes more fine with levels in
+order to increase difficulty~\ref{fig:abc-writer}.\\
+
+Since Henry was frequently drawing horizontaly inverted numbers, or even
+absolutly unrecognizable shapes, we programmed the robot to refuse shapes that
+was too distant to a reference with a threshold we arbitrary fixed. In that way,
+the child was forced to take care on what he was providing to the robot as
+demonstration. Also, to make sure the robot was going to improve its handwriting
+and to clearly show this improvement, we decided to make it completly starting
+from scratch : for all numbers, the first try of the robot resulted in
+a simple vertical stroke (see the first robot's try in~\ref{learning_6_demos}).
+
 \begin{figure}
@@
 the red-bordered area.}
-    \label{}
+    \label{fig:abc-writer}
 \end{figure}
@@
 process of the robot.}
-    \label{henry_demos}
+    \label{learning_6_demos}
 \end{figure}

### 2015-07-07 main.tex
More work on technical implementation

@@
 
-The development of the necessary components for embedding the handwriting
-learning algorithm presented in Section \ref{sec:learningAlgorithm} in the
-humanoid agent are presented in the sections which follow.
-
-\begin{figure}[ht]
+\begin{figure}[ht!]
 \centering
@@
 \caption{\small Overview of the system. In total, the system runs about 10 ROS nodes,
-    distributed over the robot itself, a central laptop and Android tablets. Since
-    the system does not require any CPU-intensive process, the laptop could be
-    removed and the whole logic could run on the robot. Due to the relative
-    difficulty to deploy and debug ROS nodes directly on the robot, we however
-    relied on the laptop during our experiments.}
+    distributed over the robot itself, a central laptop and Android tablets.}
 
@@
 \end{figure}
+
+The overall architecture of the system (Figure~\ref{fig:archi}) is therefore
+spread over several devices: the {\sc nao} robot itself, that we address via
+both a ROS API\footnote{The ROS stack for {\sc nao} is available at
+\url{http://wiki.ros.org/nao_robot}.} and the Aldebaran-provided NaoQI API, one
+to four Android tablets (the main tablet is used to print the robot's letter and
+to acquire the children' demonstrations. More tablets have been used in some
+studies, either to let the child input words to be written, or for the
+experimenter to qualitatively annotate the interaction in a synchronized
+fashion), and a central laptop running the machine learning algorithms, the
+robot's handwriting gesture generation and high level control of the activity.
+
+Since the system does not actually require any CPU-intensive process, the laptop
+could be removed and the whole logic could run on the robot. Due to the relative
+difficulty to deploy and debug ROS nodes directly on the robot, we however
+effectively kept the laptop during our experiments.
+
+Most of the nodes are written in Python, and the whole source code of the
+project is available online\footnote{The primary repository is
+\url{https://github.com/chili-epfl/cowriter_letter_learning}.}
 
@@
 
-\subsubsection{Robot Trajectory Following Movements}
+\subsubsection*{Trajectory Following Movements}
 
@@
 capture the engagement of the child participant in the action. Aldebaran's NaoQi
-API is used for the inverse kinematics of the trajectory following. The Robot
-Operating System (ROS)\footnote{The ROS stack for {\sc nao} is available at:
-\url{http://wiki.ros.org/nao_robot}.} is used for integration of the {\sc nao}
-with external reference frames, such as the tablet's location, using the $tf$
-transformation library \cite{Foote2013}.
+API is used for the inverse kinematics of the trajectory following. ROS is used
+for integration of {\sc nao} with external reference frames, such as the
+tablet's location, using the $tf$ transformation library~\cite{Foote2013}.
 
@@
 engages in the typical style of handwriting of using a writing instrument at a
-desk.  Having the robot point at a vertical writing surface to cause the
-trajectory to appear (as in Figure \ref{fig:naoWriting}) has several advantages:
+desk. Having instead the robot point at the writing surface to cause the
+trajectory to appear (as seen in Figure \ref{fig:schools}, left) has several advantages:
 
@@
         the robot from across the room and have it still respond, without
-        needing the tablet to be within arm's reach.
+        needing the tablet to be within arm's reach,
 
     \item Concerns about whether or not the child would start mimicking the
-        robot's incorrect writing form (\eg pen grip) are mitigated.
+        robot's incorrect writing form (\eg pen grip) are mitigated,
 
@@
 \end{itemize}
-
-\begin{figure}[thpb]
-     \begin{center}
-            \includegraphics[width=0.7\linewidth]{naoWriting2.png}
-    \end{center}
-    \caption{\small A demonstration of the robot simulating the writing of a word with
-    its finger. The motion of the robot is synchronised with the display of the
-    tablet, communicating over ROS.}%
-
-   \label{fig:naoWriting}
-\end{figure}
+%
+%\begin{figure}[thpb]
+%     \begin{center}
+%            \includegraphics[width=0.7\linewidth]{naoWriting2.png}
+%    \end{center}
+%    \caption{\small A demonstration of the robot simulating the writing of a word with
+%    its finger. The motion of the robot is synchronised with the display of the
+%    tablet, communicating over ROS.}%
+%
+%   \label{fig:naoWriting}
 

### 2015-07-07 main.tex
Learning letters via PCA

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Algorithmic Approach}
+\subsection{Generating and Learning Letters}
+
+Since our application is about teaching a robot to write, generating (initially
+bad) letters and learning from demonstrations is a core aspect of the project.
+
+The main insight for both the generation and the learning of letters is to
+reason in the eigenspace of the letters, instead of the cartesian space. The
+eigen space of each letters is spanned by the first $n$ eigenvectors (in our
+experiments, $3 < n < 6$) of the covariance matrix observed in a standard
+dataset of adult letters (the UJI Pen Characters 2 dataset~\cite{Llorens2008}).
+This procedure, based on a Principle Component Analysis (PCA), is explained in
+details in a previously published article~\cite{hood2015when}.
+
+By changing the eigenvalues associated to these eigenvectors and regenerating
+letters with the reverse procedure, it becomes then easy to generate new
+letters, with distortions that are actually plausible handwriting errors: they
+are actually \emph{exaggerations} of variance of writing styles that naturally
+exist amongst adult writers.  Figure~\ref{fig:sampleLetters} shows examples of
+deformed ``e'' and ``g'' generated with such a technique.
 
@@
 \centering
-    \includegraphics[width=0.6\columnwidth]{learningSdemo}
-    \caption{\label{fig:demonstrationShapes2}Example of the learning algorithm
-    responding (top) to user demonstration of shapes (bottom) for the letter `s'
-    (demonstrations received from two 7-8 year-old children taking turns).}
+\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.08\columnwidth]{e1.png}}}
+\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.08\columnwidth]{e4.png}}}
+\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.08\columnwidth]{e3.png}}}
+\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.08\columnwidth]{e5.png}}}~~~~~
+\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.06\columnwidth]{g1.png}}}
+\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.06\columnwidth]{g3.png}}}
+\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.06\columnwidth]{g4.png}}}
+\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.06\columnwidth]{g5.png}}}
+\caption{\small \label{fig:sampleLetters}Sample letters generated from the PCA shape
+model on ``e'' (left) and ``g'' paths (right), generated randomly from
+parameters with $3\times$ the standard deviation observed in the dataset.}
 
@@
 
-\subsubsection{Path Manipulation in the Letters' Eigenspace}
+The same technique can be applied to \emph{classify} demonstrations,
+\emph{assess} their quality and
+\emph{learn} from them. Figure~\ref{fig:h} shows 9 allographs of ``h'' written by a 6
+years old child, along with the mean letter from the
+dataset (reference letter). By projecting each of the demonstrations onto the eigenspace
+of ``h'' (Figure~\ref{fig:h}b), we observe that:
+
+\begin{itemize}
+    \item the different allographs can by clustered (with a $k$-means or
+        mean-shift algorithm) by their visual styles,
+    \item we can compute an euclidian distance to the reference letter to assess
+        the visual proximity of the demonstration with the expected letter, thus
+        providing a quantitative metric of writing performance.
+\end{itemize}
 
@@
     \centering
-    \subfigure[Nine samples of the cursive ``h'', next to the reference]{
-        \includegraphics[height=6cm]{h}
+    \subfigure[Nine allographs of the cursive ``h'', next to the reference]{
+        \includegraphics[height=4.1cm]{h}
     }
-    \subfigure[Raw samples projected in the Eigenspace of the dataset]{
-        \includegraphics[height=4cm]{eigenspace-base}
-    }
-    \subfigure[Normalized samples, projected in the Eigenspace. The letters are
-    clustered according to their styles.]{
-        \includegraphics[height=4cm]{eigenspace-uniformization}
+    \subfigure[Same samples, normalized and projected in the eigen space spanned from the
+        3 first eigenvectors: clusters arise, that actually match writing styles.]{
+        \includegraphics[height=4.5cm]{eigenspace-uniformization}
     }
 
-    \caption{\small Projecting demonstrated letters onto the Eigenspace
+    \caption{\small Projecting demonstrated letters onto the eigenspace
     generated from the reference dataset effectively clusters the samples
-according to their topological similarity.}
-    \label{fig:stimuli}
+    according to their topological similarity. Allographs that are visually similar to
+    the reference are close to it in the eigenspace.}
+    \label{fig:h}
 \end{figure}
+
+The algorithm for machine-learning becomes then a simple matter of converging at
+a specific pace towards the child demonstration in the eigenspace.
+Figure~\ref{learning_6_demos}, at the end of the article, shows a complete learning cycle
+with the number ``6''.
 
@@
 
-No surprinsingly, the role of the robot \textit{vis-à-vis} the teacher has also
+No surprinsingly, the role of the robot \textit{vis-a-vis} the teacher has also
 been questionned (including an angry email from one teacher: ``You want to

### 2015-07-07 main.tex
Added figure on engagement level
Diego Results
need to count demo-letters

@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{engagement}
+    \caption{Engagement level of the child, manually annotated by two judges
+        over the 13 min long interaction (Cronbach’s alpha test: $\alpha = .82$,
+        good inter-judge agreement). The vertical lines represent activity
+    switches between handwriting teaching and story-telling (used as a
+    distruber).}
+    \label{}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsubsection{Results}
+Diego completly accepted our scenario and seriously played the game. At the
+begining he seemed not confident and intimidated.  But quickly, he gain in
+confidence and complicity with the robot. Finally, he showed a real affinity, so
+that he sent an handwrited mail to it some days after the last
+session.\\
+
+In all sessions, Diego was fully focused on the activity during forty minutes up to one
+hour. In total, he provided X letters as demonstration.\\
+
+Diego felt he improved the robot's handwriting. To make clearly visible this
+improvement, we printed the letter that the robot writted during the third
+session before and after the help of diego~\ref{fig:stimuli}.
 

### 2015-07-08 main.tex
First version of abstract

@@
 \begin{abstract}
-    TDB
+
+    Robots for education are not limited to support ICT teaching, and robots are
+    finding new ways to enter and blend in the classroom. This article reports
+    on such a new paradigm for educative robots, that involves \emph{learning by
+    teaching} and strong \emph{social engagement} to help children struggling
+    with handwriting. Our system relies on machine-learning and child-robot
+    interaction with a small humanoid robot, and we present several real-world
+    studies in schools and with occupational therapists that led us to promising
+    initial results.
+
 \end{abstract}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A different paradigm for educative robots}
+\section{A Different Paradigm for Educative Robots}
 
@@
 \subsubsection{Context, Study Design}
-It was the first time we where trying a long-terme interaction between a child
+It was the first time we where trying a long-term interaction between a child
 and the CoWriter robot. So the first aim of this study was to see if such an
-interaction was simply possible. If we could create an environement to keep a child engaged
+interaction was simply possible. If we could create an environment to keep a child engaged
 during one hour just in writing words with the robot.
@@
 Before the experiment, we could create a dataset of letters based on his main
-mistakes, but amplifiyed by PCA. Our idea was to make the robot
+mistakes, but amplified by PCA. Our idea was to make the robot
 exaggerating the mistakes of Diego. Thus, by correcting the robot, 
@@
 two hold friends. Mimi was left, and the two robots were corresponding by
-handwrited mails. But Clem was begining in writing and it needed the help of
+handwritten mails. But Clem was beginning in writing and it needed the help of
 Diego to write its responses to Mimi.\\
@@
 During the first session, Diego was just teaching single letters to the robot.
-progressively we increased the difficulty : the second session he was teaching
+Gradually we increased the difficulty : the second session he was teaching
 shirt words and in third session he was teaching long sentences. The last session was more like a test for Clem
@@
 \subsubsection{Results}
-Diego completly accepted our scenario and seriously played the game. At the
-begining he seemed not confident and intimidated.  But quickly, he gain in
+Diego completely accepted our scenario and seriously played the game. At the
+beginning he seemed not confident and intimidated.  But quickly, he gain in
 confidence and complicity with the robot. Finally, he showed a real affinity, so
-that he sent an handwrited mail to it some days after the last
+that he sent an handwritten mail to it some days after the last
 session.\\
@@
 In all sessions, Diego was fully focused on the activity during forty minutes up to one
-hour. In total, he provided X letters as demonstration.\\
+hour. In total, he provided 154 letters as demonstration.\\
 
 Diego felt he improved the robot's handwriting. To make clearly visible this
-improvement, we printed the letter that the robot writted during the third
-session before and after the help of diego~\ref{fig:stimuli}.
+improvement, we printed the letter that the robot wrote during the third
+session before and after the help of Diego~\ref{fig:stimuli}.
 
@@
 activity to another.
-Hence, the goal of this study is to face those two chalenging issues
+Hence, the goal of this study is to face those two challenging issues
 (visuo-constructive deficits and inattention) in order to keep Henry focused
@@
 This experiment took place in the therapist's office. It was divided in four
-sessions as well. This time, we assumed that a complex cenario like the one for Diego was
-no longer relevent with Henry. We just introduced the robot and quickly
+sessions as well. This time, we assumed that a complex scenario like the one for Diego was
+no longer relevant with Henry. We just introduced the robot and quickly
 said that it was seeking help to train for a robot handwriting contest.\\
@@
 
-Since Henry was frequently drawing horizontaly inverted numbers, or even
-absolutly unrecognizable shapes, we programmed the robot to refuse shapes that
+Since Henry was frequently drawing horizontally inverted numbers, or even
+absolutely unrecognizable shapes, we programmed the robot to refuse shapes that
 was too distant to a reference with a threshold we arbitrary fixed. In that way,
@@
 demonstration. Also, to make sure the robot was going to improve its handwriting
-and to clearly show this improvement, we decided to make it completly starting
+and to clearly show this improvement, we decided to make it absolutely starting
 from scratch : for all numbers, the first try of the robot resulted in
@@
 \subsubsection{Results}
+Despite his inattention, Henry was able to keep on the activity during more than
+forty minutes in each session.\\
+
+As soon as Henry understood that the robot was only accepting recognizable
+shapes, he started to take care on the shapes he was providing to it. He used the
+clearing button to retry when he judged that his numbers where not going to be
+accepted by the robot. According to the therapist, it was the first time Henry
+was intrinsically correcting himself.\\
+
+Since the robot was learning from scratch, each time Henry succeeded to make his
+demonstrations accepted by it, the improvement of its handwriting was
+obvious~\ref{learning_6_demos},~\ref{henry_distances}.
+Therefor, Henry quickly enter in an self-rewarding game where the rewards were
+the improvement of the robot. In total, over the three last sessions he provided 99
+demonstrations of number shapes to the robot.
 
@@
     \caption{\small Henry's demonstrations for the number ``6'' (top row) and
-        corresponding shapes generated by the robot. After eight demomstrations,
+        corresponding shapes generated by the robot. After eight demonstrations,
         Henry decided that the robot's ``6'' was good enough, and went to

### 2015-07-08 main.tex
Henry Results

@@
 \begin{abstract}
-
-    Robots for education are not limited to support ICT teaching, and robots are
-    finding new ways to enter and blend in the classroom. This article reports
-    on such a new paradigm for educative robots, that involves \emph{learning by
-    teaching} and strong \emph{social engagement} to help children struggling
-    with handwriting. Our system relies on machine-learning and child-robot
-    interaction with a small humanoid robot, and we present several real-world
-    studies in schools and with occupational therapists that led us to promising
-    initial results.
-
+    TDB
 \end{abstract}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A Different Paradigm for Educative Robots}
+\section{A different paradigm for educative robots}
 
@@
 In all sessions, Diego was fully focused on the activity during forty minutes up to one
-hour. In total, he provided 154 letters as demonstration.\\
+hour. In total, he provided X letters as demonstration.\\
 
@@
 Therefor, Henry quickly enter in an self-rewarding game where the rewards were
-the improvement of the robot. In total, over the three last sessions he provided 99
+the improvement of the robot. In total, over the four sessions he provided X
 demonstrations of number shapes to the robot.

### 2015-07-08 main.tex
number of demonstrations
Changed figure for examples of bad letters

--- 2015-07-08 main.tex
@@
 exist amongst adult writers.  Figure~\ref{fig:sampleLetters} shows examples of
-deformed ``e'' and ``g'' generated with such a technique.
+deformed ``g'' generated with such a technique.
 
-\begin{figure}[thpb]
+\begin{figure}
     \centering
-\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.08\columnwidth]{e1.png}}}
-\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.08\columnwidth]{e4.png}}}
-\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.08\columnwidth]{e3.png}}}
-\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.08\columnwidth]{e5.png}}}~~~~~
-\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.06\columnwidth]{g1.png}}}
-\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.06\columnwidth]{g3.png}}}
-\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.06\columnwidth]{g4.png}}}
-\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.06\columnwidth]{g5.png}}}
-\caption{\small \label{fig:sampleLetters}Sample letters generated from the PCA shape
-model on ``e'' (left) and ``g'' paths (right), generated randomly from
-parameters with $3\times$ the standard deviation observed in the dataset.}
+    \includegraphics[width=0.9\linewidth]{cowriter-g}
+    \caption{\small \label{fig:sampleLetters} \textbf{Generating bad letters}:
+        effect of varying the first five eigenvalues (rows) of the shape model
+        of ``g'' by different factors (columns). The percentage of the total
+        variance in the dataset explained by each eigenvalue is shown      on
+        the appropriate row.}
 

### 2015-07-08 main.tex
WIP on intro + minor other changes

@@
 facing however the same difficulty to write in a legible manner. And, hidden
-behind this impaired skill, psycho-social difficulties also arise: they
-underperform at school, Henry has to go for follow-up visits to a therapist,
-they require ``special care''. This is a source of anxiety, for the children,
-and for their parents alike.
+beyond this impaired skill, psycho-social difficulties also arise: they
+underperform at school, Henry has to go for follow-up visits every week,...
+they live under the label ``requires special care''. This is a source of
+anxiety, for the children, and for their parents alike.
 
@@
 
-Remediations for handwriting difficulties traditionally involve
-long interventions (at least 10 weeks~\cite{Hoy2011}), essentially handwriting
-training with an occu therapist, and primarly address the motor deficits.
-Improvements in self-confidence, anxiety occur (at best) as a consequence of the
-child improving his/her handwriting skills and improving his/her performances at
-school (when not too late).
+Remediations for handwriting difficulties traditionally involve long
+interventions (at least 10 weeks, \cite{Hoy2011}), essentially consisting in
+handwriting training with occupational therapists, and primarily addressing the
+\emph{motor} deficits.  Improvements in self-confidence and anxiety occur (at
+best) as a side-effect of the child improving his/her handwriting skills and,
+consequently, improving his/her performance at school.
 
 We present in this article a new take on this educative challenge, as a
-remediation procedure that involve a bad-writer robot that is taught by the
+remediation procedure that involve a \emph{bad writer} robot that is taught by the
 child: by building on the \emph{learning by teaching} paradigm, not only the
-child practises hand-writing but, as he/she takes on the role of the teacher,
-he/she also positively reinforce his/her self-esteem and motivation (he/she
-shifts from the ``underperformer'' to become ``the one who knows and teaches'').
+child practises hand-writing but, as (s)he takes on the role of the teacher,
+(s)he also positively reinforce his/her self-esteem and motivation: his/her
+social role shifts from the ``underperformer'' to ``the one who knows and teaches''.
 
@@
 
-The \emph{learning by teaching} paradigm, which engages the target student in
-the act of teaching another, has been shown to produce motivational,
-meta-cognitive, and educational benefits in a range of disciplines~\cite{Rohrbeck2003}. 
-To our best knowledge, the application of the paradigm to
-handwriting intervention remains, however, unexplored. One reason for this may
-due to the requirement of an appropriately unskilled peer for the target child
-to tutor, as this may present a logistical constraint if the target child is the
-lowest performer in their class.  In some cases, it may be appropriate for a
-peer or teacher to simulate a na\"ive learner for the target child to teach.
-For handwriting, where one's skill level is visually evident, however, this
-acting is likely to be eventually detected. As such, there is motivation for the
-use of a teachable agent which can be configured for a variety of skill levels,
-and for which children do not have preconceptions about its handwriting ability.
+The \emph{learning by teaching} paradigm, which engages the student in the act
+of teaching another, has been shown to produce motivational, meta-cognitive, and
+educational benefits in a range of disciplines~\cite{Rohrbeck2003}.  The
+application of this paradigm to handwriting intervention remains, however,
+unexplored. One reason for this may be due to the requirement of an
+appropriately unskilled peer for the child to tutor: it may prove difficult if
+the child is the lowest performer in his/her class.  In some cases, it may be
+appropriate for a peer or teacher to simulate a na\"ive learner for the target
+child to teach. For handwriting however, where one's skill level is visually
+evident, this acting is likely to be rapidly detected. This motivates the use of
+an artificial teachable agent which can be configured for a variety of skill
+levels, and for which children do not have preconceptions about its handwriting
+ability.
 
@@
 system, children learn vocabulary by teaching the {\sc nao} robot to act out
-verbs. The robot is tele-operated and mimics the actions that the children teach
-it, but with no long-term memory or learning algorithm in place.  Our project
-significantly extends this line of work in two ways. First, by investigating the
-context of children's acquisition of a challenging physical skill (handwriting),
-and second by proposing a robotic partner which is fully autonomous in its
-learning.
+verbs. The robot is tele-operated (Wizard-Of-Oz) and mimics the actions that the
+children teach it, but with no long-term memory or learning algorithm in place.
+Our project significantly extends this line of work in two ways. First, by
+investigating the context of children's acquisition of a challenging physical
+skill (handwriting), and second by proposing a robotic partner which is fully
+autonomous in its learning.
 
@@
 Since the child is assumed to take on the role of the teacher, we had to ensure
-he/she would be able to manage by him/herself the turn-taking and the overall
+(s)he would be able to manage by him/herself the turn-taking and the overall
 progression of the activity (moving to the next letter or word). In our design,
@@
     \includegraphics[width=0.6\columnwidth]{experimental_setup}
-    \caption{\small Our experimental setup: face-to-face interaction with a {\sc nao} robot.
-        The robot writes on the tactile tablet, the child then corrects the
-        robot by directly overwriting its letters on the tablet, with a stylus
-        directly. An adult (either a therapist or an experimenter, depending on
-        the studies), remains next to the child to guide the work (prompting,
-        turn taking, etc.). For some studies, a second tablet and an additional
-        camera (dashed) are employed.}
+    \caption{\small Our experimental setup: face-to-face interaction with a {\sc
+        nao} robot.  The robot writes on the tactile tablet, the child then
+        corrects the robot by directly overwriting its letters on the tablet
+        with a stylus. An adult (either a therapist or an experimenter,
+        depending on the studies), remains next to the child to guide the work
+        (prompting, turn taking, etc.). For some studies, a second tablet and an
+        additional camera (dashed) are employed.}
 
@@
     every edge/.style={draw, very thick},
-    redarrow/.style={draw=Red},
-    greenarrow/.style={draw=GreenYellow},
-    yellowarrow/.style={draw=BurntOrange},
+    redarrow/.style={draw,red, text=black},
+    greenarrow/.style={draw,GreenYellow,text=black},
+    yellowarrow/.style={draw,BurntOrange,text=black},
     cmpt/.style={draw, align=center, rounded corners, inner sep=5pt, font=\sf, fill=black!20},
@@
 to four Android tablets (the main tablet is used to print the robot's letter and
-to acquire the children' demonstrations. More tablets have been used in some
+to acquire the children's demonstrations. More tablets have been used in some
 studies, either to let the child input words to be written, or for the

### 2015-07-08 main.tex
Detailled examples of 'g'

@@
         variance in the dataset explained by each eigenvalue is shown      on
-        the appropriate row.}
+        the appropriate row. Examples noted A, B, C, D illustrate how the
+        PCA-based approach allows to \emph{automatically} generate letters whose
+        errors can be \emph{semantically} interpreted: A has a too large bottom
+        loop, B has a wide top loop, the bottom loop of C is not correctly
+        closed, the top loop of D is not closed.}
 

### 2015-07-08 main.tex
WIP intro

@@
 This research departs from the traditional approaches to educative robotics
-(\fixme{cite reviews})
+\cite{barker2007robotics,karim2015review}.
 
@@
 
-However, we believe that the strongest impact of this work is for the
-human-robot interaction community and relates to the very \emph{nature} of the
-interaction fostered by this research. The work presented here investigates a
-particular role for a robot in the education of handwriting: not only is the
-robot actively performing the activity by drawing letters, but it does so in a
-way that engages the child in a very specific social role. The child is the
-teacher in this relationship and the robot is the learner: the child must engage
-in a (meta-) cognitive relationship with the robot to try to understand why the
-robot fails and how to help it best.  Here, the robot is more than just an
-activity facilitator or orchestrator -- its physical presence and embodiment
-induce agency and anthropomorphising, and cognitively engage the child into the
-learning activity, which we predict will lead to higher learning efficacy.
+The work presented here investigates a particular role for a robot in the
+education of handwriting: not only is the robot actively performing the activity
+by drawing letters, but it does so in a way that engages the child in a very
+specific social role. The child is the teacher in this relationship and the
+robot is the learner: the child must engage in a (meta-) cognitive relationship
+with the robot to try to understand why the robot fails and how to help it best.
+Here, the robot is more than just an activity facilitator or orchestrator -- its
+physical presence and embodiment induce agency and anthropomorphising, and
+cognitively engage the child into the learning activity.
+
+Besides that, the commitment of the child into the interaction build on the
+``protégé effect'': teachable computer-based agents have previously been used to
+encourage the ``protégé effect'', wherein students invest more effort into
+learning when it is for a teachable agent than for themselves~\cite{Chase2009}.
+As we are concerned with learning of a physical skill, the learning agent
+developed is embodied in a humanoid robot which is capable of physically
+demonstrating handwriting trajectories to its child learning partner.  This is
+supported by the potential for motor mimicry to yield significant improvements
+in handwriting interventions in which letter formations are demonstrated to
+participants~\cite{Berninger1997}. Furthermore, when compared to screen-based
+agents, robotic partners have been shown in some contexts to increase users'
+compliance with tasks~\cite{Bainbridge2011}, maintain more effective long-term
+relationships~\cite{Kidd2008}, and produce greater learning gains when acting as
+tutors~\cite{Leyzberg2012}.
+
+For these two reasons, our approach is to be distinguished from previous works in
+educational robotic. Most of these do not consider the agency induced by the
+robot beyond its motivational aspect: playing with an interactive, partially autonomous
+device naturally induces some form of anthropomorphizing, which leads to some
+level of projected agency, and participate the overall excitement (at least, on
+the short-term, before the novelty effect vanishes).
+
+In our case, the role of agency is stronger: it induces meta-cognition (``I am
+interacting with an agent, so I need to reflect on how to best teach him''),
+which is beneficial for the learning process; it also induces a protégé effect
+(``I want my robot-agent to succeed!'') which support the commitment of child
+into the interaction, also for longer period of time (we report in this article
+upon to case studies where children were interacting with the same robot for several
+hours, spanned over several weeks).
 

### 2015-07-09 main.tex
Added Deanna and Aude as co-authors

@@
 
-\author{S\'everin Lemaignan, Alexis Jacq, Fernando Garcia, Ana Paiva, Pierre
-Dillenbourg \\ CHILI Lab, \'Ecole Polytechnique F\'ed\'erale de Lausanne}
+\author{S\'everin Lemaignan$^1$, Alexis Jacq$^{1,2}$, Fernando Garcia$^1$,
+    Deanna Hood$^1$, \\Aude
+    Billard$^1$, Ana Paiva$^2$, Pierre Dillenbourg$^1$ \\
+$^1$CHILI Lab, \'Ecole Polytechnique F\'ed\'erale de Lausanne, Suisse,\\
+$^2$Instituto Superior T\'{e}cnico, University of Lisbon, Portugal}
 
@@
 
-We would like to thank Deanna Hood for her key contribution to the project.
 This research was partially supported by the Funda\c{c}\~{a}o para a Ci\^{e}ncia

### 2015-07-09 main.tex
School studies section

@@
 
+Over the two years of the project, we conducted four studies in schools
+(Figure~\ref{fig:schools}). These experiments were meant to technically validate
+the system (is it actually able to autonomously write and learn from
+demonstrations?) and the interaction (is the apparatus easy to grasp and to
+interact with for children), and to study the initial acceptance of the robot in
+the school environment (we conducted several formal and informal discussions
+with teachers).
+
+We also relied on these four studies to acquire a dataset of children's
+handwritten letters and to study how they engage with the robot (and maintain or
+not this engagement).
+
 \begin{figure}
@@
 
-\paragraph{Engagement levels}
+Critically, these studies were conducted with \emph{whole classes}: we did not
+select underperforming children that requires careful beforehand organization
+with the school to limit ethical issues. Having more children (73 in total) was
+however beneficial for these preliminary studies.
 
-A conclusion drawn in a systematic review of classical handwriting
-interventions~\cite{Hoy2011} is that any of the studies considered which
-involved fewer than two practice sessions per week and fewer than a total of 20
-practice sessions, including homework, were not found to demonstrate effective
-results. This highlights the necessity to engage students in an interaction
-which will be sustainable over the long-term if we want to address research
-questions which involve the measurement of learning gains.
+\paragraph{System validation}
 
-\begin{figure}
+From a technical perspective, the system achieves an acceptable level of
+reliability and allow a technically sound autonomous interaction. For instance,
+during the second school study (\textit{Florimont}), the robot withstood
+interactions which lasted for a total of 160 minutes.  During this time the
+robot wrote 335 letters, 152 of which in response to demonstrations received
+from the 21 children. Technical intervention was only required for the three
+instances that the robot fell later in that day.  Otherwise, the technical
+components of the system operated autonomously and as expected over the
+sessions.
+
+Due to the modular software architecture (about 10 ROS nodes), crashes occurring
+during the sessions were usually quickly resolved by re-launching the faulty
+node alone, and did not significantly impact the interaction.
+
+The otherwise technical limitations were related to some letters or writing
+style (most notably, the ones requiring multiple strokes per letter) not being
+adequately processed by the learning algorithm. Support for such letters has
+been since added.
+
+\paragraph{Acceptance}
+
+For the approach to be effective, it is critical that children recognize and
+accept that the robot is writing by itself. When asked, no child indicated that
+they did not believe that the robot was writing by itself. There were, at times,
+questions about the robot's writing method at the beginning of the interaction,
+but when advised that the robot ``tells the tablet what it wants to write,''
+this was accepted by the children.  Besides, teachers interviewed for their
+feedback on the system advised that children are asked to draw letters in the
+air in a similar manner as part of their handwriting education. The behaviour is
+hence not unfamiliar to children.
+
+\paragraph{Engagement}
+
+As indicated in the introduction, literature~\cite{Hoy2011} suggests that a
+total of 20 handwriting practice sessions is found to be the minimum to demonstrate effective
+results in handwriting remediation. This highlights the necessity to sustain
+engagement over the long-term if we want to achieve measurable learning gains.
+
+Factually, the children engaged into the teaching activity: in the
+\textit{Florimont} study, for instance, they demonstrated an average of
+10.9 demonstration letters (SD = 4.4) for an average session duration of 11 minutes.
+In 9 out of the 14 sessions (64\%), the robot
+received demonstration letters even \emph{after} reaching the final stage of the
+interaction, suggesting an intrinsic motivation to further engage in the
+interaction.
+
+\begin{figure}[ht!]
     \centering
     \includegraphics[width=0.9\linewidth]{engagement}
-    \caption{Engagement level of the child, manually annotated by two judges
-        over the 13 min long interaction (Cronbach’s alpha test: $\alpha = .82$,
-        good inter-judge agreement). The vertical lines represent activity
+    \caption{Engagement level of one child, manually annotated by two judges
+        over a 13 min long interaction (Cronbach’s alpha test: $\alpha = .82$,
+        good inter-judge agreement). 0 means ``complete disengagement from the
+        task'', 6 means ``full engagement''. The vertical lines represent activity
     switches between handwriting teaching and story-telling (used as a
     distruber).}
-    \label{}
+    \label{engagement_level}
 \end{figure}
+
+We also conducted a qualitative assessment of the engagement levels during the
+interaction sessions (studies \textit{Haut-Lac} and \textit{ISG 2}).
+Figure~\ref{engagement_level} shows the result for one child, with a moderate
+level of engagement. On-going research in our group aims at assessing such
+engagement levels in real-time, and adapt accordingly its behaviour.
+
+Other qualitative assessment of engagement over longer interactions were
+conducted during the case studies, as reported in the following section.
 

### 2015-07-09 main.tex
A bit of cleaning

@@
 
-\subsection{Teaching beyond ICT}
-
-This research departs from the traditional approaches to educative robotics
-\cite{barker2007robotics,karim2015review}.
-
-Also notable, the robot is not used in the usual context of robotics or computer
-education, but instead in an activity -- handwriting -- which requires fine
-physical skills. In such activities, the embodied nature of the robot is
-appropriate as in interventions where motor mimicry is elicited
-\cite{Berninger1997} the arm motion for instance is, \emph{by itself}, part of
-the teaching. Furthermore, when facing a child with school difficulties, robots
-can play the role of a na\"ive learner which neither adults nor peers -- because
-of the social effects it would induce -- can convincingly play. Along these
-lines, we hope to see more research on non-STEM educational applications of
-robotics.
-
-Beyond handwriting, we believe that this work provides a novel perspective on
-the role for robots in the field of education. \emph{Learning by teaching} is a
-powerful paradigm because of not only its pedagogical efficacy, but its
-potential to positively impact the child's motivation and self-esteem. We hope that 
-this article shows that this is a very relevant context of use for robots. Indeed,
-when facing a child with school difficulties, robots can play the role of a na\"ive 
-learner which neither adults nor peers -- because of the social effects it would 
-induce -- can convincingly play.
-
 \subsection{Agency and Commitment}
@@
 several modules (from performing gestures and acquiring the user's input to
-the high-level state machine), spread over several machines (the robot itself,
+the high-level state machine), spread over several devices (the robot itself,
 one laptop and up to four tactile tablets for certain studies we conducted). We
@@
 
-Finally, the interaction itself, involving a robot, one or two children and, in
-one study, an occupational therapist, also raised by itself several challenges.
-
-We detail each of these points in the following sections.
+We detail each of these in the following sections.
 
@@
 reason in the eigenspace of the letters, instead of the cartesian space. The
-eigen space of each letters is spanned by the first $n$ eigenvectors (in our
+eigenspace of each letters is spanned by the first $n$ eigenvectors (in our
 experiments, $3 < n < 6$) of the covariance matrix observed in a standard
@@
 
-%
-%\begin{figure}[ht]
-%\centering
-%
-%\resizebox{0.9\linewidth}{!}{%
-%
-%\begin{tikzpicture}[
-%    >=latex,
-%    node distance=2cm,
-%    every edge/.style={draw, very thick},
-%    redarrow/.style={fill=Red, draw=Red},
-%    greenarrow/.style={fill=GreenYellow, draw=GreenYellow},
-%    yellowarrow/.style={fill=BurntOrange, draw=BurntOrange},
-%    cmpt/.style={draw, align=center, rounded corners, inner sep=5pt, font=\sf, fill=black!20},
-%    label/.style={midway, align=left, font=\scriptsize\sf, fill=white, above,opacity=0,text opacity=1}]
-%
-%    \node at (0,0)[cmpt, fill=Cyan!50, text width=2.7cm] (tactile) {tactile interaction \\ manager};
-%    \node[cmpt, fill=Cyan!50, left=of tactile] (display) {display \\ manager};
-%  
-%    \node [cmpt, fill=Green!50, below=of tactile, text width=2.7cm] (shape) {shape position \\ manager};
-%    \node [cmpt, fill=Green!50, left=of shape] (learner) {letter \\ learner};
-%    \node [cmpt, fill=Green!50, left=of learner] (card) {card \\ detector};
-%
-%    \node [cmpt, fill=YellowOrange!50, below=1.5cm of shape] (writer) {nao writer};
-%    \node [cmpt, fill=YellowOrange!50, left=of writer] (tablet) {tablet detector};
-%    \node [cmpt, fill=YellowOrange!50, left=of tablet] (camera) {nao camera};
-%    \node [cmpt, fill=YellowOrange!50, below=0.6cm of tablet] (pose) {robot state publisher};
-%
-%    %%% Relations between components
-%    \path (tactile) edge [->] node[label] {clear \\ screen} (display);
-%
-%    \path ($(shape.north west)!0.33!(shape.north east)$) edge [<-] node[label,left] {clear \\ screen} ($(tactile.south west)!0.33!(tactile.south east)$);
-%    \path ($(shape.north west)!0.66!(shape.north east)$) edge [<-, greenarrow] node[label,right] {feedback \\ + location} ($(tactile.south west)!0.66!(tactile.south east)$);
-%
-%    \path (shape) edge [->, redarrow] node[label, left, align=left] {letter \\to write \\ + time \\ + position} (display);
-%    \path (shape) edge [->,redarrow] node[label, right] {letter to write \\ + time \\ + position in\\tablet frame} (writer);
-%
-%    \path ($(shape.north west)!0.66!(shape.south west)$) edge [->, greenarrow] node[label,below] {feedback \\ + letter} ($(learner.north east)!0.66!(learner.south east)$);
-%    \path ($(shape.north west)!0.33!(shape.south west)$) edge [<-, redarrow] node[label] {letter to write \\ + time} ($(learner.north east)!0.33!(learner.south east)$);
-%
-%    \path (card) edge [->, yellowarrow] node[label] {word to learn} (learner);
-%    \path (camera) edge [->, yellowarrow] node[label, right] {camera image} (card);
-%    \path (camera) edge [->, redarrow] node[label] {camera image} (tablet);
-%    \path (tablet) edge [->, redarrow] node[label] {tablet pose} (writer);
-%    \path (pose) edge [->, redarrow] node[label, right] {head pose} (tablet);
-%
-%    \path (-8, 0) edge [->, redarrow] node[label] {Writing letters} (-7, 0);
-%    \path (-8, -0.6) edge [->, greenarrow] node[label] {Responding to feedback} (-7, -0.6);
-%    \path (-8, -1.2) edge [->, yellowarrow] node[label] {Detecting card input} (-7, -1.2);
-%    
-%\end{tikzpicture}
-%}
-%
-%\caption{\small Overview of the system. Components in the top row run on the tablet, 
-%those in the middle row on the central controller, and those in the bottom row 
-%on the robot.}
-%
-%    \label{fig:archi}
-%\end{figure}
-%
-
 \subsubsection*{Trajectory Following Movements}
@@
 \end{itemize}
-%
-%\begin{figure}[thpb]
-%     \begin{center}
-%            \includegraphics[width=0.7\linewidth]{naoWriting2.png}
-%    \end{center}
-%    \caption{\small A demonstration of the robot simulating the writing of a word with
-%    its finger. The motion of the robot is synchronised with the display of the
-%    tablet, communicating over ROS.}%
-%
-%   \label{fig:naoWriting}
 
@@
 working space for writing on the entire tablet.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Interaction Design}
 
@@
     \caption{\small Two metrics to assess the handwriting progresses: Euclidian
-    distance in the Eigenspace of the number dataset (top figure) or in
+    distance in the eigenspace of the number dataset (top figure) or in
 cartesian space (bottom figure). Green lines represent the robot performance,
@@
 
-No surprinsingly, the role of the robot \textit{vis-a-vis} the teacher has also
-been questionned (including an angry email from one teacher: ``You want to
+No surprisingly, the role of the robot \textit{vis-a-vis} the teacher has also
+been questioned (including an angry email from one teacher: ``You want to
 replace us!''): as we see it, the role of the robot within the classroom (or at
@@
 
-We initially envisionned the CoWriter system to be run in the back of a
+We initially envisioned the CoWriter system to be run in the back of a
 classroom with one child: this would have allowed an individual, face-to-face
@@
 robots are not going to replace them any time soon.
+
+Beyond handwriting, we believe that this work provides a novel perspective on
+the role for robots in the field of education. \emph{Learning by teaching} is a
+powerful paradigm because of not only its pedagogical efficacy, but its
+potential to positively impact the child's motivation and self-esteem. We hope that 
+this article shows that this is a very relevant context of use for robots. Indeed,
+when facing a child with school difficulties, robots can play the role of a na\"ive 
+learner which neither adults nor peers -- because of the social effects it would 
+induce -- can convincingly play.
 

### 2015-07-09 main.tex
Minor layout

@@
+%\documentclass[twocolumn]{article}
 \documentclass{article}
@@
     \subfigure[Nine allographs of the cursive ``h'', next to the reference]{
-        \includegraphics[height=4.1cm]{h}
+        \includegraphics[width=0.4\columnwidth]{h}
     }
@@
         3 first eigenvectors: clusters arise, that actually match writing styles.]{
-        \includegraphics[height=4.5cm]{eigenspace-uniformization}
+        \includegraphics[width=0.4\columnwidth]{eigenspace-uniformization}
     }
@@
 
-To date, the presented system has been deployed and tested in three different
-schools (totalling child-robot interactions with more than 70 children, aged 5
-to 8) and we also conducted 2 case studies over several weeks.
-Table~\ref{studies} gives an overview of these studies.
-
 \begin{table}[ht!]
@@
 \label{studies}
-\begin{tabular}{@{}lp{4.5cm}p{2.2cm}p{2.5cm}l@{}}
+\footnotesize
+\begin{tabular}{@{}lp{4cm}p{2.2cm}p{1.8cm}l@{}}
 \toprule
@@
 \end{table}
+
+To date, the presented system has been deployed and tested in three different
+schools (totalling child-robot interactions with more than 70 children, aged 5
+to 8) and we also conducted 2 case studies over several weeks.
+Table~\ref{studies} gives an overview of these studies.
 
@@
 \subsubsection{Context, Study Design}
-It was the first time we where trying a long-term interaction between a child
-and the CoWriter robot. So the first aim of this study was to see if such an
-interaction was simply possible. If we could create an environment to keep a child engaged
-during one hour just in writing words with the robot.
-The second question was focused on the content of the interaction. Our goal was
-to figure out what extent the child would actually improve the robot's
-writing.\\
+The CoWriter activity needs a child engaged as interaction leader. In this
+study we consider the problem of long-term interactions : is it possible to sustain this
+engagement over several one-hour sessions ?\\
 
-Diego is a six years old child. 
-Before the experiment, we could create a dataset of letters based on his main
-mistakes, but amplified by PCA. Our idea was to make the robot
-exaggerating the mistakes of Diego. Thus, by correcting the robot, 
-Diego was actually going to correct himself.\\
+Diego is a six years old child with troubles in handwriting and a poor
+self-esteem. Our goal was to figure out an environment in
+order to make Diego sustaining engagement over four sessions of one hour, one
+session per week. We decided to introduce an appealing scenario that justifyed to the child
+the activity where a robot wants to learn handwriting. We used two Nao robots :
+a blue one (called Mimi) and an orange one (called Clem). Mimi was away for a
+mission,and the two robots had to communicate by mails. But they decided to do
+it ``like humans", with handwritten messages. While Mimi was a good in
+handwriting, Clem had strong difficulties and needed the help of Diego.\\
 
-The experiment took one month. It was divided in four sessions of one hour, one session per week.
-In order to justify to the child an activity where a robot wants to learn
-handwriting, we decided to adduce a scenario. There where two Nao robots: a
-blue one (Mimi) and an orange one (Clem). They where introduced to the child as
-two hold friends. Mimi was left, and the two robots were corresponding by
-handwritten mails. But Clem was beginning in writing and it needed the help of
-Diego to write its responses to Mimi.\\
+During the first session, Diego just teached single letters to the robot.
+To chose a letter, he had to pick it in a set of magneted plastic letters, and to
+show it to the robot. But the robot was not programmed to detect and recognize
+it. A third person was entering the letter and was sending it to the
+system with the computer. With following sessions we gradually increased the
+difficulty : in the second session the child teached full words. To chose the
+word, he had to assemble plastic letters to build it and to show it to the
+robot. In the two last sessions, he teached long sentences. This time he just
+had to say the sentence to the robot, and then the third person was using the
+compurter to enter it. \\
 
-During the first session, Diego was just teaching single letters to the robot.
-Gradually we increased the difficulty : the second session he was teaching
-shirt words and in third session he was teaching long sentences. The last session was more like a test for Clem
-: it had to prove to Mimi that it actually improved its handwriting thanks to
-Diego.
+We designed a special scenario for the last session : Mimi came back to the
+laboratory. In front of Diego, Mimi said to Clem that it did not believe 
+it realy improved its handwritting, and that the responses it recieved
+was written by Diego. To prove that it actualy improved its writing,
+it gave to Clem thirty minutes for a final training. After the training, it
+asked to Clem to write some words. Finally, he admitted that Clem actually
+improved its handwriting.
 
@@
 \subsubsection{Results}
-Diego completely accepted our scenario and seriously played the game. At the
+In all sessions, Diego was focused on the activity during forty minutes up to
+one hour. In total, he provided 154 letters as demonstration.\\
+
+Diego accepted our scenario and seriously played the game. At the
 beginning he seemed not confident and intimidated.  But quickly, he gain in
-confidence and complicity with the robot. Finally, he showed a real affinity, so
-that he sent an handwritten mail to it some days after the last
-session.\\
+confidence to play with the robot. He even showed real affective bonds : he
+sent an handwritten mail to it some days after the last session.\\
 
-In all sessions, Diego was fully focused on the activity during forty minutes up to one
-hour. In total, he provided 154 letters as demonstration.\\
-
-Diego felt he improved the robot's handwriting. To make clearly visible this
-improvement, we printed the letter that the robot wrote during the third
-session before and after the help of Diego~\ref{fig:stimuli}.
+As a technical result, we do not know if Diego improved his handwriting but he
+likely conviced himself he was good enough at writing to help someone else (the
+robot in that case). To show to Diego how much he improved the robot's
+handwriting, we printed the letter that the robot wrote during the third session
+before and after his help (Fig.~\ref{fig:stimuli}).
 
@@
 This experiment took place in the therapist's office. It was divided in four
-sessions as well. This time, we assumed that a complex scenario like the one for Diego was
+sessions as well. This time, we assumed that a scenario like the one we used for Diego was
 no longer relevant with Henry. We just introduced the robot and quickly
@@
 Hence we decided to turn the CoWriter activity to teach numbers to the robot. Build on
-existing iPad application actually used by occupational therapists (Dawson Toth's \emph{ABC's
+existing iPad applications actually used by occupational therapists (Dawson Toth's \emph{ABC's
 Writer}), we designed an Android application for pre-test and post-test. It consisted in drawing
@@
 Therefor, Henry quickly enter in an self-rewarding game where the rewards were
-the improvement of the robot. In total, over the three last sessions he provided 99
+the improvement of the robot. Over the three last sessions he provided 99
 demonstrations of number shapes to the robot.

### 2015-07-10 main.tex
diego re-writing
Proof-reading first half

@@
 
-    Robots for education are not limited to support ICT teaching, and robots are
-    finding new ways to enter and blend in the classroom. This article reports
-    on such a new paradigm for educative robots, that involves \emph{learning by
-    teaching} and strong \emph{social engagement} to help children struggling
-    with handwriting. Our system relies on machine-learning and child-robot
+    Robots for education are not limited to support ICT teaching, and they are
+    indeed finding new roles in the classroom. This article reports on such a
+    new paradigm for educative robots, that involves \emph{learning by teaching}
+    and strong \emph{social engagement} to help children struggling with
+    handwriting. Our system relies on machine-learning and child-robot
     interaction with a small humanoid robot, and we present several real-world
@@
 beyond this impaired skill, psycho-social difficulties also arise: they
-underperform at school, Henry has to go for follow-up visits every week,...
-they live under the label ``requires special care''. This is a source of
-anxiety, for the children, and for their parents alike.
+underperform at school, Henry has to go for follow-up visits every week,
+they both live under the label ``requires special care''. This is a source of
+anxiety for the children and for their parents alike.
 
@@
 We present in this article a new take on this educative challenge, as a
-remediation procedure that involve a \emph{bad writer} robot that is taught by the
-child: by building on the \emph{learning by teaching} paradigm, not only the
-child practises hand-writing but, as (s)he takes on the role of the teacher,
+remediation procedure that involve a ``bad writer'' robot that is \emph{taught}
+by the child: by building on the \emph{learning by teaching} paradigm, not only
+the child practises hand-writing but, as (s)he takes on the role of the teacher,
 (s)he also positively reinforce his/her self-esteem and motivation: his/her
-social role shifts from the ``underperformer'' to ``the one who knows and teaches''.
+social role shifts from the ``underperformer'' to ``the one who knows and
+teaches''.
 
@@
 unexplored. One reason for this may be due to the requirement of an
-appropriately unskilled peer for the child to tutor: it may prove difficult if
-the child is the lowest performer in his/her class.  In some cases, it may be
-appropriate for a peer or teacher to simulate a na\"ive learner for the target
+appropriately unskilled peer for the child to tutor: this may indeed prove
+difficult if the child is the lowest performer in the class.  In some cases, it
+may be appropriate for a peer or teacher to simulate a na\"ive learner for the
 child to teach. For handwriting however, where one's skill level is visually
@@
 
-The work presented here investigates a particular role for a robot in the
+We also investigate here a particular role for a robot in the
 education of handwriting: not only is the robot actively performing the activity
@@
 
-Besides that, the commitment of the child into the interaction build on the
-``protégé effect'': teachable computer-based agents have previously been used to
-encourage the ``protégé effect'', wherein students invest more effort into
+Besides the commitment of the child into the interaction build on the
+``prot\'eg\'e effect'': the teacher feels responsible for his student, commits
+to the student's success and possibly experiences student's failure as his own
+failure to teach. Teachable computer-based agents have previously been used to
+encourage this ``prot\'eg\'e effect'', wherein students invest more effort into
 learning when it is for a teachable agent than for themselves~\cite{Chase2009}.
-As we are concerned with learning of a physical skill, the learning agent
-developed is embodied in a humanoid robot which is capable of physically
-demonstrating handwriting trajectories to its child learning partner.  This is
-supported by the potential for motor mimicry to yield significant improvements
-in handwriting interventions in which letter formations are demonstrated to
-participants~\cite{Berninger1997}. Furthermore, when compared to screen-based
-agents, robotic partners have been shown in some contexts to increase users'
-compliance with tasks~\cite{Bainbridge2011}, maintain more effective long-term
-relationships~\cite{Kidd2008}, and produce greater learning gains when acting as
-tutors~\cite{Leyzberg2012}.
+We rely on this cognitive mechanism to reinforce the child's commitment into the
+robot-mediated handwriting activity.
 
@@
 In our case, the role of agency is stronger: it induces meta-cognition (``I am
-interacting with an agent, so I need to reflect on how to best teach him''),
-which is beneficial for the learning process; it also induces a protégé effect
-(``I want my robot-agent to succeed!'') which support the commitment of child
-into the interaction, also for longer period of time (we report in this article
-upon to case studies where children were interacting with the same robot for several
-hours, spanned over several weeks).
+interacting with an agent, so I need to reflect on how to best teach him'')
+which is beneficial for the learning process; it also induces a prot\'eg\'e
+effect (``I want my robot-agent to succeed!'') which support the commitment of
+child into the interaction, also for longer period of time (we report in this
+article upon two case studies where children were indeed interacting with the
+same robot for several hours, spanned over several weeks).
 
-\subsection*{Overview of the article}
+Building on these socio-cognitive mechanisms, our intent is therefore to design
+a robotic system that would effectively support handwriting remediation in an
+original way: by getting children to teach a robotic agent how to write, those
+children would both practise without knowing it and recover self-confidence and
+self-esteem by supporting a worst-than-themselves robotic ``student''.
 
-We have sofar frame our intents: to design a robotic system that would support
-handwriting remediation in an original way: by getting children to teach a
-robotic agent how to write, those children would both practise without knowing
-it and recover self-confidence and self-esteem by supporting a
-worst-than-themselves robotic ``student''.
-
-The following sections of the paper go into details. The next section provides a
+The following sections of the paper go into details. We first provides a
 brief overview of the robotic system and the interaction it induces. We then
 give the details of machine-learning techniques that allow the robot to learn
-from the children, followed by the implementation on the Nao robot.
+from the children, followed by the implementation on a {\sc nao} robot.
 
-Next, we present and report on the four school experiments and the two case
+We then present and report on the four school experiments and the two case
 studies that we conducted over the last two years. While the focus of the school
@@
 Figure~\ref{experimental_setup} illustrates our general experimental setup: a
-face-to-face child-robot interaction with an (autonomous) Aldebran's Nao robot.
+face-to-face child-robot interaction with an (autonomous) Aldebran's {\sc nao}
+robot.
 
@@
 something (a single letter, a number or a full word), and push the tablet
-towards the robot, the robot writes on the tablet by gesturing the writing, but
-\emph{without actually physically touching the tablet}), the child then pull
-back the tablet, correct the robot's attempt by writing him/herself on top or
-next to the robot's writing (see the picture in Figure~\ref{fig:diego}), and
-``send'' his/her demonstration to the robot by pressing a small button on the
-tablet. The robot ``learns'' from this demonstration and tries again.
+towards the robot, the robot writes on the tablet by gesturing the writing (but
+without actually physically touching the tablet), the child then pull back the
+tablet, corrects the robot's attempt by writing him/herself on top or next to
+the robot's writing (see Figure~\ref{fig:diego}), and ``send'' his/her
+demonstration to the robot by pressing a small button on the tablet. The robot
+``learns'' from this demonstration and tries again.
 
@@
 
-Implementing such a system raises several challenges: first, acquiring,
-analyzing and learning from hand-written demonstration lays at the core of the
-our approach and we developed several algorithms for the robot to generate
+Implementing such a system raises several challenges: first, the acquisition,
+analyse and learning from hand-written demonstration, which lays at the core of the
+our approach, necessitates the development of several algorithms for the robot to generate
 initial bad writing and to respond in an adequate manner, showing visible (but
@@
 
-Then, the actual implementation on the robot required the coordination of
+Then, the actual implementation on the robot requires the coordination of
 several modules (from performing gestures and acquiring the user's input to
@@
 The main insight for both the generation and the learning of letters is to
-reason in the eigenspace of the letters, instead of the cartesian space. The
-eigenspace of each letters is spanned by the first $n$ eigenvectors (in our
-experiments, $3 < n < 6$) of the covariance matrix observed in a standard
-dataset of adult letters (the UJI Pen Characters 2 dataset~\cite{Llorens2008}).
-This procedure, based on a Principle Component Analysis (PCA), is explained in
-details in a previously published article~\cite{hood2015when}.
+reason about the shape of letters in their eigenspace, instead of the natural
+cartesian space. The eigenspace of each letters is spanned by the first $n$
+eigenvectors (in our experiments, $3 < n < 6$) of the covariance matrix
+generated from a standard dataset of adult letters (the UJI Pen Characters 2
+dataset~\cite{Llorens2008}).  This procedure, based on a Principle Component
+Analysis (PCA), is explained in details in a previously published
+article~\cite{hood2015when}.
 
@@
 are actually \emph{exaggerations} of variance of writing styles that naturally
-exist amongst adult writers.  Figure~\ref{fig:sampleLetters} shows examples of
+occurs amongst adult writers.  Figure~\ref{fig:sampleLetters} shows examples of
 deformed ``g'' generated with such a technique.
@@
         PCA-based approach allows to \emph{automatically} generate letters whose
-        errors can be \emph{semantically} interpreted: A has a too large bottom
+        errors can be \emph{semantically} interpreted, \eg A has a too large bottom
         loop, B has a wide top loop, the bottom loop of C is not correctly
-        closed, the top loop of D is not closed.}
+        closed, the top loop of D is not closed, etc.}
 
@@
 
-The same technique can be applied to \emph{classify} demonstrations,
-\emph{assess} their quality and
-\emph{learn} from them. Figure~\ref{fig:h} shows 9 allographs of ``h'' written by a 6
-years old child, along with the mean letter from the
-dataset (reference letter). By projecting each of the demonstrations onto the eigenspace
-of ``h'' (Figure~\ref{fig:h}b), we observe that:
+The same technique can also be applied to \emph{classify} demonstrations,
+\emph{assess} their quality and \emph{learn} from them. Figure~\ref{fig:h} shows
+for example 9 allographs of ``h'' written by a 6 years old child, along with the
+mean letter from the dataset (reference letter). By projecting each of the
+demonstrations onto the eigenspace of ``h'' (Figure~\ref{fig:h}b), we observe
+that:
 
@@
     generated from the reference dataset effectively clusters the samples
-    according to their topological similarity. Allographs that are visually similar to
+    according to their topological similarity. Allographs that are similar to
     the reference are close to it in the eigenspace.}
@@
 a specific pace towards the child demonstration in the eigenspace.
-Figure~\ref{learning_6_demos}, at the end of the article, shows a complete learning cycle
-with the number ``6''.
+Figure~\ref{learning_6_demos}, at the end of the article, illustrates the
+process with the complete learning cycle of the number ``6''.
 
@@
 
-Our system is embodied through an Aldebaran's {\sc nao} (V4 or V5, depending on the studies)
-humanoid robot. This choice is motivated by its approachable
-design~\cite{Gouaillier2008}, its size (58cm) and inherently safe structure (lightweight
-plastic) (making it suitable for close interaction with children), its low price
-(making it closer to what school may afford in the coming years) and finally its
-ease of deployment on the field.
-
-{\sc nao} is a humanoid biped, has 25 degrees of freedom,
-two cameras, speech capabilities and the ability to autonomously execute a range
-of tasks. While it can walk, this capability is not used in this project, and
-the robot remains in a crouching posture during the whole interaction.
+Our system is embodied in an Aldebaran's {\sc nao} (V4 or V5, depending on the
+studies) humanoid robot. This choice is motivated by its approachable
+design~\cite{Gouaillier2008}, its size (58cm) and inherently safe structure
+(lightweight plastic) making it suitable for close interaction with children,
+its low price (making it closer to what school may afford in the coming years)
+and finally its ease of deployment on the field.
 
 Robotic handwriting requires precise closed-loop control of the arm and hand
-motion. Because of the limited fine motor skills possible
-with such an affordable robot, in addition to the absence of force feedback and
-other technical necessities, we have opted for \emph{simulated
-handwriting}: the robot draws letters in the air, and the actual writing is
-displayed on a synchronised tablet.
+motion. Because of the limited fine motor skills possible with such an
+affordable robot, in addition to the absence of force feedback, we have opted
+for \emph{simulated handwriting}: the robot draws letters in the air, and the
+actual writing is displayed on a synchronised tablet.
 

### 2015-07-10 main.tex
Changed children's names + blurred face

@@
 
-Henry is five and an half, and has been diagnosed with visuo-constructive
+Thomas is five and an half, and has been diagnosed with visuo-constructive
 deficits. He is under the care of an occupational therapist, and tries to
-workaround his inability to draw letters in a consistent manner. Diego is six
+workaround his inability to draw letters in a consistent manner. Vincent is six
 and struggles at school with his poor handwriting and even poorer
@@
 
-While Henry is lively and always quick at shifting his attention from one
-activity to another, Diego is shy and poised. Two very different children,
+While Thomas is lively and always quick at shifting his attention from one
+activity to another, Vincent is shy and poised. Two very different children,
 facing however the same difficulty to write in a legible manner. And, hidden
 beyond this impaired skill, psycho-social difficulties also arise: they
-underperform at school, Henry has to go for follow-up visits every week,
+underperform at school, Thomas has to go for follow-up visits every week,
 they both live under the label ``requires special care''. This is a source of
@@
     \includegraphics[width=0.9\linewidth]{henry}
-    \caption{\small Henry teaching Nao how to write numbers, with the help of an
+    \caption{\small Thomas teaching Nao how to write numbers, with the help of an
     occupational therapist.}
@@
 {\it Florimont} & Individual/Pair interaction at school         & 11 min/group               & 7 (individual) + 7 $\times$ 2 (pairs) & 7-8        \\
-{\it Diego}     & Case-study, spanning over 4 weeks              & 1.5 hours $\times$ 4 weeks & 1                                     & 5          \\
+{\it Vincent}     & Case-study, spanning over 4 weeks              & 1.5 hours $\times$ 4 weeks & 1                                     & 5          \\
 {\it Haut-Lac}  & Pair (with turn-taking) interaction at school & 26 min/group                            & 7 $\times$ 2                          & 5-6        \\
 {\it ISG 2}     & Individual interaction at school              & 18 min                            & 6                                     & 5-6        \\
-{\it Henry}     & Case-study, spanning over 4 weeks              & 1h $\times$ 4 weeks        & 1                                     & 5          \\ \bottomrule
+{\it Thomas}     & Case-study, spanning over 4 weeks              & 1h $\times$ 4 weeks        & 1                                     & 5          \\ \bottomrule
 \end{tabular}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Case Study 1: Diego}
+\subsection{Case Study 1: Vincent}
 
@@
 
-Diego is a six years old child with troubles in handwriting and a poor
+Vincent is a six years old child with troubles in handwriting and a poor
 self-esteem. Our goal was to figure out an environment in
-order to make Diego sustaining engagement over four sessions of one hour, one
+order to make Vincent sustaining engagement over four sessions of one hour, one
 session per week. We decided to introduce an appealing scenario that justifyed to the child
@@
 it ``like humans", with handwritten messages. While Mimi was a good in
-handwriting, Clem had strong difficulties and needed the help of Diego.\\
+handwriting, Clem had strong difficulties and needed the help of Vincent.\\
 
-During the first session, Diego just teached single letters to the robot.
+During the first session, Vincent just teached single letters to the robot.
 To chose a letter, he had to pick it in a set of magneted plastic letters, and to
@@
 We designed a special scenario for the last session : Mimi came back to the
-laboratory. In front of Diego, Mimi said to Clem that it did not believe 
+laboratory. In front of Vincent, Mimi said to Clem that it did not believe 
 it realy improved its handwritting, and that the responses it recieved
-was written by Diego. To prove that it actualy improved its writing,
+was written by Vincent. To prove that it actualy improved its writing,
 it gave to Clem thirty minutes for a final training. After the training, it
@@
     \includegraphics[width=0.5\linewidth]{diego}
-    \caption{\small Diego working with Nao}
+    \caption{\small Vincent working with Nao}
     \label{fig:diego}
@@
 \subsubsection{Results}
-In all sessions, Diego was focused on the activity during forty minutes up to
+In all sessions, Vincent was focused on the activity during forty minutes up to
 one hour. In total, he provided 154 letters as demonstration.\\
 
-Diego accepted our scenario and seriously played the game. At the
+Vincent accepted our scenario and seriously played the game. At the
 beginning he seemed not confident and intimidated.  But quickly, he gain in
@@
 
-As a technical result, we do not know if Diego improved his handwriting but he
+As a technical result, we do not know if Vincent improved his handwriting but he
 likely conviced himself he was good enough at writing to help someone else (the
-robot in that case). To show to Diego how much he improved the robot's
+robot in that case). To show to Vincent how much he improved the robot's
 handwriting, we printed the letter that the robot wrote during the third session
@@
     }
-    \subfigure[Final letter, after training with Diego]{
+    \subfigure[Final letter, after training with Vincent]{
         \includegraphics[height=6cm]{diego-final-letter}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Case Study 2: Henry}
+\subsection{Case Study 2: Thomas}
 
 \subsubsection{Context, Study Design}
-Henry is a five and half years old child. It is followed by an occupational
+Thomas is a five and half years old child. It is followed by an occupational
 therapist. According to the therapist, he presents visuo-constructive deficits.
 From our perspective, the main effect of this trouble during handwriting
-activity was that Henry
+activity was that Thomas
 did not have any mental template to draw the shape of a letter. So he was 
 apparently performing random attempts and then was comparing with the provided template.
-What is more, Henry is strongly careless : he rarely payed attention to
+What is more, Thomas is strongly careless : he rarely payed attention to
 advices, even to what he was doing when he was currently drawing, and he was
@@
 Hence, the goal of this study is to face those two challenging issues
-(visuo-constructive deficits and inattention) in order to keep Henry focused
+(visuo-constructive deficits and inattention) in order to keep Thomas focused
 on the activity during forty-minutes interactions, and to make the robot
@@
 This experiment took place in the therapist's office. It was divided in four
-sessions as well. This time, we assumed that a scenario like the one we used for Diego was
-no longer relevant with Henry. We just introduced the robot and quickly
+sessions as well. This time, we assumed that a scenario like the one we used for Vincent was
+no longer relevant with Thomas. We just introduced the robot and quickly
 said that it was seeking help to train for a robot handwriting contest.\\
 
-Before the experiment, Henry was working on writing numbers with the therapist.
+Before the experiment, Thomas was working on writing numbers with the therapist.
 Hence we decided to turn the CoWriter activity to teach numbers to the robot. Build on
@@
 
-Since Henry was frequently drawing horizontally inverted numbers, or even
+Since Thomas was frequently drawing horizontally inverted numbers, or even
 absolutely unrecognizable shapes, we programmed the robot to refuse shapes that
@@
 \subsubsection{Results}
-Despite his inattention, Henry was able to keep on the activity during more than
+Despite his inattention, Thomas was able to keep on the activity during more than
 forty minutes in each session.\\
 
-As soon as Henry understood that the robot was only accepting recognizable
+As soon as Thomas understood that the robot was only accepting recognizable
 shapes, he started to take care on the shapes he was providing to it. He used the
 clearing button to retry when he judged that his numbers where not going to be
-accepted by the robot. According to the therapist, it was the first time Henry
+accepted by the robot. According to the therapist, it was the first time Thomas
 was intrinsically correcting himself.\\
 
-Since the robot was learning from scratch, each time Henry succeeded to make his
+Since the robot was learning from scratch, each time Thomas succeeded to make his
 demonstrations accepted by it, the improvement of its handwriting was
 obvious~\ref{learning_6_demos},~\ref{henry_distances}.
-Therefor, Henry quickly enter in an self-rewarding game where the rewards were
+Therefor, Thomas quickly enter in an self-rewarding game where the rewards were
 the improvement of the robot. Over the three last sessions he provided 99
@@
     \includegraphics[width=0.9\linewidth]{learning_6_demos}
-    \caption{\small Henry's demonstrations for the number ``6'' (top row) and
+    \caption{\small Thomas's demonstrations for the number ``6'' (top row) and
         corresponding shapes generated by the robot. After eight demonstrations,
-        Henry decided that the robot's ``6'' was good enough, and went to
+        Thomas decided that the robot's ``6'' was good enough, and went to
     another character: in that respect, he was the one leading the learning
@@
 cartesian space (bottom figure). Green lines represent the robot performance,
-blue lines Henry's performance. The round IDs correspond to the demonstrations
+blue lines Thomas's performance. The round IDs correspond to the demonstrations
 pictured on Figure~\ref{henry_demos}.}
@@
 child to give feedback to the robot or to go to the next letter or word.  At a
-higher orchestration level (and as reported in the two case studies with Diego
-and Henry), the educational scenarios were also always designed and monitored by
+higher orchestration level (and as reported in the two case studies with Vincent
+and Thomas), the educational scenarios were also always designed and monitored by
 the adults.

### 2015-07-10 main.tex
Proof-reading second half, except case studies

@@
 
-\title{Learning by Teaching a Robot: the Case of Handwriting}
+\title{Learning by Teaching a Robot:\\The Case of Handwriting}
 
@@
 
-\resizebox{0.6\linewidth}{!}{%
+\resizebox{0.7\linewidth}{!}{%
 
@@
 
+    \node[draw,above right=2 of laptop,anchor=north west,text width=4cm] (processes)
+    {\sf\scriptsize machine-learning, \\letters/gestures
+    generation, \\interaction supervision};
+    \path (laptop) edge[dashed] (processes);
+
     \path (nao) edge [->,redarrow, bend left] node[label, auto] {robot state} (laptop);
-    \path (laptop) edge [->,greenarrow, bend left] node[label, auto] {writing gesture} (nao);
+    \path (laptop) edge [->,greenarrow, bend left] node[label, auto] {writing gestures} (nao);
 
     \path (tablet) edge [->,redarrow, bend left] node[label, auto] {demonstrations,\\turn taking} (laptop);
-    \path (laptop) edge [->,redarrow, bend left] node[label, auto=right] {letter to write} (tablet);
+    \path (laptop) edge [->,redarrow, bend left] node[label,
+    auto=right,align=right]
+    {path of\\ letters to display} (tablet);
 
-    \path (selection) edge [->,redarrow] node[label, auto] {letter/word to write} (laptop);
+    \path (selection) edge [->,redarrow] node[label, auto=right] {letter/word to write} (laptop);
 
@@
 
-\caption{\small Overview of the system. In total, the system runs about 10 ROS nodes,
+\caption{\small \textbf{Overview of the system}. In total, the system runs about 10 ROS nodes,
     distributed over the robot itself, a central laptop and Android tablets.}
@@
 to four Android tablets (the main tablet is used to print the robot's letter and
-to acquire the children's demonstrations. More tablets have been used in some
+to acquire the children's demonstrations; more tablets have been used in some
 studies, either to let the child input words to be written, or for the
@@
 Since the system does not actually require any CPU-intensive process, the laptop
-could be removed and the whole logic could run on the robot. Due to the relative
-difficulty to deploy and debug ROS nodes directly on the robot, we however
-effectively kept the laptop during our experiments.
+can be removed and the whole logic run on the robot. Due to the relative
+difficulty to deploy and debug ROS nodes directly on the robot, the laptop
+remains however convenient during the development phase and we kept if during
+our experiments.
 
@@
 \url{https://github.com/chili-epfl/cowriter_letter_learning}.}
-
-\subsubsection*{Trajectory Following Movements}
-
-Using simulated handwriting provides an opportunity for the robot's writing to
-appear smoother than would be achievable with a writing instrument. However, the
-robot's motions must still sufficiently match the displayed trajectory in order
-capture the engagement of the child participant in the action. Aldebaran's NaoQi
-API is used for the inverse kinematics of the trajectory following. ROS is used
-for integration of {\sc nao} with external reference frames, such as the
-tablet's location, using the $tf$ transformation library~\cite{Foote2013}.
-
-When using simulated handwriting, it is no longer necessary that the robot
-engages in the typical style of handwriting of using a writing instrument at a
-desk. Having instead the robot point at the writing surface to cause the
-trajectory to appear (as seen in Figure \ref{fig:schools}, left) has several advantages:
-
-\begin{itemize}
-
-    \item The working space of the robot increases, both in the technical sense
-        and the interaction sense: someone can, in theory, show the tablet to
-        the robot from across the room and have it still respond, without
-        needing the tablet to be within arm's reach,
-
-    \item Concerns about whether or not the child would start mimicking the
-        robot's incorrect writing form (\eg pen grip) are mitigated,
-
-    \item Perhaps most significantly, the accuracy of the matching of the
-        robot's motion with the trajectory displayed on the tablet is not as
-        critical. This is because a pen tip would be expected to touch the
-        tablet exactly at the trajectory point, while a fingertip may not.
-
-\end{itemize}
-
-We have therefore designed the system in such a way that the robot is simulating
-handwriting by pointing at the tablet\footnote{Teachers interviewed for their
-feedback on the system advised that children are asked to draw letters in
-the air in a similar manner as part of their handwriting education. The
-behaviour is hence not unfamiliar to children.}. As interacting with a
-tablet with one's finger is not uncommon, this may aid the acceptance of the
-writing style by users.
-
-Because motion planning is performed with respect to the hand of the robot,
-rather than its fingertip, one or two of the orientation degrees of freedom of
-the hand are fixed to keep the finger approximately perpendicular to the writing
-surface, depending on the desired accuracy. The remaining free orientation(s),
-coupled with the whole-body motion control available, allow for a sufficient
-working space for writing on the entire tablet.
 
@@
 
+To date, the system has been deployed and tested in three different
+schools (totalling child-robot interactions with more than 70 children, aged 5
+to 8) and we also conducted 2 case studies that lasted several weeks.
+Table~\ref{studies} gives an overview of these studies.
+
 \begin{table}[ht!]
@@
 \footnotesize
-\begin{tabular}{@{}lp{4cm}p{2.2cm}p{1.8cm}l@{}}
+\begin{tabular}{@{}lp{4cm}p{2.2cm}p{2cm}l@{}}
 \toprule
-{\bf Study}     & {\bf Type}                                    & {\bf Duration}              & {\bf Children \#}                     & {\bf Ages} \\ \midrule
+{\bf Study}     & {\bf Type}                                    & {\bf Avg. duration}              & {\bf Children \#}                     & {\bf Ages} \\ \midrule
 {\it ISG 1}     & Unstructured group interaction at school      & 16 min/group               & 4 $\times$ 8 children                 & 6-7        \\
 {\it Florimont} & Individual/Pair interaction at school         & 11 min/group               & 7 (individual) + 7 $\times$ 2 (pairs) & 7-8        \\
-{\it Vincent}     & Case-study, spanning over 4 weeks              & 1.5 hours $\times$ 4 weeks & 1                                     & 5          \\
-{\it Haut-Lac}  & Pair (with turn-taking) interaction at school & 26 min/group                            & 7 $\times$ 2                          & 5-6        \\
+{\it Vincent}     & Case-study, spanning over 4 weeks              & 1.5h $\times$ 4 weeks & 1                                     & 6          \\
+{\it Haut-Lac}  & Pair interaction at school & 26 min/group                            & 7 $\times$ 2                          & 5-6        \\
 {\it ISG 2}     & Individual interaction at school              & 18 min                            & 6                                     & 5-6        \\
@@
 \end{table}
-
-To date, the presented system has been deployed and tested in three different
-schools (totalling child-robot interactions with more than 70 children, aged 5
-to 8) and we also conducted 2 case studies over several weeks.
-Table~\ref{studies} gives an overview of these studies.
 
@@
 the system (is it actually able to autonomously write and learn from
-demonstrations?) and the interaction (is the apparatus easy to grasp and to
-interact with for children), and to study the initial acceptance of the robot in
-the school environment (we conducted several formal and informal discussions
-with teachers).
-
-We also relied on these four studies to acquire a dataset of children's
-handwritten letters and to study how they engage with the robot (and maintain or
-not this engagement).
+demonstrations?) and test the interaction (is the apparatus easy to grasp and to
+interact with for children?). We also studied the initial acceptance of the
+robot in the school environment (through several formal and informal discussions
+with teachers) and how children engage with the robot (and maintain or not this
+engagement).
 
@@
 
-Critically, these studies were conducted with \emph{whole classes}: we did not
-select underperforming children that requires careful beforehand organization
-with the school to limit ethical issues. Having more children (73 in total) was
-however beneficial for these preliminary studies.
+Critically, these studies were conducted with \emph{whole classes}: we decided
+not to select specifically underperforming children as having more children (73
+in total) was beneficial for these preliminary studies, and this would have
+besides required careful organization with the school due to ethical concerns.
 
@@
 from the 21 children. Technical intervention was only required for the three
-instances that the robot fell later in that day.  Otherwise, the technical
+instances that the robot fell during that day.  Otherwise, the technical
 components of the system operated autonomously and as expected over the
@@
 
-Due to the modular software architecture (about 10 ROS nodes), crashes occurring
-during the sessions were usually quickly resolved by re-launching the faulty
-node alone, and did not significantly impact the interaction.
+Due to the modular software architecture (about 10 ROS nodes), the occasional
+crashes occurring during others sessions were usually quickly resolved by
+re-launching the faulty node alone, and did not significantly impact the
+interaction.
 
@@
 adequately processed by the learning algorithm. Support for such letters has
-been since added.
+been added as a follow-up of the studies.
 
@@
 
-For the approach to be effective, it is critical that children recognize and
-accept that the robot is writing by itself. When asked, no child indicated that
-they did not believe that the robot was writing by itself. There were, at times,
-questions about the robot's writing method at the beginning of the interaction,
-but when advised that the robot ``tells the tablet what it wants to write,''
-this was accepted by the children.  Besides, teachers interviewed for their
-feedback on the system advised that children are asked to draw letters in the
-air in a similar manner as part of their handwriting education. The behaviour is
-hence not unfamiliar to children.
+Children's recognition that the robot is writing by itself is critical for our
+approach to be effective. When asked, no child indicated that they did not
+believe that the robot was writing by itself. There were, at times, questions
+about the robot's writing method at the beginning of the interaction, but when
+advised that the robot ``tells the tablet what it wants to write,'' this was
+accepted by the children.  Besides, teachers interviewed for their feedback on
+the system advised that children are asked to draw letters in the air in a
+similar manner as part of their handwriting education. The behaviour is hence
+not unfamiliar to children.
 
-\paragraph{Engagement}
+More generally, the approach was well accepted and recognized as useful and
+promising by teachers and parents; this was \textit{a posteriori} confirmed by
+the numerous spontaneous contacts made by parents and therapists who were
+looking forward to use the system with their children.
 
-As indicated in the introduction, literature~\cite{Hoy2011} suggests that a
-total of 20 handwriting practice sessions is found to be the minimum to demonstrate effective
-results in handwriting remediation. This highlights the necessity to sustain
-engagement over the long-term if we want to achieve measurable learning gains.
+\paragraph{Sustained engagement}
+
+As indicated in the introduction, the literature suggests that 20 handwriting
+practice sessions is found to be the minimum to demonstrate effective results in
+handwriting remediation~\cite{Hoy2011}. This highlights the necessity to sustain
+a child-robot engagement over the long-term if we want to achieve measurable learning gains.
 
@@
     \includegraphics[width=0.9\linewidth]{engagement}
-    \caption{Engagement level of one child, manually annotated by two judges
+    \caption{\small Engagement level of one child, manually annotated by two judges
         over a 13 min long interaction (Cronbach’s alpha test: $\alpha = .82$,
@@
 
-We also conducted a qualitative assessment of the engagement levels during the
-interaction sessions (studies \textit{Haut-Lac} and \textit{ISG 2}).
-Figure~\ref{engagement_level} shows the result for one child, with a moderate
-level of engagement. On-going research in our group aims at assessing such
-engagement levels in real-time, and adapt accordingly its behaviour.
+We also conducted a qualitative assessment of the engagement levels during
+several interaction sessions (studies \textit{Haut-Lac} and \textit{ISG 2}).
+While the overall engagement was high, Figure~\ref{engagement_level} shows an
+example of a child with moderate level of engagement. On-going research in our
+group aims at assessing such engagement levels in real-time, so that the robot
+can address possible disengagement (by changing/adapting its physical behaviour,
+by switching to different activities, etc.)
 
-Other qualitative assessment of engagement over longer interactions were
+Further qualitative assessment of engagement over longer interactions were
 conducted during the case studies, as reported in the following section.
@@
     \caption{\small Text generated by the robot, before and after training with
-    the child. As an example, the red box emphasizes the changes on the word
+        the child (in French). As an example, the red box emphasizes the changes on the word
     ``envoyer''.}
@@
 
-The CoWriter project, as presented above, gives a novel perspective on educative
-robots, at several levels:
+We believe that this research provides a novel perspective on educative
+robots at several levels. We show that:
 
@@
         hours) around a fundamentally routine -- yet challenging -- educational
-        task (hand-writing learning).
+        task (hand-writing learning),
+
+    \item besides, the social commitment induces a \textbf{cognitive engagement}
+        of the child with the robot which is \textbf{a key learning lever} as
+        its elicits \textbf{reflective, meta-cognitive mechanisms on the
+        learning task}.
 \end{itemize}
 
-No surprisingly, the role of the robot \textit{vis-a-vis} the teacher has also
-been questioned (including an angry email from one teacher: ``You want to
-replace us!''): as we see it, the role of the robot within the classroom (or at
-the therapist's surgery) does not infringe upon the role of the adult (teacher
-or therapist).  The core of the \emph{learning by teaching} paradigm relies on
-the child becoming the teacher of an underperforming pupil (the robot): from
-that perspective, the robot does not replace the teacher, on the contrary. It
-plays a different role in the classroom, which happen to be novel as well: the
-robot is the \emph{least} performing student, and however a very patient, always
-eager to improve, one.
+The place and role of the robot \textit{vis-\`a-vis} the teacher can be
+questioned: as we see it, the role of the robot within the classroom (or at the
+therapist's surgery) does not infringe upon the role of the adult (teacher or
+therapist).  The core of the \emph{learning by teaching} paradigm relies on the
+child becoming the teacher of an underperforming pupil (the robot): from that
+perspective, the robot does not replace the teacher, on the contrary. It plays a
+different role in the classroom, which happen to be novel as well: the robot is
+the \emph{least} performing student, and still a very patient, always eager to
+improve, one.
 
@@
 experiments we conducted, the adult always played a key role at prompting the
-child to give feedback to the robot or to go to the next letter or word.  At a
+child to give feedback to the robot or to move to the next letter or word.  At a
 higher orchestration level (and as reported in the two case studies with Vincent
-and Thomas), the educational scenarios were also always designed and monitored by
-the adults.
+and Thomas), the educational scenarios were also always designed and monitored
+by the adults.
 
-We initially envisioned the CoWriter system to be run in the back of a
-classroom with one child: this would have allowed an individual, face-to-face
-remediation approach, not otherwise tractable for a teacher with 20 pupils.
-This is unlikely to happen. In our experience, the teacher keeps such an
-important role that the interaction and the learning would hardly occur if the
-child is left alone (or even semi-alone). At the end, we can reassure teachers:
-robots are not going to replace them any time soon.
+We initially envisioned our system to be run in the back of a classroom with one
+child: this would have allowed an individual, face-to-face remediation approach,
+not otherwise tractable for a teacher with 20 pupils.  This is however unlikely
+to happen soon. In our experience, the teacher keeps such an important role that
+the interaction (and the learning!) would hardly occur if the child is left
+alone (or even semi-alone). We can therefore reassure teachers: robots are not
+going to replace them any time soon.
 
-Beyond handwriting, we believe that this work provides a novel perspective on
-the role for robots in the field of education. \emph{Learning by teaching} is a
-powerful paradigm because of not only its pedagogical efficacy, but its
-potential to positively impact the child's motivation and self-esteem. We hope that 
-this article shows that this is a very relevant context of use for robots. Indeed,
-when facing a child with school difficulties, robots can play the role of a na\"ive 
-learner which neither adults nor peers -- because of the social effects it would 
-induce -- can convincingly play.
+Beyond handwriting, we however do believe that this work provides a novel
+perspective on the role for robots in the field of education. \emph{Learning by
+teaching} is a powerful paradigm because of not only its pedagogical efficacy,
+but its potential to positively impact the child's motivation and self-esteem.
+We hope that this article shows that this is a very relevant context of use for
+robots: when facing a child with school difficulties, robots can play the role
+of a na\"ive learner which neither adults nor peers -- because of the social
+effects it would induce -- can convincingly play.
 

### 2015-07-10 main.tex
re-writing henry part

@@
 \subsubsection{Context, Study Design}
-Thomas is a five and half years old child. It is followed by an occupational
-therapist. According to the therapist, he presents visuo-constructive deficits.
-From our perspective, the main effect of this trouble during handwriting
-activity was that Thomas
-did not have any mental template to draw the shape of a letter. So he was 
-apparently performing random attempts and then was comparing with the provided template.
-What is more, Thomas is strongly careless : he rarely payed attention to
+
+This study focused technical adaptations of the CoWriter activity for a
+child diagnosed with real writing deficits.\\
+
+Thomas is a six years old child. It is under the care of an occupational
+therapist. According to the therapist, he shows visuo-constructive deficits.
+As an effect in writing activities, he was frequently performing random attempts and then was comparing
+with the provided template. What is more, Thomas is strongly careless : he rarely payed attention to
 advices, even to what he was doing when he was currently drawing, and he was
 quickly shifting his attention from one
-activity to another.
-Hence, the goal of this study is to face those two challenging issues
-(visuo-constructive deficits and inattention) in order to keep Thomas focused
-on the activity during forty-minutes interactions, and to make the robot
+activity to another.\\
+
+Our objective is to investigate small modifications of the activity adapted to
+the troubles of Thomas (visuo-constructive deficits and inattention) in order to
+sustain him focused
+on the activity during forty-minutes session, and to make the robot
 evidently learning from his demonstrations.\\
 
-This experiment took place in the therapist's office. It was divided in four
-sessions as well. This time, we assumed that a scenario like the one we used for Vincent was
+The experiment took place in the therapist's office. It was divided in four
+sessions. This time, we assumed that a scenario like the one we used for Vincent was
 no longer relevant with Thomas. We just introduced the robot and quickly
@@
 
-Since Thomas was frequently drawing horizontally inverted numbers, or even
-absolutely unrecognizable shapes, we programmed the robot to refuse shapes that
-was too distant to a reference with a threshold we arbitrary fixed. In that way,
+Since Thomas was frequently drawing horizontally-inverted numbers, or even
+unrecognizable allographs, the learning algorithm of the robot was converging to
+meaningless scrawls. To fix this problem, we programmed the robot to refuse allographs that
+were too distant to a reference with a threshold we arbitrary fixed. In that way,
 the child was forced to take care on what he was providing to the robot as
 demonstration. Also, to make sure the robot was going to improve its handwriting
-and to clearly show this improvement, we decided to make it absolutely starting
+and to clearly show this improvement, we decided to make it starting
 from scratch : for all numbers, the first try of the robot resulted in
-a simple vertical stroke (see the first robot's try in~\ref{learning_6_demos}).
+a simple vertical stroke (see the first robot's try
+in~\ref{learning_6_demos}).\\
+
+In this setup, we added a second tablet with one button per numbers. It was used
+by the child to chose a new number to teach to the robot. It also provided the
+possiblity to enter letters or words, and to switch to another activity (the
+robot telling a story).
 
@@
 \subsubsection{Results}
+
 Despite his inattention, Thomas was able to keep on the activity during more than
-forty minutes in each session.\\
+forty minutes in each session. In total, 55 allographs out of 82 
+provided by the child as demonstration were acceptable by the robot (with a
+progressive improvement from 13 out of 28 in the first session up to 26 out
+of 29 in the last session).\\
 
 As soon as Thomas understood that the robot was only accepting recognizable
-shapes, he started to take care on the shapes he was providing to it. He used the
+allographs, he started to take care on the allographs he was providing to it. He used the
 clearing button to retry when he judged that his numbers where not going to be
@@
 demonstrations accepted by it, the improvement of its handwriting was
-obvious~\ref{learning_6_demos},~\ref{henry_distances}.
-Therefor, Thomas quickly enter in an self-rewarding game where the rewards were
-the improvement of the robot. Over the three last sessions he provided 99
-demonstrations of number shapes to the robot.
+visible~\ref{learning_6_demos},~\ref{henry_distances}.
+Therefor, Thomas quickly entered in an self-rewarding game where the rewards were
+the improvements of the robot, that enhanced his attention.
 

### 2015-07-10 main.tex
Reworked Diego's study part

@@
 by the child: by building on the \emph{learning by teaching} paradigm, not only
-the child practises hand-writing but, as (s)he takes on the role of the teacher,
+the child practises handwriting but, as (s)he takes on the role of the teacher,
 (s)he also positively reinforce his/her self-esteem and motivation: his/her
@@
 
-Further qualitative assessment of engagement over longer interactions were
-conducted during the case studies, as reported in the following section.
+We also performed further qualitative assessment of engagement over longer
+interaction periods: two 4-weeks long case studies with one child at a time have
+been conducted so far, the first one (\emph{Vincent}) focusing on
+\emph{engagement on the long run}, the second one (\emph{Thomas}) looking at
+\emph{acceptance and impact in a real therapeutic context}.
 
@@
 \subsubsection{Context, Study Design}
-The CoWriter activity needs a child engaged as interaction leader. In this
-study we consider the problem of long-term interactions : is it possible to sustain this
-engagement over several one-hour sessions ?\\
 
-Vincent is a six years old child with troubles in handwriting and a poor
-self-esteem. Our goal was to figure out an environment in
-order to make Vincent sustaining engagement over four sessions of one hour, one
-session per week. We decided to introduce an appealing scenario that justifyed to the child
-the activity where a robot wants to learn handwriting. We used two Nao robots :
-a blue one (called Mimi) and an orange one (called Clem). Mimi was away for a
-mission,and the two robots had to communicate by mails. But they decided to do
-it ``like humans", with handwritten messages. While Mimi was a good in
-handwriting, Clem had strong difficulties and needed the help of Vincent.\\
-
-During the first session, Vincent just teached single letters to the robot.
-To chose a letter, he had to pick it in a set of magneted plastic letters, and to
-show it to the robot. But the robot was not programmed to detect and recognize
-it. A third person was entering the letter and was sending it to the
-system with the computer. With following sessions we gradually increased the
-difficulty : in the second session the child teached full words. To chose the
-word, he had to assemble plastic letters to build it and to show it to the
-robot. In the two last sessions, he teached long sentences. This time he just
-had to say the sentence to the robot, and then the third person was using the
-compurter to enter it. \\
-
-We designed a special scenario for the last session : Mimi came back to the
-laboratory. In front of Vincent, Mimi said to Clem that it did not believe 
-it realy improved its handwritting, and that the responses it recieved
-was written by Vincent. To prove that it actualy improved its writing,
-it gave to Clem thirty minutes for a final training. After the training, it
-asked to Clem to write some words. Finally, he admitted that Clem actually
-improved its handwriting.
+During this first case study, we invited and followed Vincent, a six years old
+child, once per week over a period of a month. Our primary aim was to address the
+question of whether we can sustain Vincent's engagement and commitment to the
+writing activity over such a period.
 
@@
     \includegraphics[width=0.5\linewidth]{diego}
-    \caption{\small Vincent working with Nao}
+    \caption{\small Vincent correcting {\sc nao}'s attempt by rewriting the
+        whole word. Empty boxes are drawn on the screen to serve as template for the child
+        and to make word segmentation more robust.}
     \label{fig:diego}
@@
 
+The study took place at our laboratory (Figure~\ref{fig:diego}), and we chose to
+design the activity around a storyline meant to be attractive for a 6 years old
+boy: one of our {\sc nao} was away for a mysterious scientific mission, and it
+needed the support of another one -- which stayed at the lab -- to interpret
+curious pictures that were sent every week. Vincent had to help the second robot
+understanding the pictures, and since the two robots had somehow beforehand
+agreed to communicate with ``letters written like humans'' (\ie handwritten),
+Vincent also had to help the robot to write good looking letters (because, well,
+this robot was terrible at writing!) The experimental setup was similar to
+Figure~\ref{experimental_setup}, except that Vincent had to tell the robot what
+to write with small plastic letters (visible behind the robot on Figure~\ref{fig:diego}).
+
+To complement the intrinsic motivation of helping a robot to communicate with another one, we
+gradually increased the complexity of Vincent's task to keep it challenging and
+interesting (the first week: demonstration of single letters; the second week:
+short words; the third week: a full letter -- Figure~\ref{fig:stimuli}).
+
+The last session was set as a test: the ``explorer'' robot
+had come back from its mission and it actually challenged the other robot in
+front of Vincent: \emph{``I don't believe you wrote yourself these nice letters that I
+received! Prove it to me by writing something in front of me!''} This situation
+was meant to evidence the Prot\'eg\'e effect: by judging the other robot's
+handwriting, the ``explorer'' robot would implicitly judge Vincent's skills as
+teacher, and in turn, Vincent's handwriting.
+
 \subsubsection{Results}
-In all sessions, Vincent was focused on the activity during forty minutes up to
-one hour. In total, he provided 154 letters as demonstration.\\
-
-Vincent accepted our scenario and seriously played the game. At the
-beginning he seemed not confident and intimidated.  But quickly, he gain in
-confidence to play with the robot. He even showed real affective bonds : he
-sent an handwritten mail to it some days after the last session.\\
-
-As a technical result, we do not know if Vincent improved his handwriting but he
-likely conviced himself he was good enough at writing to help someone else (the
-robot in that case). To show to Vincent how much he improved the robot's
-handwriting, we printed the letter that the robot wrote during the third session
-before and after his help (Fig.~\ref{fig:stimuli}).
 
@@
 
-    \caption{\small Text generated by the robot, before and after training with
-        the child (in French). As an example, the red box emphasizes the changes on the word
-    ``envoyer''.}
+    \caption{\small (French) text generated by the robot, before and after a one
+        hour long interaction session with the child. As an example, the red box
+        highlights the changes on the word ``envoyer''.}
+
     \label{fig:stimuli}
 \end{figure}
+
+Overall, Vincent provided 154 demonstrations to the robot, and he remained
+actively engaged over the four weeks. The story was well accepted by Vincent and
+he seriously engaged into the game. After the first week, he showed good
+confidence to play with the robot and he built affective bonds with the robot
+over the course of the study, as evidenced by some cries on the last session,
+and several letters sent by him to the robot \emph{after} the end of the study
+(one of them 4 months later) to get news. This represents a promising initial
+result: we can effectively keep a child engaged with the robot for a relatively
+long period of time (about 5 hours).
+
+No conclusion can be drawn in terms of actual handwriting remediation: we did
+not design this study to formally assess possible improvements.
+
+However, as pictured on Figure~\ref{fig:stimuli}, Vincent was able to
+significantly improve the robot's skill, and he acknowledged that he had been
+able to help the robot: in that regard, Vincent convinced himself that he was
+``good enough'' at writing to help someone else, and this is likely to have
+positively impacted his self-esteem.
 
@@
         hours) around a fundamentally routine -- yet challenging -- educational
-        task (hand-writing learning),
+        task (handwriting learning),
 

### 2015-07-10 main.tex
Minor fixes in Diego study

@@
 
-\subsubsection{Context, Study Design}
+\paragraph{Context, Study Design}
 
@@
 
-\subsubsection{Results}
+\paragraph{Results}
 
@@
 over the course of the study, as evidenced by some cries on the last session,
-and several letters sent by him to the robot \emph{after} the end of the study
-(one of them 4 months later) to get news. This represents a promising initial
-result: we can effectively keep a child engaged with the robot for a relatively
-long period of time (about 5 hours).
+and a letter sent by him to the robot 4 months \emph{after} the end of the study
+to get news. This represents a promising initial result: we can effectively keep
+a child engaged with the robot for a relatively long period of time (about 5
+hours).
 
 No conclusion can be drawn in terms of actual handwriting remediation: we did
-not design this study to formally assess possible improvements.
-
-However, as pictured on Figure~\ref{fig:stimuli}, Vincent was able to
-significantly improve the robot's skill, and he acknowledged that he had been
-able to help the robot: in that regard, Vincent convinced himself that he was
-``good enough'' at writing to help someone else, and this is likely to have
-positively impacted his self-esteem.
+not design this study to formally assess possible improvements.  However, as
+pictured on Figure~\ref{fig:stimuli}, Vincent was able to significantly improve
+the robot's skill, and he acknowledged that he had been able to help the robot:
+in that regard, Vincent convinced himself that he was ``good enough'' at writing
+to help someone else, and this is likely to have positively impacted his
+self-esteem.
 

### 2015-07-10 main.tex
Reworked Henry section

@@
 and struggles at school with his poor handwriting and even poorer
-self-confidence.
+self-confidence.\footnote{Children's names have been changed.}
 
@@
     }
-    \subfigure[Same samples, normalized and projected in the eigen space spanned from the
+    \subfigure[Same samples, normalized and projected in the eigenspace spanned from the
         3 first eigenvectors: clusters arise, that actually match writing styles.]{
@@
 
-\subsubsection{Context, Study Design}
+\paragraph{Context, Study Design}
 
-This study focused technical adaptations of the CoWriter activity for a
-child diagnosed with real writing deficits.\\
+The second long-term study was designed with the help of an occupational therapist in
+Geneva and aimed at deploying the system for the first time on a real
+therapeutic case.
 
-Thomas is a six years old child. It is under the care of an occupational
-therapist. According to the therapist, he shows visuo-constructive deficits.
-As an effect in writing activities, he was frequently performing random attempts and then was comparing
-with the provided template. What is more, Thomas is strongly careless : he rarely payed attention to
-advices, even to what he was doing when he was currently drawing, and he was
-quickly shifting his attention from one
-activity to another.\\
+Thomas is a 5.5 years old child, under the care of an occupational therapist. He
+has been diagnosed with visuo-constructive deficits, which translate into
+difficulties for him to consistently draw letters. Besides, focusing on a task
+is difficult for Thomas, who tends to rapidly shift his attention to something
+else. Since the robot's learning algorithm requires repeated demonstrations of
+similarly shaped letters to converge, the occupational therapist was especially
+interested in observing if the robot would induce a strong enough motivation for
+Thomas to focus on producing regular, consistent letters, thus overcoming his
+deficit.
 
-Our objective is to investigate small modifications of the activity adapted to
-the troubles of Thomas (visuo-constructive deficits and inattention) in order to
-sustain him focused
-on the activity during forty-minutes session, and to make the robot
-evidently learning from his demonstrations.\\
-
-The experiment took place in the therapist's office. It was divided in four
-sessions. This time, we assumed that a scenario like the one we used for Vincent was
-no longer relevant with Thomas. We just introduced the robot and quickly
-said that it was seeking help to train for a robot handwriting contest.\\
-
-Before the experiment, Thomas was working on writing numbers with the therapist.
-Hence we decided to turn the CoWriter activity to teach numbers to the robot. Build on
-existing iPad applications actually used by occupational therapists (Dawson Toth's \emph{ABC's
-Writer}), we designed an Android application for pre-test and post-test. It consisted in drawing
-numbers following an helping pattern, that becomes more fine with levels in
-order to increase difficulty~\ref{fig:abc-writer}.\\
-
-Since Thomas was frequently drawing horizontally-inverted numbers, or even
-unrecognizable allographs, the learning algorithm of the robot was converging to
-meaningless scrawls. To fix this problem, we programmed the robot to refuse allographs that
-were too distant to a reference with a threshold we arbitrary fixed. In that way,
-the child was forced to take care on what he was providing to the robot as
-demonstration. Also, to make sure the robot was going to improve its handwriting
-and to clearly show this improvement, we decided to make it starting
-from scratch : for all numbers, the first try of the robot resulted in
-a simple vertical stroke (see the first robot's try
-in~\ref{learning_6_demos}).\\
-
-In this setup, we added a second tablet with one button per numbers. It was used
-by the child to chose a new number to teach to the robot. It also provided the
-possiblity to enter letters or words, and to switch to another activity (the
-robot telling a story).
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\linewidth]{abc-writer}
-    \caption{\small Screenshot of the Android application developed to be used as
-    pre-test and post-test: the child must follow with his finger the path of
-the letter or number. We count the number of times the finger goes outside of
-the red-bordered area.}
-    \label{fig:abc-writer}
-\end{figure}
-
-\subsubsection{Results}
-
-Despite his inattention, Thomas was able to keep on the activity during more than
-forty minutes in each session. In total, 55 allographs out of 82 
-provided by the child as demonstration were acceptable by the robot (with a
-progressive improvement from 13 out of 28 in the first session up to 26 out
-of 29 in the last session).\\
-
-As soon as Thomas understood that the robot was only accepting recognizable
-allographs, he started to take care on the allographs he was providing to it. He used the
-clearing button to retry when he judged that his numbers where not going to be
-accepted by the robot. According to the therapist, it was the first time Thomas
-was intrinsically correcting himself.\\
-
-Since the robot was learning from scratch, each time Thomas succeeded to make his
-demonstrations accepted by it, the improvement of its handwriting was
-visible~\ref{learning_6_demos},~\ref{henry_distances}.
-Therefor, Thomas quickly entered in an self-rewarding game where the rewards were
-the improvements of the robot, that enhanced his attention.
+The experiment took place at the therapist's surgery (four sessions spanning
+over 5 weeks). Contrary to Vincent's experiment, we chose not to introduce any
+storyline beyond a simple prompt (``the robot wants to participate to a robotic
+handwriting contest, will you help him to prepare?''), only provided during the
+first session. Hence, we also tested during this case-study if the robot (and
+the Prot\'eg\'e effect) would induce a strong enough intrinsic motivation to
+keep the child engage over the five weeks.
 
@@
     \includegraphics[width=0.9\linewidth]{learning_6_demos}
-    \caption{\small Thomas's demonstrations for the number ``6'' (top row) and
-        corresponding shapes generated by the robot. After eight demonstrations,
-        Thomas decided that the robot's ``6'' was good enough, and went to
-    another character: in that respect, he was the one leading the learning
-process of the robot.}
+    \caption{\small Thomas' demonstrations for the number ``6'' (top row) and
+        the corresponding shapes generated by the robot (bottom row). After
+        eight demonstrations, Thomas decided that the robot's ``6'' was good enough,
+        and went to another character. In that regard, he was the one
+        responsible for the learning process of the robot.}
+
     \label{learning_6_demos}
 \end{figure}
+
+The occupational therapist had recently carried activities with Thomas about
+writing numbers, so we decided with her to focus on these as well: Thomas would
+use a secondary tablet to tell the robot what number to write, and would then
+correct the robot's attempts like in the other experiments.
+Figure~\ref{learning_6_demos} shows the attempts/corrections cycles that
+occurred during one of the session, on the number ``6''.
+
+%Before the experiment, Thomas was working on writing numbers with the therapist.
+%Hence we decided to turn the CoWriter activity to teach numbers to the robot.
+%Build on existing iPad applications actually used by occupational therapists
+%(Dawson Toth's \emph{ABC's Writer}), we designed an Android application for
+%pre-test and post-test. It consisted in drawing numbers following an helping
+%pattern, that becomes more fine with levels in order to increase
+%difficulty~\ref{fig:abc-writer}.
+
+Since Thomas would frequently draw mirrored numbers, or hard-to-recognize
+shapes, the learning algorithm of the robot tended to converge towards
+meaningless scribblings. We addressed this issue by having the robot to
+\emph{refuse} allographs that were too far from the reference shape (\emph{``I'm
+not sure to understand what you are drawing...''}), so that the child had to
+pay good attention to what he would demonstrate to the robot. Also, to make
+the robot's progresses evident, we modified the initialization step of the
+learning algorithm to start with a roughly vertical stroke instead of a
+deformed number (round 0 on Figure~\ref{learning_6_demos}).
+
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=0.9\linewidth]{abc-writer}
+%    \caption{\small Screenshot of the Android application developed to be used as
+%    pre-test and post-test: the child must follow with his finger the path of
+%the letter or number. We count the number of times the finger goes outside of
+%the red-bordered area.}
+%    \label{fig:abc-writer}
+%\end{figure}
+%
+
+\paragraph{Results}
+
+Despite his attention deficit, Thomas was able to conduct the activity during
+more than forty minutes in each session. In total, 55 allographs out of the 82
+shapes demonstrated by the child were accepted by the robot (with a progressive
+improvement from 13 out of 28 in the first session up to 26 out of 29 in the
+last session).
+
+As soon as Thomas understood that the robot was only accepting well-formed
+allographs, he started to focus on it and he would typically draw 5 or 6 times
+the number before actually sending to the robot (the tablet let the children
+clear their drawing and try again before sending it to the robot). According to
+the therapist, it was the first time that Thomas would correct himself in such a
+way, explicitly having to reflect on how \emph{another agent} (the robot) would
+interpret and understand his writing.
+
+Since the robot's handwriting started from a simple primitive (a stroke), each
+time Thomas succeeded to have his demonstration accepted by it, the robot's
+improvement was clearly visible (as measured in Figure~\ref{henry_distances}).
+This led to a self-rewarding situation that effectively supported Thomas'
+engagement.
 
@@
     \includegraphics[width=0.9\linewidth]{learning_6_distances}
-    \caption{\small Two metrics to assess the handwriting progresses: Euclidian
-    distance in the eigenspace of the number dataset (top figure) or in
-cartesian space (bottom figure). Green lines represent the robot performance,
-blue lines Thomas's performance. The round IDs correspond to the demonstrations
-pictured on Figure~\ref{henry_demos}.}
+    \caption{\small \textbf{Assessing handwriting progress}: Euclidian
+    distance in the eigenspace of the number ``6''. Green lines represent the robot performance,
+    blue lines Thomas's performance. The round IDs ($x$-axis) correspond to the demonstrations
+pictured on Figure~\ref{learning_6_demos}.}
     \label{henry_distances}

### 2015-07-10 main.tex
Final version

@@
 
-\begin{abstract}
-
-    Robots for education are not limited to support ICT teaching, and they are
-    indeed finding new roles in the classroom. This article reports on such a
-    new paradigm for educative robots, that involves \emph{learning by teaching}
-    and strong \emph{social engagement} to help children struggling with
-    handwriting. Our system relies on machine-learning and child-robot
-    interaction with a small humanoid robot, and we present several real-world
-    studies in schools and with occupational therapists that led us to promising
-    initial results.
-
-\end{abstract}
+%\begin{abstract}
+%
+%    Robots for education are not limited to support ICT teaching, and they are
+%    indeed finding new roles in the classroom. This article reports on such a
+%    new paradigm for educative robots, that involves \emph{learning by teaching}
+%    and strong \emph{social engagement} to help children struggling with
+%    handwriting. Our system relies on machine-learning and child-robot
+%    interaction with a small humanoid robot, and we present several real-world
+%    studies in schools and with occupational therapists that led us to promising
+%    initial results.
+%
+%\end{abstract}
 

### 2015-12-11 main.tex
Imported some bits (including Normandy study) from HRI paper

@@
 \documentclass{article}
+
+\usepackage[utf8]{inputenc}
 
@@
 \usepackage{paralist}
-
+\usepackage{amsfonts}
 \usepackage{hyperref}
@@
 
-\author{S\'everin Lemaignan$^1$, Alexis Jacq$^{1,2}$, Fernando Garcia$^1$,
+\author{Séverin Lemaignan$^1$, Alexis Jacq$^{1,2}$, Fernando Garcia$^1$,
     Deanna Hood$^1$, \\Aude
     Billard$^1$, Ana Paiva$^2$, Pierre Dillenbourg$^1$ \\
-$^1$CHILI Lab, \'Ecole Polytechnique F\'ed\'erale de Lausanne, Suisse,\\
-$^2$Instituto Superior T\'{e}cnico, University of Lisbon, Portugal}
+$^1$CHILI Lab, École Polytechnique Fédérale de Lausanne, Suisse,\\
+$^2$Instituto Superior Técnico, University of Lisbon, Portugal}
 
@@
 Besides the commitment of the child into the interaction build on the
-``prot\'eg\'e effect'': the teacher feels responsible for his student, commits
+``protégé effect'': the teacher feels responsible for his student, commits
 to the student's success and possibly experiences student's failure as his own
 failure to teach. Teachable computer-based agents have previously been used to
-encourage this ``prot\'eg\'e effect'', wherein students invest more effort into
+encourage this ``protégé effect'', wherein students invest more effort into
 learning when it is for a teachable agent than for themselves~\cite{Chase2009}.
@@
 interacting with an agent, so I need to reflect on how to best teach him'')
-which is beneficial for the learning process; it also induces a prot\'eg\'e
+which is beneficial for the learning process; it also induces a protégé
 effect (``I want my robot-agent to succeed!'') which support the commitment of
@@
 received! Prove it to me by writing something in front of me!''} This situation
-was meant to evidence the Prot\'eg\'e effect: by judging the other robot's
+was meant to evidence the Protégé effect: by judging the other robot's
 handwriting, the ``explorer'' robot would implicitly judge Vincent's skills as
@@
 over the course of the study, as evidenced by some cries on the last session,
-and a letter sent by him to the robot 4 months \emph{after} the end of the study
-to get news. This represents a promising initial result: we can effectively keep
-a child engaged with the robot for a relatively long period of time (about 5
-hours).
+and several letters sent to the robot \emph{after} the end of the study
+(one of them 4 months later) to get news. This represents a promising initial
+result: we can effectively keep a child engaged with the robot for a relatively
+long periods of time (about 5 hours).
 
 No conclusion can be drawn in terms of actual handwriting remediation: we did
-not design this study to formally assess possible improvements.  However, as
-pictured on Figure~\ref{fig:stimuli}, Vincent was able to significantly improve
-the robot's skill, and he acknowledged that he had been able to help the robot:
-in that regard, Vincent convinced himself that he was ``good enough'' at writing
-to help someone else, and this is likely to have positively impacted his
-self-esteem.
+not design this study to formally assess possible improvements. However, as pictured on Figure~\ref{fig:stimuli}, Vincent was able to
+significantly improve the robot's skill, and he acknowledged that he had been
+able to help the robot: in that regard, Vincent convinced himself that he was
+``good enough'' at writing to help someone else, which is likely to have
+a positive impact on his self-esteem.
 
@@
 first session. Hence, we also tested during this case-study if the robot (and
-the Prot\'eg\'e effect) would induce a strong enough intrinsic motivation to
+the Protégé effect) would induce a strong enough intrinsic motivation to
 keep the child engage over the five weeks.
@@
     \includegraphics[width=0.9\linewidth]{learning_6_demos}
-    \caption{\small Thomas' demonstrations for the number ``6'' (top row) and
-        the corresponding shapes generated by the robot (bottom row). After
-        eight demonstrations, Thomas decided that the robot's ``6'' was good enough,
-        and went to another character. In that regard, he was the one
-        responsible for the learning process of the robot.}
-
+    \caption{\small Demonstrations provided by Thomas for the number ``6'' (top row) and
+        corresponding shapes generated by the robot. After eight demonstrations,
+        Thomas decided that the robot's ``6'' was good enough, and went to
+    another character: in that respect, he was the one leading the learning
+process of the robot.}
     \label{learning_6_demos}
@@
 
-Despite his attention deficit, Thomas was able to conduct the activity during
-more than forty minutes in each session. In total, 55 allographs out of the 82
-shapes demonstrated by the child were accepted by the robot (with a progressive
-improvement from 13 out of 28 in the first session up to 26 out of 29 in the
-last session).
+Despite his attention deficit, Thomas was able to remain engaged in the activity during more than
+forty minutes in each session. In total, 55 allographs out of 82 
+demonstrated by the child were acceptable considering our threshold (with a
+progressive improvement from 13 out of 28 in the first session up to 26 out
+of 29 in the last session).
 
@@
 allographs, he started to focus on it and he would typically draw 5 or 6 times
-the number before actually sending to the robot (the tablet let the children
-clear their drawing and try again before sending it to the robot). According to
+the number before actually sending to the robot (the tablet lets children
+clear their drawing and try again before sending it). According to
 the therapist, it was the first time that Thomas would correct himself in such a
 way, explicitly having to reflect on how \emph{another agent} (the robot) would
-interpret and understand his writing.
+interpret and understand his writing. Figure~\ref{Thomas_progress} shows how
+he gradually improved his demonstrations for some numbers, according to the
+metric we used to make the robot accept/refuse samples.
 
@@
 time Thomas succeeded to have his demonstration accepted by it, the robot's
-improvement was clearly visible (as measured in Figure~\ref{henry_distances}).
+improvement was clearly visible (as measured in Figure~\ref{Thomas_distances}).
 This led to a self-rewarding situation that effectively supported Thomas'
@@
     \includegraphics[width=0.9\linewidth]{learning_6_distances}
-    \caption{\small \textbf{Assessing handwriting progress}: Euclidian
-    distance in the eigenspace of the number ``6''. Green lines represent the robot performance,
-    blue lines Thomas's performance. The round IDs ($x$-axis) correspond to the demonstrations
-pictured on Figure~\ref{learning_6_demos}.}
-    \label{henry_distances}
+    \caption{\small Two metrics to assess the handwriting progresses: Euclidean
+    distance in the subspace of the number dataset (top figure) or in
+Cartesian space (bottom figure). Green lines represent the robot performance,
+blue lines performance of the child. The round IDs correspond to the demonstrations
+pictured on Figure~\ref{Thomas_demos}.}
+    \label{Thomas_distances}
 \end{figure}
+
+\begin{figure}
+    \centering
+    \subfigure[number 1]{
+        \includegraphics[width=6cm]{henry1}
+    }
+    \subfigure[number 2]{
+        \includegraphics[width=6cm]{henry2}
+    }
+	\subfigure[number 5]{
+        \includegraphics[width=6cm]{henry5}
+    }
+    \caption{\small Improvement of Thomas demonstrations for some numbers: a) drawing the allograph of the number 1, b) the number 2 and c) the number 5.
+Thomas progressively took care of the demonstrations he was providing to the robot for those numbers.
+We used for this figure the same metric than the one used for the acceptance algorithm.
+Distances are normalized with respect to the biggest value.
+The dashed line correspond to the threshold of robot's acceptance.}
+
+    \label{Thomas_progress}
+\end{figure}
+
+\section{Case study 3: when children evaluate the robot}\label{auto}
+
+\subsection{Context}
+
+Each of previous studies was adapted to individual subjects : we used a special
+design for each case in order to sustain the child's engagement.
+In our next experiment, we conducted case studies with eight children using a single
+design. Those children have in common difficulties to learn
+cursive writing but the natures and intensities of those troubles are radically
+different from one child to another. Valentine (7 years old), Alexandre (6.5) and
+Jonathan (7) are under the care of an
+occupational therapist. Enzo (8) and Matenzo (7) are repeating their school year
+because of writing. Mona (6) and Adele (8) are bottom of their respective
+classes in writing activities. Nathan (7) is under the care of a neurologist, and
+has been diagnosed with specific language impairment. All of those children are
+expected, given their school level, to know the allographs of
+cursive letters.
+
+\subsection{Questions}
+
+The main purpose of this study is to test the ergonomic aspects of CoWriter. When introducing the robot, we did not
+provide scenario and gave to children minimal explanations. We then observed how easily children assumed the role of the teacher and how seriously they tried to help the robot.
+
+\subsection{Experimental settings}
+
+This experiment took place in the coffee room of a therapists shared surgery
+in Normandy, France. Over two weeks, each child came three time for one-hour
+session, except Adele and Mona who did just one session.
+A facilitator was present to explain the rules of the game and tablet usage.
+As in the previous experiment, children were provided with two tablets : one to choose a word (or a single letter) to teach, and one
+used by both the child and the robot to write. We also provided the allograph's template if the child asked for.
+
+The starting point of the robot's writing was the same for all children: we
+used middle point between simple vertical strokes and letters. For this study,
+we wanted the robot to be only influenced by the demonstrations provided by the
+child, so we did not project allographs in a subspace. The new generated
+sample by the robot were calculated as the middle way between demonstration and
+last state in Cartesian space.
+
+The robot was programmed to accept all demonstrations, giving to the child the
+full responsibility of the teacher.
+
+We added two buttons on the tablet  interface: a green one with a ``thumbs up", and
+a red one with a ``thumbs down". Those buttons could be used by children to evaluate the
+robot (the green one was for rewards while the red one was for punishment). This
+way, we could measure the perception of the robot by the child: the more the
+child used evaluation buttons, the more he was playing the teacher, judging the
+robot instead of himself. It became possible to estimate if a child is playing
+seriously given correlation between his evaluation and robot's actual
+progression.
+
+\subsection{Results}
+
+All children maintained their engagement during the whole sessions. They
+provided on average 42 demonstrations per session. All children used evaluation
+buttons and had preference to reward the robot (in total, 99 rewards and 33
+punishments were recorded).
+
+Since sessions took place over only two weeks, we did not studied possible
+handwriting remediation in children, but we focused on correlation between the
+children's evaluations and the robot's progression.  We estimated the robot's
+progression as the difference between a starting score (score of the first
+robot's try when children have chosen a new word/letter to work on) and the
+current robot's score (after being taught by the child). The score is calculated
+as the average of euclidean distance between robot's try and a reference
+allograph over all letters of the word. Those references for letter allographs
+where drawn by us beforehand, taking inspiration in education.com cursive
+letters
+template\footnote{\url{http://www.education.com/slideshow/cursive-handwriting-z/}}.
+Let $P_i$ be the estimated progression of the robot at time $ i$. Of course, if
+the child chose to switch to a new word at time $ j$ , we have $ P_{j}=0$ and
+clearly $ P_0=0$.  To measure whether the fact that the child was rewarding the
+robot when it was progressing was significant or not, we generated 10000 times
+the same number of reward/punishment but accorded at random times. Let $ R_i^n$
+be the $n^{th}$ generated evaluation at time $ i$ ($ R_i^n = 0$ if no evaluation
+occurred at time $ i$ , $ R_i^n=1$ if a reward occurred at time $i$ and
+$R_i^n=-1$ if a punishment occurred at time $i$), and $\overline{R}_i$ be the
+actual evaluation at time $i$. For each $n^{th}$ generated sequence of
+evaluation, we compute a score of evaluation: $$ S^n = \sum\limits_i{R_i^n
+P_i}$$ Then we can estimate the p-value $p$ of the actual score: $$ \overline{S}
+= \sum\limits_i{\overline{R}_i P_i}$$ given the distribution of the generated
+scores $\left(S^n\right)_n$, which is assumed to be Gaussian:
+
+$$
+p(\overline{S})= \mathbb{P}{\left[X>\overline{S}\right]} = 1-\phi{\left(\frac{\overline{S}-\mu}{\sigma}\right)}
+$$
+
+where $\phi$ denotes the cumulative distribution function of the standard normal
+distribution, $\mu$ the mean and $\sigma$ the deviation of the generated scores
+$\left(S^n\right)_n$.  As a result, we found that 5 of the 8 children obtained a
+score of evaluation significantly hight ($p(\overline{S})<0.05$). Score of
+evaluation p-values of each child is reported in the second-to-last column of
+Table~\ref{table:scores}.
+
+We also studied correlation between children's evaluations and their own
+progression. The analysis was conducted in the same way, using distances between
+children demonstrations and reference allographs to compute children
+progressions.  3 of 5 children that played ``seriously" obtained score of
+evaluation of their own progression significantly high (last column of
+Table~\ref{table:scores}). For those children, it seems that the robot was
+reflecting their own performances, and while they were judging the robot
+positively (three times more rewards than punishments) they were actually
+evaluating themselves.
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|c|c|c|}
+        \hline
+        Child & demo & rew & pun & p(robot) & p(child)\\ \hline
+        Valentine & 127 & 24 & 6 & 2.4e-03 & 5.5e-02\\ \hline
+        Enzo & 223 & 20 & 9 & 1.7e-01 & 3.5e-01\\ \hline
+        Matenzo & 131 & 10 & 3 & 3.8e-03 & 7.9e-03\\ \hline
+        Jonathan & 98 & 10 & 5 &  1.5e-01 & 3.8e-01\\ \hline
+        Nathan & 115 & 16 & 4 & 5.3e-04 & 2.7e-03\\ \hline
+        Alexandre & 83 & 10 & 3 & 3.1e-02 & 6.0e-01\\ \hline
+        Adele & 35 & 4 & 2 & 5.0e-02 & 3.7e-02\\ \hline
+        Mona & 40 & 5 & 1 &  5.4e-01 & 2.0e-01\\ \hline
+    \end{tabular}
+    \caption{\small Results of evaluations. demo: number of demonstrations provided by
+        the child over all session. rew: number of rewards accorded by
+        the child. pun: number of punishments. p(robot): are the evaluations
+        significantly corresponding to the progression of the robot ? p(child): are the
+        evaluations significantly corresponding to the child's own
+progression ?}
+    \label{table:scores}
+\end{table}
 
@@
 \begin{itemize}
-    \item robots in an educative context are \textbf{relevant and effective beyond ICT
-        teaching},
+    \item robots in an educative context are \textbf{relevant and effective
+        beyond STEM} (Science, Technology, Engineering and Mathematics)
+    \textbf{teaching},
 

### 2015-12-11 main.tex
Expanded discussion on ethical concerns

@@
 
-The place and role of the robot \textit{vis-\`a-vis} the teacher can be
+\paragraph{Possible Ethical Concerns} Two aspects of this research ought to be
+discussed in term of their possible ethical implications: the perceived role of
+the robot \textit{vis-à-vis} the teacher, and the implications of the
+mentor-protégé relationship for children, especially vulnerable ones.
+
+The place and role of the robot \textit{vis-à-vis} the teacher can be
 questioned: as we see it, the role of the robot within the classroom (or at the
@@
 
+The implication of the mentor-protégé relationship on the children is less
+clearly understood. The children can establish strong affective bonds with the robot (as
+witnessed by the letter sent by Vincent several months after he interacted with
+the robot), but we are not yet able to precisely characterize these bonds. The
+ethical implication of the mentor-protégé relationship have been explored before
+in the context of human teaching~\cite{brad1999mentor,wendelyn2008context}, but
+they mostly look at the question from the perspective of the protégé, whereas in our
+case, the child is the mentor. As such, relatively little is known on the
+psychological implications for a child to commit in helping a robot, and as
+advocated by Belpaeme and Morse~\cite{belpaeme2010time}, we likely need to first
+gain more field experience before being able to draw conclusions.
+
 Beyond handwriting, we however do believe that this work provides a novel
@@
 robots: when facing a child with school difficulties, robots can play the role
-of a na\"ive learner which neither adults nor peers -- because of the social
+of a naïve learner which neither adults nor peers -- because of the social
 effects it would induce -- can convincingly play.

### 2015-12-11 main.tex
Started to rewrite the Normandy study

@@
 
-To date, the system has been deployed and tested in three different
-schools (totalling child-robot interactions with more than 70 children, aged 5
-to 8) and we also conducted 2 case studies that lasted several weeks.
-Table~\ref{studies} gives an overview of these studies.
+The system has been deployed and tested in several situations: in three
+different schools (resulting in relatively short child-robot interactions with
+more than 70 children, aged 5 to 8), in the surgery of an occupational
+therapist (with 8 children, each interacting several hours with the system),
+and for two case studies that each lasted several weeks. Table~\ref{studies}
+gives an overview of these studies.
 
@@
 {\it ISG 2}     & Individual interaction at school              & 18 min                            & 6                                     & 5-6        \\
-{\it Thomas}     & Case-study, spanning over 4 weeks              & 1h $\times$ 4 weeks        & 1                                     & 5          \\ \bottomrule
+{\it Thomas}     & Case-study, spanning over 4 weeks              & 1h $\times$ 4 weeks        & 1                                     & 5          \\
+{\it Normandie}     & Individual interactions at surgery               & 1h $\times$ 3 sessions & 8                                     & 6-8          \\ \bottomrule
 \end{tabular}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Studies at School}
+\subsection{System Validation at Schools}
 
@@
 \emph{acceptance and impact in a real therapeutic context}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Study: How Children Take on the Role of a Teacher}\label{normandie}
+
+\paragraph{Context, Study Design}
+
+\begin{inparaenum}
+This experiment was conducted to study \item how easily children with actual
+deficit take on the role of a teacher, and \item if they engage \emph{seriously}
+in this role.
+\end{inparaenum}
+
+The experiment took place at the surgery of an occupational therapist in
+Normandy, France. Eight children participated, all showing troubles with
+handwriting acquisition: Valentine (7 years old), Alexandre (6.5) and Jonathan
+(7) are under the care of the occupational therapist. Enzo (8) and Matenzo (7)
+are repeating their school year because of writing difficulties. Mona (6) and
+Adele (8) are both ranked at the bottom of their respective classes in writing
+activities. Nathan (7) is under the care of a neurologist, and has been
+diagnosed with specific language impairment. Given their school level, all those
+children would be expected to know how to correctly shape cursive letters.
+
+Over a period of two weeks, each child attended three time a one hour long
+session, except for Adele and Mona who only attended one session. The
+experimenter role was limited to the explaination of the rules of the game and
+the basic tablet usage. For this experiment, the children were provided with two
+tablets: one to choose the word (or single letter) to teach to the robot, one
+used to write. We also provided templates of the cursive letters, where the
+children to ask for.
+
+We only provided the children with minimal explanations on the task (they would
+have to help the robot to improve its writing style), so as to assess if and how
+children would naturally take on the role of the teacher, and how seriously they
+engaged into helping the robot.
+
+This assessment was carried through two additional buttons on the tablet: a
+green ``thumbs up'' and a red ``thumbs down''. The children were told to use
+them to evaluate the robot's improvements (``thumb up'' being a way to give
+positive feedback while ``thumb down'' would be convey negative feedback). Our
+assessment builds on the hypothesis that the more the child provides feedback to
+the robot with the green/red buttons, the more he/she assumes the role of a
+teacher. Then, by correlating the feedback with the actual performance of the
+robot, we can measure to what extend the child was seriously assuming his/her
+role: since the child knows what a correct letter should look like, if his/her
+feedback does not correlate with the actual performance, the child was likely
+engaged into a playful instead of serious role-playing.
+
+\paragraph{Results}
+
+All children maintained their engagement during the whole sessions. They
+provided on average 42 demonstrations per session. All children used the
+feedback buttons and had preference to give positive feedback the robot (99
+``thumb up'' \vs 33 ``thumb down'' were recorded).
+
+As the children participated at most 3 hours of interactions spanning over two
+week, we did not attempted to measure possible handwriting improvements by the
+children, and instead we focused on the possible correlation between the
+children's feedback and the robot's progression.  We estimated the robot's
+progression as the difference between a starting score (score of the first
+robot's try when children have chosen a new word/letter to work on) and the
+current robot's score (after being taught by the child). The score is calculated
+as the average of euclidean distance between robot's try and a reference
+allograph over all letters of the word. Those references for letter allographs
+where drawn by us beforehand, taking inspiration in education.com cursive
+letters
+template\footnote{\url{http://www.education.com/slideshow/cursive-handwriting-z/}}.
+Let $P_i$ be the estimated progression of the robot at time $ i$. Of course, if
+the child chose to switch to a new word at time $ j$ , we have $ P_{j}=0$ and
+clearly $ P_0=0$.  To measure whether the fact that the child was rewarding the
+robot when it was progressing was significant or not, we generated 10000 times
+the same number of reward/punishment but accorded at random times. Let $ R_i^n$
+be the $n^{th}$ generated evaluation at time $ i$ ($ R_i^n = 0$ if no evaluation
+occurred at time $ i$ , $ R_i^n=1$ if a reward occurred at time $i$ and
+$R_i^n=-1$ if a punishment occurred at time $i$), and $\overline{R}_i$ be the
+actual evaluation at time $i$. For each $n^{th}$ generated sequence of
+evaluation, we compute a score of evaluation: $$ S^n = \sum\limits_i{R_i^n
+P_i}$$ Then we can estimate the p-value $p$ of the actual score: $$ \overline{S}
+= \sum\limits_i{\overline{R}_i P_i}$$ given the distribution of the generated
+scores $\left(S^n\right)_n$, which is assumed to be Gaussian:
+
+$$
+p(\overline{S})= \mathbb{P}{\left[X>\overline{S}\right]} = 1-\phi{\left(\frac{\overline{S}-\mu}{\sigma}\right)}
+$$
+
+where $\phi$ denotes the cumulative distribution function of the standard normal
+distribution, $\mu$ the mean and $\sigma$ the deviation of the generated scores
+$\left(S^n\right)_n$.  As a result, we found that 5 of the 8 children obtained a
+score of evaluation significantly hight ($p(\overline{S})<0.05$). Score of
+evaluation p-values of each child is reported in the second-to-last column of
+Table~\ref{table:scores}.
+
+We also studied correlation between children's evaluations and their own
+progression. The analysis was conducted in the same way, using distances between
+children demonstrations and reference allographs to compute children
+progressions.  3 of 5 children that played ``seriously" obtained score of
+evaluation of their own progression significantly high (last column of
+Table~\ref{table:scores}). For those children, it seems that the robot was
+reflecting their own performances, and while they were judging the robot
+positively (three times more rewards than punishments) they were actually
+evaluating themselves.
+
+\begin{table}
+    \centering
+    \begin{tabular}{|c|c|c|c|c|c|}
+        \hline
+        Child & demo & rew & pun & p(robot) & p(child)\\ \hline
+        Valentine & 127 & 24 & 6 & 2.4e-03 & 5.5e-02\\ \hline
+        Enzo & 223 & 20 & 9 & 1.7e-01 & 3.5e-01\\ \hline
+        Matenzo & 131 & 10 & 3 & 3.8e-03 & 7.9e-03\\ \hline
+        Jonathan & 98 & 10 & 5 &  1.5e-01 & 3.8e-01\\ \hline
+        Nathan & 115 & 16 & 4 & 5.3e-04 & 2.7e-03\\ \hline
+        Alexandre & 83 & 10 & 3 & 3.1e-02 & 6.0e-01\\ \hline
+        Adele & 35 & 4 & 2 & 5.0e-02 & 3.7e-02\\ \hline
+        Mona & 40 & 5 & 1 &  5.4e-01 & 2.0e-01\\ \hline
+    \end{tabular}
+    \caption{\small Results of evaluations. demo: number of demonstrations provided by
+        the child over all session. rew: number of rewards accorded by
+        the child. pun: number of punishments. p(robot): are the evaluations
+        significantly corresponding to the progression of the robot ? p(child): are the
+        evaluations significantly corresponding to the child's own
+progression ?}
+    \label{table:scores}
+\end{table}
 
@@
     \centering
-    \subfigure[number 1]{
-        \includegraphics[width=6cm]{henry1}
+    \subfigure[Number ``2'']{
+        \includegraphics[width=0.45\linewidth]{henry2}
     }
-    \subfigure[number 2]{
-        \includegraphics[width=6cm]{henry2}
-    }
-	\subfigure[number 5]{
-        \includegraphics[width=6cm]{henry5}
+	\subfigure[Number ``5'']{
+        \includegraphics[width=0.45\linewidth]{henry5}
     }
@@
 \end{figure}
-
-\section{Case study 3: when children evaluate the robot}\label{auto}
-
-\subsection{Context}
-
-Each of previous studies was adapted to individual subjects : we used a special
-design for each case in order to sustain the child's engagement.
-In our next experiment, we conducted case studies with eight children using a single
-design. Those children have in common difficulties to learn
-cursive writing but the natures and intensities of those troubles are radically
-different from one child to another. Valentine (7 years old), Alexandre (6.5) and
-Jonathan (7) are under the care of an
-occupational therapist. Enzo (8) and Matenzo (7) are repeating their school year
-because of writing. Mona (6) and Adele (8) are bottom of their respective
-classes in writing activities. Nathan (7) is under the care of a neurologist, and
-has been diagnosed with specific language impairment. All of those children are
-expected, given their school level, to know the allographs of
-cursive letters.
-
-\subsection{Questions}
-
-The main purpose of this study is to test the ergonomic aspects of CoWriter. When introducing the robot, we did not
-provide scenario and gave to children minimal explanations. We then observed how easily children assumed the role of the teacher and how seriously they tried to help the robot.
-
-\subsection{Experimental settings}
-
-This experiment took place in the coffee room of a therapists shared surgery
-in Normandy, France. Over two weeks, each child came three time for one-hour
-session, except Adele and Mona who did just one session.
-A facilitator was present to explain the rules of the game and tablet usage.
-As in the previous experiment, children were provided with two tablets : one to choose a word (or a single letter) to teach, and one
-used by both the child and the robot to write. We also provided the allograph's template if the child asked for.
-
-The starting point of the robot's writing was the same for all children: we
-used middle point between simple vertical strokes and letters. For this study,
-we wanted the robot to be only influenced by the demonstrations provided by the
-child, so we did not project allographs in a subspace. The new generated
-sample by the robot were calculated as the middle way between demonstration and
-last state in Cartesian space.
-
-The robot was programmed to accept all demonstrations, giving to the child the
-full responsibility of the teacher.
-
-We added two buttons on the tablet  interface: a green one with a ``thumbs up", and
-a red one with a ``thumbs down". Those buttons could be used by children to evaluate the
-robot (the green one was for rewards while the red one was for punishment). This
-way, we could measure the perception of the robot by the child: the more the
-child used evaluation buttons, the more he was playing the teacher, judging the
-robot instead of himself. It became possible to estimate if a child is playing
-seriously given correlation between his evaluation and robot's actual
-progression.
-
-\subsection{Results}
-
-All children maintained their engagement during the whole sessions. They
-provided on average 42 demonstrations per session. All children used evaluation
-buttons and had preference to reward the robot (in total, 99 rewards and 33
-punishments were recorded).
-
-Since sessions took place over only two weeks, we did not studied possible
-handwriting remediation in children, but we focused on correlation between the
-children's evaluations and the robot's progression.  We estimated the robot's
-progression as the difference between a starting score (score of the first
-robot's try when children have chosen a new word/letter to work on) and the
-current robot's score (after being taught by the child). The score is calculated
-as the average of euclidean distance between robot's try and a reference
-allograph over all letters of the word. Those references for letter allographs
-where drawn by us beforehand, taking inspiration in education.com cursive
-letters
-template\footnote{\url{http://www.education.com/slideshow/cursive-handwriting-z/}}.
-Let $P_i$ be the estimated progression of the robot at time $ i$. Of course, if
-the child chose to switch to a new word at time $ j$ , we have $ P_{j}=0$ and
-clearly $ P_0=0$.  To measure whether the fact that the child was rewarding the
-robot when it was progressing was significant or not, we generated 10000 times
-the same number of reward/punishment but accorded at random times. Let $ R_i^n$
-be the $n^{th}$ generated evaluation at time $ i$ ($ R_i^n = 0$ if no evaluation
-occurred at time $ i$ , $ R_i^n=1$ if a reward occurred at time $i$ and
-$R_i^n=-1$ if a punishment occurred at time $i$), and $\overline{R}_i$ be the
-actual evaluation at time $i$. For each $n^{th}$ generated sequence of
-evaluation, we compute a score of evaluation: $$ S^n = \sum\limits_i{R_i^n
-P_i}$$ Then we can estimate the p-value $p$ of the actual score: $$ \overline{S}
-= \sum\limits_i{\overline{R}_i P_i}$$ given the distribution of the generated
-scores $\left(S^n\right)_n$, which is assumed to be Gaussian:
-
-$$
-p(\overline{S})= \mathbb{P}{\left[X>\overline{S}\right]} = 1-\phi{\left(\frac{\overline{S}-\mu}{\sigma}\right)}
-$$
-
-where $\phi$ denotes the cumulative distribution function of the standard normal
-distribution, $\mu$ the mean and $\sigma$ the deviation of the generated scores
-$\left(S^n\right)_n$.  As a result, we found that 5 of the 8 children obtained a
-score of evaluation significantly hight ($p(\overline{S})<0.05$). Score of
-evaluation p-values of each child is reported in the second-to-last column of
-Table~\ref{table:scores}.
-
-We also studied correlation between children's evaluations and their own
-progression. The analysis was conducted in the same way, using distances between
-children demonstrations and reference allographs to compute children
-progressions.  3 of 5 children that played ``seriously" obtained score of
-evaluation of their own progression significantly high (last column of
-Table~\ref{table:scores}). For those children, it seems that the robot was
-reflecting their own performances, and while they were judging the robot
-positively (three times more rewards than punishments) they were actually
-evaluating themselves.
-
-\begin{table}
-    \centering
-    \begin{tabular}{|c|c|c|c|c|c|}
-        \hline
-        Child & demo & rew & pun & p(robot) & p(child)\\ \hline
-        Valentine & 127 & 24 & 6 & 2.4e-03 & 5.5e-02\\ \hline
-        Enzo & 223 & 20 & 9 & 1.7e-01 & 3.5e-01\\ \hline
-        Matenzo & 131 & 10 & 3 & 3.8e-03 & 7.9e-03\\ \hline
-        Jonathan & 98 & 10 & 5 &  1.5e-01 & 3.8e-01\\ \hline
-        Nathan & 115 & 16 & 4 & 5.3e-04 & 2.7e-03\\ \hline
-        Alexandre & 83 & 10 & 3 & 3.1e-02 & 6.0e-01\\ \hline
-        Adele & 35 & 4 & 2 & 5.0e-02 & 3.7e-02\\ \hline
-        Mona & 40 & 5 & 1 &  5.4e-01 & 2.0e-01\\ \hline
-    \end{tabular}
-    \caption{\small Results of evaluations. demo: number of demonstrations provided by
-        the child over all session. rew: number of rewards accorded by
-        the child. pun: number of punishments. p(robot): are the evaluations
-        significantly corresponding to the progression of the robot ? p(child): are the
-        evaluations significantly corresponding to the child's own
-progression ?}
-    \label{table:scores}
-\end{table}
 

### 2015-12-15 main.tex
Added proper acknowledgements

@@
 
+We warmly thank all the children, teachers and staff of the International School
+of Geneva, the Institut Florimont and the Haut-Lac school for their
+participation and support. We express our great gratitude to Anne Gossin and [TBD] for
+inviting us to their surgery, for their involvement in the experiments and for
+bearing with our disruptive robots. And finally, we want to say a very special \emph{Thank
+you!} to Vincent, Thomas, Valérie, Émilien, Mathieu, Johan, Nicolas, Antoine,
+Adèle and Marie: by spending several hours being patient teachers for our
+robots, you have paved the way for what may become tomorrow a new way of
+learning how to deal with all these letters and words!
+
 This research was partially supported by the Funda\c{c}\~{a}o para a Ci\^{e}ncia

### 2015-12-15 main.tex
Added quote of Vincent's mother + anonymized children name + plenty of little changes

@@
 
-Thomas is five and an half, and has been diagnosed with visuo-constructive
+Thomas is five and half, and has been diagnosed with visuo-constructive
 deficits. He is under the care of an occupational therapist, and tries to
@@
 facing however the same difficulty to write in a legible manner. And, hidden
-beyond this impaired skill, psycho-social difficulties also arise: they
+beyond this impaired skill, psycho-social difficulties arise: they
 underperform at school, Thomas has to go for follow-up visits every week,
@@
 Remediations for handwriting difficulties traditionally involve long
-interventions (at least 10 weeks, \cite{Hoy2011}), essentially consisting in
+interventions (at least 10 weeks~\cite{Hoy2011}), essentially consisting in
 handwriting training with occupational therapists, and primarily addressing the
@@
 
-We present in this article a new take on this educative challenge, as a
-remediation procedure that involve a ``bad writer'' robot that is \emph{taught}
-by the child: by building on the \emph{learning by teaching} paradigm, not only
-the child practises handwriting but, as (s)he takes on the role of the teacher,
+We present in this article a new take on this educative challenge: a remediation
+procedure that involve a ``bad writer'' robot that is \emph{taught} by the
+child. By building on the \emph{learning by teaching} paradigm, not only the
+child practises handwriting but, as (s)he takes on the role of the teacher,
 (s)he also positively reinforce his/her self-esteem and motivation: his/her
 social role shifts from the ``underperformer'' to ``the one who knows and
-teaches''.
+teaches''. And by relying on a robot, we can tailor the exercises and the
+learning rate to each children' needs, as we show in this article.
 
@@
 physical presence and embodiment induce agency and anthropomorphising, and
-cognitively engage the child into the learning activity.
+cognitively engage the child into the learning activity (be it consciously or
+not).
 
-Besides the commitment of the child into the interaction build on the
-``protégé effect'': the teacher feels responsible for his student, commits
+Besides the commitment of the child into the interaction build on a
+psychological effect known as the
+``protégé effect''~\cite{Chase2009}: the teacher feels responsible for his student, commits
 to the student's success and possibly experiences student's failure as his own
@@
 We rely on this cognitive mechanism to reinforce the child's commitment into the
-robot-mediated handwriting activity.
+robot-mediated handwriting activity, and we indeed show sustained child-robot
+engagement over extended periods of time (several hours spread of a month).
 
@@
 effect (``I want my robot-agent to succeed!'') which support the commitment of
-child into the interaction, also for longer period of time (we report in this
-article upon two case studies where children were indeed interacting with the
-same robot for several hours, spanned over several weeks).
+child into the interaction, also for longer period of time.
 
@@
 The following sections of the paper go into details. We first provides a
-brief overview of the robotic system and the interaction it induces. We then
-give the details of machine-learning techniques that allow the robot to learn
+brief overview of the robotic system and the interaction it induces. We also
+present the machine-learning techniques that allow the robot to learn
 from the children, followed by the implementation on a {\sc nao} robot.
 
-We then present and report on the four school experiments and the two case
-studies that we conducted over the last two years. While the focus of the school
-experiments was mostly technical validation and data acquisition, the two case
-studies that lasted each about 4 weeks, gave us initial insights on the
+We then present and report on the field experiments that we conducted over the
+last two years, including four studies at school, one longer experiment
+with eight children at the surgery of an occupational therapist, and two
+one-month long case studies. While the focus of the school
+experiments was mostly technical validation and data acquisition, the three
+other experiments involved children with actual deficits, and gave us initial insights on the
 relevance and effectiveness of our approach.
@@
 Figure~\ref{experimental_setup} illustrates our general experimental setup: a
-face-to-face child-robot interaction with an (autonomous) Aldebran's {\sc nao}
+face-to-face child-robot interaction with an autonomous Aldebran's {\sc nao}
 robot.
@@
 something (a single letter, a number or a full word), and push the tablet
-towards the robot, the robot writes on the tablet by gesturing the writing (but
-without actually physically touching the tablet), the child then pull back the
+towards the robot. The robot writes on the tablet by gesturing the writing -- the
+letters being actually drawn by the tablet application itself. The child then pull back the
 tablet, corrects the robot's attempt by writing him/herself on top or next to
@@
 
-Since the child is assumed to take on the role of the teacher, we had to ensure
-(s)he would be able to manage by him/herself the turn-taking and the overall
-progression of the activity (moving to the next letter or word). In our design,
+Since the children are assumed to take on the role of the teachers, we had to ensure
+they would be able to manage by themselves the turn-taking and the overall
+progression of the activity (moving forwards to the next letter or word). In our design,
 the turn-taking relies on the robot prompting for feedback once it is done with
-its writing (simple sentences like ``What do you think?''), and pressing on a
+its writing (through simple sentences like ``What do you think?''), and pressing on a
 small robot icon on the tablet once the child has finished correcting. In our
@@
 several modules (from performing gestures and acquiring the user's input to
-the high-level state machine), spread over several devices (the robot itself,
-one laptop and up to four tactile tablets for certain studies we conducted). We
+the state machine implementing the high-level behaviour), spread over several devices (the robot itself,
+one laptop and up to four tactile tablets for some of the studies that we conducted). We
 relied on ROS to ensure the synchronization and communication between these
@@
 letters, with distortions that are actually plausible handwriting errors: they
-are actually \emph{exaggerations} of variance of writing styles that naturally
+are actually exaggerations of variance of writing styles that naturally
 occurs amongst adult writers.  Figure~\ref{fig:sampleLetters} shows examples of
@@
 \emph{assess} their quality and \emph{learn} from them. Figure~\ref{fig:h} shows
-for example 9 allographs of ``h'' written by a 6 years old child, along with the
+for example nine allographs of ``h'' written by a 6 years old child, along with the
 mean letter from the dataset (reference letter). By projecting each of the
@@
     \subfigure[Nine allographs of the cursive ``h'', next to the reference]{
-        \includegraphics[width=0.4\columnwidth]{h}
+        \includegraphics[width=0.45\columnwidth]{h}
     }
@@
         3 first eigenvectors: clusters arise, that actually match writing styles.]{
-        \includegraphics[width=0.4\columnwidth]{eigenspace-uniformization}
+        \includegraphics[width=0.45\columnwidth]{eigenspace-uniformization}
     }
@@
 a specific pace towards the child demonstration in the eigenspace.
-Figure~\ref{learning_6_demos}, at the end of the article, illustrates the
+Figure~\ref{learning_6_demos} (p.~\pageref{learning_6_demos}) illustrates the
 process with the complete learning cycle of the number ``6''.
@@
 spread over several devices: the {\sc nao} robot itself, that we address via
-both a ROS API\footnote{The ROS stack for {\sc nao} is available at
-\url{http://wiki.ros.org/nao_robot}.} and the Aldebaran-provided NaoQI API, one
-to four Android tablets (the main tablet is used to print the robot's letter and
+both a ROS API\footnote{The ROS documentation for {\sc nao} is available at
+\url{http://wiki.ros.org/nao}.} and the Aldebaran-provided NaoQI API, one
+to four Android tablets (the main tablet is used to draw the robot's letter and
 to acquire the children's demonstrations; more tablets have been used in some
@@
 fashion), and a central laptop running the machine learning algorithms, the
-robot's handwriting gesture generation and high level control of the activity.
+robot's handwriting gesture generation (based on the NaoQI inverse kinematics
+library) and high level control of the activity.
 
@@
 difficulty to deploy and debug ROS nodes directly on the robot, the laptop
-remains however convenient during the development phase and we kept if during
+remains however convenient during the development phase and we kept it during
 our experiments.
@@
 project is available online\footnote{The primary repository is
-\url{https://github.com/chili-epfl/cowriter_letter_learning}.}
+\url{https://github.com/chili-epfl/cowriter_letter_learning}.}. Further details
+on the technical implementation are available in~\cite{hood2015when}.
 
@@
 The system has been deployed and tested in several situations: in three
-different schools (resulting in relatively short child-robot interactions with
+different schools (relatively short child-robot interactions with
 more than 70 children, aged 5 to 8), in the surgery of an occupational
 therapist (with 8 children, each interacting several hours with the system),
-and for two case studies that each lasted several weeks. Table~\ref{studies}
+and in two case studies that each lasted several weeks. Table~\ref{studies}
 gives an overview of these studies.
@@
 {\bf Study}        & {\bf Type}                               & {\bf Avg. duration}    & {\bf Children \#}                     & {\bf Ages} \\ \midrule
-{\it ISG 1}     & Unstructured group interaction at school      & 16 min/group               & 4 $\times$ 8 children                 & 6-7        \\
-{\it Florimont} & Individual/Pair interaction at school         & 11 min/group               & 7 (individual) + 7 $\times$ 2 (pairs) & 7-8        \\
-{\it Vincent}     & Case-study, spanning over 4 weeks              & 1.5h $\times$ 4 weeks & 1                                     & 6          \\
-{\it Haut-Lac}  & Pair interaction at school & 26 min/group                            & 7 $\times$ 2                          & 5-6        \\
-{\it ISG 2}     & Individual interaction at school              & 18 min                            & 6                                     & 5-6        \\
-{\it Thomas}     & Case-study, spanning over 4 weeks              & 1h $\times$ 4 weeks        & 1                                     & 5          \\
-{\it Normandie}     & Individual interactions at surgery               & 1h $\times$ 3 sessions & 8                                     & 6-8          \\ \bottomrule
+{\it School A (1)} & Unstructured group interaction at school & 16 min/group           & 4 $\times$ 8 children                 & 6-7        \\
+{\it School B}     & Individual/Pair interaction at school    & 11 min/group           & 7 (individual) + 7 $\times$ 2 (pairs) & 7-8        \\
+{\it School C}     & Pair interaction at school               & 26 min/group           & 7 $\times$ 2                          & 5-6        \\
+{\it School A (2)} & Individual interaction at school         & 18 min                 & 6                                     & 5-6        \\ \midrule
+{\it Surgery}      & Individual interaction at surgery        & 3 sessions $\times$ 1h & 8                                     & 6-8        \\ \midrule 
+{\it Vincent}      & Case-study, spanning over 4 weeks        & 4 weeks $\times$ 1.5h  & 1                                     & 6          \\
+{\it Thomas}       & Case-study, spanning over 4 weeks        & 4 weeks $\times$ 1h    & 1                                     & 5          \\ \bottomrule
 \end{tabular}
 \end{table}
+
+We report hereafter the main design choices and results for each of these
+studies and experiments. The interested reader can find supplementary details
+in~\cite{jacq2016building, hood2015when}.
 
@@
     \caption{\small The first field studies were focused on technical validation, with
-    above 70 pupils interacting with the robot over short periods (around 15
+    more than 70 pupils interacting with the robot over short periods (around 15
 minutes), either alone or in small groups.}
@@
 Normandy, France. Eight children participated, all showing troubles with
-handwriting acquisition: Valentine (7 years old), Alexandre (6.5) and Jonathan
-(7) are under the care of the occupational therapist. Enzo (8) and Matenzo (7)
+handwriting acquisition: Valérie (7 years old), Antoine (6.5) and Johan
+(7) are under the care of the occupational therapist. Émilien (8) and Mathieu (7)
 are repeating their school year because of writing difficulties. Mona (6) and
-Adele (8) are both ranked at the bottom of their respective classes in writing
-activities. Nathan (7) is under the care of a neurologist, and has been
+Adèle (8) are both ranked at the bottom of their respective classes in writing
+activities. Nicolas (7) is under the care of a neurologist, and has been
 diagnosed with specific language impairment. Given their school level, all those
@@
         \hline
-        Child & demo & rew & pun & p(robot) & p(child)\\ \hline
-        Valentine & 127 & 24 & 6 & 2.4e-03 & 5.5e-02\\ \hline
-        Enzo & 223 & 20 & 9 & 1.7e-01 & 3.5e-01\\ \hline
-        Matenzo & 131 & 10 & 3 & 3.8e-03 & 7.9e-03\\ \hline
-        Jonathan & 98 & 10 & 5 &  1.5e-01 & 3.8e-01\\ \hline
-        Nathan & 115 & 16 & 4 & 5.3e-04 & 2.7e-03\\ \hline
-        Alexandre & 83 & 10 & 3 & 3.1e-02 & 6.0e-01\\ \hline
-        Adele & 35 & 4 & 2 & 5.0e-02 & 3.7e-02\\ \hline
-        Mona & 40 & 5 & 1 &  5.4e-01 & 2.0e-01\\ \hline
+        child & demo & rew & pun & p(robot) & p(child)\\ \hline
+        Valérie & 127 & 24 & 6 & 2.4e-03 & 5.5e-02\\ \hline
+        Émilien & 223 & 20 & 9 & 1.7e-01 & 3.5e-01\\ \hline
+        Mathieu & 131 & 10 & 3 & 3.8e-03 & 7.9e-03\\ \hline
+        Johan & 98 & 10 & 5 &  1.5e-01 & 3.8e-01\\ \hline
+        Nicolas & 115 & 16 & 4 & 5.3e-04 & 2.7e-03\\ \hline
+        Antoine & 83 & 10 & 3 & 3.1e-02 & 6.0e-01\\ \hline
+        Adèle & 35 & 4 & 2 & 5.0e-02 & 3.7e-02\\ \hline
+        Marie & 40 & 5 & 1 &  5.4e-01 & 2.0e-01\\ \hline
     \end{tabular}
-    \caption{\small Results of evaluations. demo: number of demonstrations provided by
+    \caption{\small results of evaluations. demo: number of demonstrations provided by
         the child over all session. rew: number of rewards accorded by
@@
 
-During this first case study, we invited and followed Vincent, a six years old
+For the first case study, we invited and followed Vincent, a six years old
 child, once per week over a period of a month. Our primary aim was to address the
@@
 boy: one of our {\sc nao} was away for a mysterious scientific mission, and it
-needed the support of another one -- which stayed at the lab -- to interpret
+needed the support of another one -- which would remain at the lab -- to interpret
 curious pictures that were sent every week. Vincent had to help the second robot
@@
     \centering
-    \subfigure[Initial letter, generated by the robot]{
+    \subfigure[Initial text, generated by the robot]{
         \includegraphics[height=6cm]{diego-initial-letter}
     }
-    \subfigure[Final letter, after training with Vincent]{
+    \subfigure[Final text, after training with Vincent]{
         \includegraphics[height=6cm]{diego-final-letter}
@@
     \caption{\small (French) text generated by the robot, before and after a one
-        hour long interaction session with the child. As an example, the red box
-        highlights the changes on the word ``envoyer''.}
+        hour long interaction session with the child. The red box
+        highlights one instance of striking improvement of the robot's
+        handwriting legibility.}
 
@@
 
-Overall, Vincent provided 154 demonstrations to the robot, and he remained
-actively engaged over the four weeks. The story was well accepted by Vincent and
+Over the whole duration of the study, Vincent had provided 154 demonstrations to the robot, and he remained
+actively engaged over the four weeks. The story was well accepted by the child and
 he seriously engaged into the game. After the first week, he showed good
-confidence to play with the robot and he built affective bonds with the robot
-over the course of the study, as evidenced by some cries on the last session,
-and several letters sent to the robot \emph{after} the end of the study
-(one of them 4 months later) to get news. This represents a promising initial
-result: we can effectively keep a child engaged with the robot for a relatively
-long periods of time (about 5 hours).
+confidence to play with the robot and by the end of the study he had built
+affective bonds with the robot, as evidenced by several letters he did send to the
+robot \emph{after} the end of the study (one of them four months later) to get
+news. This represents an initial validation of our hypothesis: our system can effectively keep a
+child engaged with the robot for a relatively long periods of time (about 5
+hours spread over a month).
 
-No conclusion can be drawn in terms of actual handwriting remediation: we did
-not design this study to formally assess possible improvements. However, as pictured on Figure~\ref{fig:stimuli}, Vincent was able to
-significantly improve the robot's skill, and he acknowledged that he had been
-able to help the robot: in that regard, Vincent convinced himself that he was
-``good enough'' at writing to help someone else, which is likely to have
-a positive impact on his self-esteem.
+No hard conclusion can be drawn in terms of actual handwriting remediation: we did
+not design this study to formally assess possible improvements. However, as
+visible on Figure~\ref{fig:stimuli}, Vincent was able to significantly improve
+the robot's skill, and post-hoc interviews revealed that he acknowledged that he
+was the one helping the robot. In that respect, Vincent realized that he was
+``good enough'' at writing to help someone else. Quoting the mother a week
+after the end of the experiment: ``Vincent's handwriting has changed over the
+last weeks, going from a mix of standalone and cursive letters to full words in
+cursive. This requires a lot of efforts and concentration from him, but he did
+succeed during the sessions with the robot as he knew he had to show a
+consistent writing to the robot''.
 
@@
 
-Thomas is a 5.5 years old child, under the care of an occupational therapist. He
+Thomas is a 5.5 years old child, under the care of this occupational therapist. He
 has been diagnosed with visuo-constructive deficits, which translate into
 difficulties for him to consistently draw letters. Besides, focusing on a task
-is difficult for Thomas, who tends to rapidly shift his attention to something
-else. Since the robot's learning algorithm requires repeated demonstrations of
+is difficult for Thomas, who tends to rapidly shift his attention to other
+things. Since the robot's learning algorithm requires repeated demonstrations of
 similarly shaped letters to converge, the occupational therapist was especially
@@
 first session. Hence, we also tested during this case-study if the robot (and
-the Protégé effect) would induce a strong enough intrinsic motivation to
-keep the child engage over the five weeks.
+the Protégé effect) would induce by itself a strong enough intrinsic motivation to
+keep the child engaged over the five weeks.
 
@@
 Since Thomas would frequently draw mirrored numbers, or hard-to-recognize
-shapes, the learning algorithm of the robot tended to converge towards
+shapes, the learning algorithm of the robot initially tended to converge towards
 meaningless scribblings. We addressed this issue by having the robot to
-\emph{refuse} allographs that were too far from the reference shape (\emph{``I'm
+\emph{refuse} allographs that were too far from the reference shape (the robt
+would instead say \emph{``I'm
 not sure to understand what you are drawing...''}), so that the child had to
@@
 interpret and understand his writing. Figure~\ref{Thomas_progress} shows how
-he gradually improved his demonstrations for some numbers, according to the
+he gradually improved his demonstrations for two different numbers, according to the
 metric we used to make the robot accept/refuse samples.
@@
 time Thomas succeeded to have his demonstration accepted by it, the robot's
-improvement was clearly visible (as measured in Figure~\ref{Thomas_distances}).
+improvement was clearly visible (as shown in Figure~\ref{Thomas_distances}).
 This led to a self-rewarding situation that effectively supported Thomas'
@@
     }
-    \caption{\small Improvement of Thomas demonstrations for some numbers: a) drawing the allograph of the number 1, b) the number 2 and c) the number 5.
-Thomas progressively took care of the demonstrations he was providing to the robot for those numbers.
-We used for this figure the same metric than the one used for the acceptance algorithm.
-Distances are normalized with respect to the biggest value.
-The dashed line correspond to the threshold of robot's acceptance.}
+    \caption{\small Improvement of Thomas demonstrations for the numbers ``2''
+        and ``5''. Thomas progressively took care of the demonstrations he was providing to the
+    robot for those numbers. We used for this figure the same metric than the one
+    used for the acceptance algorithm. Distances are normalized with respect to the
+    biggest value. The dashed line correspond to the threshold of robot's
+    acceptance.}
 

### 2015-12-15 main.tex
Report on With-me-ness measurements

@@
 {\it School C}     & Pair interaction at school               & 26 min/group           & 7 $\times$ 2                          & 5-6        \\
-{\it School A (2)} & Individual interaction at school         & 18 min                 & 6                                     & 5-6        \\ \midrule
+{\it School A (2)} & Individual interaction at school         & 20 min                 & 6                                     & 5-6        \\ \midrule
 {\it Surgery}      & Individual interaction at surgery        & 3 sessions $\times$ 1h & 8                                     & 6-8        \\ \midrule 
@@
 reliability and allow a technically sound autonomous interaction. For instance,
-during the second school study (\textit{Florimont}), the robot withstood
+during the second school study (\textit{School B}), the robot withstood
 interactions which lasted for a total of 160 minutes.  During this time the
@@
 Factually, the children engaged into the teaching activity: in the
-\textit{Florimont} study, for instance, they demonstrated an average of
+\textit{School B} study, for instance, they demonstrated an average of
 10.9 demonstration letters (SD = 4.4) for an average session duration of 11 minutes.
@@
 
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=0.9\linewidth]{engagement}
-    \caption{\small Engagement level of one child, manually annotated by two judges
-        over a 13 min long interaction (Cronbach’s alpha test: $\alpha = .82$,
-        good inter-judge agreement). 0 means ``complete disengagement from the
-        task'', 6 means ``full engagement''. The vertical lines represent activity
-    switches between handwriting teaching and story-telling (used as a
-    distruber).}
-    \label{engagement_level}
-\end{figure}
-
-We also conducted a qualitative assessment of the engagement levels during
-several interaction sessions (studies \textit{Haut-Lac} and \textit{ISG 2}).
-While the overall engagement was high, Figure~\ref{engagement_level} shows an
-example of a child with moderate level of engagement. On-going research in our
-group aims at assessing such engagement levels in real-time, so that the robot
-can address possible disengagement (by changing/adapting its physical behaviour,
-by switching to different activities, etc.)
-
-We also performed further qualitative assessment of engagement over longer
-interaction periods: two 4-weeks long case studies with one child at a time have
-been conducted so far, the first one (\emph{Vincent}) focusing on
-\emph{engagement on the long run}, the second one (\emph{Thomas}) looking at
-\emph{acceptance and impact in a real therapeutic context}.
+%\begin{figure}[ht!]
+%    \centering
+%    \includegraphics[width=0.9\linewidth]{engagement}
+%    \caption{\small Engagement level of one child, manually annotated by two judges
+%        over a 13 min long interaction (Cronbach’s alpha test: $\alpha = .82$,
+%        good inter-judge agreement). 0 means ``complete disengagement from the
+%        task'', 6 means ``full engagement''. The vertical lines represent activity
+%    switches between handwriting teaching and story-telling (used as a
+%    disturber).}
+%    \label{engagement_level}
+%\end{figure}
+
+\begin{table}[h!]
+    \centering
+    \caption{\textbf{Levels of with-me-ness}. Percentage of interaction time
+        during which the child was effectively focusing his/her attention on the
+        task. The six children are those from the second study at School A.
+        Interaction duration: M = 19.6 min, SD = 1.58. Results
+        taken from~\cite{lemaignan2016realtime}.}
+
+    \begin{tabular}{p{1cm}cccccccc}
+        \toprule
+        Child & 1 & 2 & 3 & 4 & 5 & 6 & {\bf M} & {\it SD} \\
+        \midrule
+        $\mathcal{W}$ & 79.4\% & 81.6\%  & 90.5\% & 87.9\% & 90.7\% & 80.9\% & {\bf 85.2\%} & {\it 5.1} \\ 
+        \bottomrule
+    \end{tabular}
+    \label{tab:results-with-me-ness}
+\end{table}
+
+We also conducted a quantitative assessment of the engagement levels of the
+children. Table~\ref{tab:results-with-me-ness} reports the levels of
+\emph{with-me-ness} of the children during the second study at school A.
+\emph{With-me-ness} is a quantifiable precursor of engagement: it measures the
+percentage of time spent by the child focusing onto the task at end. It was
+first devised in the context of computer-supported
+learning~\cite{sharma2014me}, and we have previously studied its applicability to
+human-robot interaction and formalized the exact methodology
+in~\cite{lemaignan2016realtime}\footnote{Note that with-me-ness is a metric that
+allows real-time computation by the robot itself: while we did not yet make
+use of it, it would allow the robot to detect possible disengagements, and
+eventually address them by adapting its physical behaviour, switching to
+different activities, etc.}.
+
+The average with-me-ness is well above 80\%, and confirms that the children
+where very much engaged into these 20 minutes of interaction with the robot,
+paying close attention to the task.
+
+The other three experiments (the group study at the surgery and the two
+case-studies) provide further qualitative assessment of engagement over longer
+interaction periods. In particular, as reported hereafter, the two 4-weeks long
+case studies that have been conducted so far show that our system can sustain
+children' engagement over durations (5 hours) that are closer to what is
+expected to have an impact in a real therapeutic context.
 

### 2015-12-15 rebuttal.tex
Added a first version of the rebuttal

\documentclass{article}
\usepackage[utf8]{inputenc}

\begin{document}

Dear Editor, Dear Reviewers,

\vspace{2em}

We would like to thank both the reviewers and the editor for going through our
manuscript. As a consequence of your comments, the manuscript underwent
several significant changes, that we summarize hereafter.

While the reviews are generally positive (quoting one of the reviewer, ``this
is an excellent model for this kind of research, and the results are
promising''), we certainly take the two main concerns raised from the reviews:

\begin{itemize}
    \item The reporting of the studies is lacking
    \item Possible ethical concerns are not emphasized
\end{itemize}

This revision of our manuscript attempts to address both these concerns.

\section*{Reporting of the studies}

We have clarified our wording to only draw conclusion from factual measured
data or feedback from the parents/teachers/therapists (this is in particular
the case for the case-study 1),

We have added new quantitative data in the reporting of the case study 2, that
show a quantifiable improvement of handwriting. We want however to remain
cautious in our claims: handwriting acquisition is a complex cognitive skill.
While the child demonstrated clear improvements over the course of our
experiment, measuring the lasting effects of the robot-based remediation is
difficult (as many other individual and social factors may have impacted the
child's performance as well).

Finally, we have added the results of one more experiment (conducted during the
summer 2015), that involved 8 children with handwriting deficits. This
experiment focused on measuring how \emph{serious} the children were at
teaching the robot,

Other specific points were raised:

\begin{itemize}
    \item inclusions criteria should now be clearer,
    \item

\end{itemize}

\section*{Ethical concerns}

An explicit paragraph has been added to the discussion that states the possible
ethical concerns raised by this work. While the psychological impact of the
mentor-protégé relationship has been studied in the literature, it has been
done \emph{from the perspective of the protégé}. In our situation, the protégé
is the robot, and the child is the mentor: we could not find existing
literature discussing this situation, but would be nevertheless glad to further
expand the discussion if necessary.

\section*{Other comments}

One of the reviewers asked for clarification on a few other points:

\begin{itemize}

\item \emph{Did we include children who had difficulty with the dynamics of
turn-taking? If so, how did the system fare with those children?} In the case
the children did not pro-actively engaged into turn-taking, the facilitator (be
it one of the experimenter or the therapist) would suggest it.

\item \emph{The handwriting/letter generation is sufficiently described, but
how are the corresponding robot gestures generated?} We relied on NaoQI inverse
kinematics solver. This has been added to the text.

\item \emph{Was an interaction in which the child moves the hand of the NAO to
write the letter considered? Why did the authors choose the current interaction
model in which the child demonstrates on a tablet?} The Nao robot does not
feature force/torque sensing, which prevents most of the interesting direct
physical interaction like moving the hand of the robot. With appropriate
hardware, this interaction modality would be interesting to explore, though.

\item \emph{Could the measure of engagement plotted in Figure 7 be confounded
with the physically-active aspect of the hand-writing activity versus the
passive storytelling activity?} We have decided to replace this figure with an
objective metric (with-me-ness) that relies on the focus of attention of the
children to assess to which extend the child is into the task.

\item \emph{The paper would benefit from further explanation of the adult's
active role throughout the robot-child interaction} The manuscript hopefully
clarifies the role of the facilitator in the different studies.

\end{itemize}

\end{document}

### 2015-12-15 main.tex
Improved reporting on Thomas progress with simple satistical analysis

@@
     \includegraphics[width=0.9\linewidth]{learning_6_demos}
+
+    \vspace{1em}
+
+    \includegraphics[width=0.7\linewidth]{learning_6_distances}
     \caption{\small Demonstrations provided by Thomas for the number ``6'' (top row) and
-        corresponding shapes generated by the robot. After eight demonstrations,
-        Thomas decided that the robot's ``6'' was good enough, and went to
-    another character: in that respect, he was the one leading the learning
+        corresponding shapes generated by the robot. The plot beneath shows the distance
+        (in the eigenspace of the shape) between Thomas demonstrations, the
+        robot's attempts and our reference shape. After eight demonstrations,
+        Thomas decided that the robot's ``6'' was good enough, and switched to
+    another character. In that respect, he was the one leading the learning
 process of the robot.}
@@
 learning algorithm to start with a roughly vertical stroke instead of a
-deformed number (round 0 on Figure~\ref{learning_6_demos}).
+deformed number (see the initial state on Figure~\ref{learning_6_demos}).
 
@@
 time Thomas succeeded to have his demonstration accepted by it, the robot's
-improvement was clearly visible (as shown in Figure~\ref{Thomas_distances}).
+improvement was clearly visible (as shown in Figure~\ref{learning_6_demos}).
 This led to a self-rewarding situation that effectively supported Thomas'
 engagement.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\linewidth]{learning_6_distances}
-    \caption{\small Two metrics to assess the handwriting progresses: Euclidean
-    distance in the subspace of the number dataset (top figure) or in
-Cartesian space (bottom figure). Green lines represent the robot performance,
-blue lines performance of the child. The round IDs correspond to the demonstrations
-pictured on Figure~\ref{Thomas_demos}.}
-    \label{Thomas_distances}
-\end{figure}
 
@@
     }
-    \caption{\small Improvement of Thomas demonstrations for the numbers ``2''
-        and ``5''. Thomas progressively took care of the demonstrations he was providing to the
-    robot for those numbers. We used for this figure the same metric than the one
-    used for the acceptance algorithm. Distances are normalized with respect to the
-    biggest value. The dashed line correspond to the threshold of robot's
-    acceptance.}
+    \caption{\small Normalized distance between Thomas' demonstrations and
+        reference allographs for the numbers ``2'' and ``5''. The horizontal
+        dashed line correspond to the threshold for the robot to accept a
+        demonstration. Thomas' progress is visible on these figures: we find a
+        significant negative regression equation ($r=-0.023, F(1,19)=8.69,
+        p<.02$, adjusted $R^2=.461$) for the number ``2'' (dotted red line),
+        indicating that Thomas' shapes are getting closer to the reference. The
+        regression is not significant for the number ``5'', but we can observe
+        that after about 10 repetitions, all the demonstrations are deemed of
+        acceptable quality by the robot.  }
 

### 2015-12-15 main.tex
WIP on the conclusion

@@
 
+As we show through extensive experimental validation, these claims are not just
+words: we have effectively deployed robots on the field, with children suffering
+actual handwriting impairments. Our results are promising in terms of sustaining
+the interest and attention of children on otherwise boring writing exercises,
+and we evidence clear handwriting improvements.
+
+It is however still early to quantify the lasting effects of this remediation:
+handwriting is a complex cognitive skill, that builds on many individual and
+social factors. Self-confidence is one of them. We expect that our
+approach that endows the child with the role of a teacher who can help a robot,
+may help some children to recover self-esteem and self-confidence. However the
+experiments that we have conducted so do not allow us to confirm this
+hypothesis, and more research need to be conducted in this direction.
+
 \paragraph{Possible Ethical Concerns} Two aspects of this research ought to be

### 2015-12-15 main.tex
First lines of a real discussion

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+
+The four school studies, the experiment at the surgery and the two case-studies
+provide a first broad picture of how a robot-based remediation to handwriting
+is accepted by the children and practitioners, and what are the outcomes that
+can be expected.
+
+In total, more than 80 children have interacted with the system, for a total
+duration of more than 38 hours in multiple experimental configurations
+(individual interaction \vs pairs \vs groups; at school, in the lab, at a
+surgery; short interactions \vs long, repeated interactions; 5 to 8 years old
+children).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}

### 2015-12-15 main.tex
Added missing name in acknowledgmeents

@@
 of Geneva, the Institut Florimont and the Haut-Lac school for their
-participation and support. We express our great gratitude to Anne Gossin and [TBD] for
-inviting us to their surgery, for their involvement in the experiments and for
-bearing with our disruptive robots. And finally, we want to say a very special \emph{Thank
-you!} to Vincent, Thomas, Valérie, Émilien, Mathieu, Johan, Nicolas, Antoine,
-Adèle and Marie: by spending several hours being patient teachers for our
-robots, you have paved the way for what may become tomorrow a new way of
-learning how to deal with all these letters and words!
+participation and support. We express our great gratitude to Anne Gossin and
+Fanny Lebel for inviting us to their surgery, for their involvement in the
+experiments and for bearing with our disruptive robots. And finally, we want to
+say a very special \emph{Thank you!} to Vincent, Thomas, Valérie, Émilien,
+Mathieu, Johan, Nicolas, Antoine, Adèle and Marie: by spending several hours
+being patient teachers for our robots, you have paved the way for what may
+become tomorrow a new way of learning how to deal with all these letters and
+words!
 

### 2015-12-15 main.tex
WIP on the Normandy study

@@
 The other three experiments (the group study at the surgery and the two
-case-studies) provide further qualitative assessment of engagement over longer
+case-studies) provide further qualitative evidence of engagement over longer
 interaction periods. In particular, as reported hereafter, the two 4-weeks long
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Study: How Children Take on the Role of a Teacher}\label{normandie}
+\subsection{Surgery Study: How Children Take on the Role of a Teacher}\label{normandie}
 
@@
 
-\begin{inparaenum}
+\begin{inparaenum}[\itshape 1\upshape)]
 This experiment was conducted to study \item how easily children with actual
 deficit take on the role of a teacher, and \item if they engage \emph{seriously}
-in this role.
+in this role. As such, measuring actual handwriting improvements was not a primary goal
+of this experiment.
 \end{inparaenum}
@@
 The experiment took place at the surgery of an occupational therapist in
-Normandy, France. Eight children participated, all showing troubles with
-handwriting acquisition: Valérie (7 years old), Antoine (6.5) and Johan
-(7) are under the care of the occupational therapist. Émilien (8) and Mathieu (7)
-are repeating their school year because of writing difficulties. Mona (6) and
-Adèle (8) are both ranked at the bottom of their respective classes in writing
-activities. Nicolas (7) is under the care of a neurologist, and has been
-diagnosed with specific language impairment. Given their school level, all those
-children would be expected to know how to correctly shape cursive letters.
+Normandy, France. Eight children participated, selected by the occupational
+therapist based on their age and type of deficit (related to handwriting).
+Valérie (female, 7 years old), Antoine (male, 6.5) and Johan (male, 7) are under
+the care of the occupational therapist. Émilien (male, 8) and Mathieu (male, 7)
+are repeating their school year because of writing difficulties. Marie (female,
+6) and Adèle (female, 8) are both ranked at the bottom of their respective
+classes in writing activities. Nicolas (male, 7) is under the care of a
+neurologist, and has been diagnosed with specific language impairment. Given
+their school level, all those children would be expected to know how to
+correctly shape cursive letters.
 
 Over a period of two weeks, each child attended three time a one hour long
-session, except for Adele and Mona who only attended one session. The
-experimenter role was limited to the explaination of the rules of the game and
+session (except for Marie and Adèle who only attended one session). The
+experimenter role was limited to the explanation of the rules of the game and
 the basic tablet usage. For this experiment, the children were provided with two
-tablets: one to choose the word (or single letter) to teach to the robot, one
-used to write. We also provided templates of the cursive letters, where the
-children to ask for.
+tablets: one to choose the words (or letters) to teach to the robot, the other one
+to write, like in the previous studies. We also provided templates of the
+cursive letters, where the children to ask for it.
 
@@
 children would naturally take on the role of the teacher, and how seriously they
-engaged into helping the robot.
-
-This assessment was carried through two additional buttons on the tablet: a
-green ``thumbs up'' and a red ``thumbs down''. The children were told to use
+engage into helping the robot.
+
+This assessment is carried through two additional buttons on the tablet: a
+green ``thumbs up'' and a red ``thumbs down''. The children are told to use
 them to evaluate the robot's improvements (``thumb up'' being a way to give
-positive feedback while ``thumb down'' would be convey negative feedback). Our
-assessment builds on the hypothesis that the more the child provides feedback to
-the robot with the green/red buttons, the more he/she assumes the role of a
+positive feedback while ``thumb down'' would convey negative feedback). Our
+assessment builds on the hypothesis that the more the children provides feedback to
+the robot with the green/red buttons, the more they assume the role of a
 teacher. Then, by correlating the feedback with the actual performance of the
-robot, we can measure to what extend the child was seriously assuming his/her
-role: since the child knows what a correct letter should look like, if his/her
-feedback does not correlate with the actual performance, the child was likely
-engaged into a playful instead of serious role-playing.
+robot, we can measure to what extend the children are seriously assuming their
+role: since the child knows what a correct letter should look like, if their
+feedback does not correlate with the actual performance of the robot, the child
+is likely engaged into a \emph{playful} instead of \emph{serious} role-playing.
 
@@
 
-All children maintained their engagement during the whole sessions. They
-provided on average 42 demonstrations per session. All children used the
-feedback buttons and had preference to give positive feedback the robot (99
-``thumb up'' \vs 33 ``thumb down'' were recorded).
-
-As the children participated at most 3 hours of interactions spanning over two
-week, we did not attempted to measure possible handwriting improvements by the
-children, and instead we focused on the possible correlation between the
-children's feedback and the robot's progression.  We estimated the robot's
-progression as the difference between a starting score (score of the first
-robot's try when children have chosen a new word/letter to work on) and the
-current robot's score (after being taught by the child). The score is calculated
-as the average of euclidean distance between robot's try and a reference
-allograph over all letters of the word. Those references for letter allographs
-where drawn by us beforehand, taking inspiration in education.com cursive
-letters
-template\footnote{\url{http://www.education.com/slideshow/cursive-handwriting-z/}}.
+All children maintained their engagement during the entire sessions. They
+provided on average 42 demonstrations per session. All children did use the
+feedback buttons (in total, 99 ``thumb up'' and 33 ``thumb down'' were recorded,
+see Table~\ref{table:scores}).
+
+\begin{table}
+    \centering
+    \begin{tabular}{cccccc}
+        \toprule
+        Child          & \# demos & \# positive & \# negative & $p_{robot}$ & $p_{child}$\\ \midrule
+        \emph{Valérie} & 127 & 24 & 6 & $< .003$ \tt ** & $< .055$ \tt .\\ 
+        \emph{Émilien} & 223 & 20 & 9 & $.17$ \it ns & $.35$ \it ns \\
+        \emph{Mathieu} & 131 & 10 & 3 & $< .004$ \tt ** & $< .008$ \tt **\\
+        \emph{Johan}   & 98 & 10 & 5 &  $.15$ \it ns & $.38$ \it ns  \\
+        \emph{Nicolas} & 115 & 16 & 4 & $< .0006$ \tt *** & $< .003$ \tt ** \\
+        \emph{Antoine} & 83 & 10 & 3 & $< .031$ \tt * & $.6$ \it ns \\
+        \emph{Adèle}   & 35 & 4 & 2 & $< .05$ \tt * & $< .04$ \tt * \\
+        \emph{Marie}   & 40 & 5 & 1 &  $.54$ \it ns & $.2$ \it ns \\ \midrule
+        Total          & 852 & 99 & 33 &   & \\ \bottomrule
+    \end{tabular}
+    \caption{\small results of evaluations. demo: number of demonstrations provided by
+        the child over all session. rew: number of rewards accorded by
+        the child. pun: number of punishments. p(robot): are the evaluations
+        significantly corresponding to the progression of the robot ? p(child): are the
+        evaluations significantly corresponding to the child's own
+progression ?}
+    \label{table:scores}
+\end{table}
+
+To study a possible correlation between the children's feedback and the robot's
+improvements, estimated the robot's progress as the difference between an
+initial score (average euclidian distance between the shape drawn by the robot
+at its first attempt and the shape of the reference letter) and the robot's
+score at the current round of demonstrations.
+
 Let $P_i$ be the estimated progression of the robot at time $ i$. Of course, if
@@
 
-\begin{table}
-    \centering
-    \begin{tabular}{|c|c|c|c|c|c|}
-        \hline
-        child & demo & rew & pun & p(robot) & p(child)\\ \hline
-        Valérie & 127 & 24 & 6 & 2.4e-03 & 5.5e-02\\ \hline
-        Émilien & 223 & 20 & 9 & 1.7e-01 & 3.5e-01\\ \hline
-        Mathieu & 131 & 10 & 3 & 3.8e-03 & 7.9e-03\\ \hline
-        Johan & 98 & 10 & 5 &  1.5e-01 & 3.8e-01\\ \hline
-        Nicolas & 115 & 16 & 4 & 5.3e-04 & 2.7e-03\\ \hline
-        Antoine & 83 & 10 & 3 & 3.1e-02 & 6.0e-01\\ \hline
-        Adèle & 35 & 4 & 2 & 5.0e-02 & 3.7e-02\\ \hline
-        Marie & 40 & 5 & 1 &  5.4e-01 & 2.0e-01\\ \hline
-    \end{tabular}
-    \caption{\small results of evaluations. demo: number of demonstrations provided by
-        the child over all session. rew: number of rewards accorded by
-        the child. pun: number of punishments. p(robot): are the evaluations
-        significantly corresponding to the progression of the robot ? p(child): are the
-        evaluations significantly corresponding to the child's own
-progression ?}
-    \label{table:scores}
-\end{table}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 pay good attention to what he would demonstrate to the robot. Also, to make
-the robot's progresses evident, we modified the initialization step of the
+the robot's progress evident, we modified the initialization step of the
 learning algorithm to start with a roughly vertical stroke instead of a

### 2015-12-16 main.tex
WIP on the Normandy study

@@
 \newcommand{\vs}{\textit{vs.}\xspace}
+\newcommand{\resp}{\textit{resp.}\xspace}
 
@@
 
+\begin{table}
+    \centering
+    \begin{tabular}{ccccll}
+        \toprule
+        Child          & \#demos & \#positive & \#negative & $r_{robot}$ & \\ \midrule
+        \emph{Valérie} & 127 & 24 & 6 & 0.25 \small\tt **  \\ 
+        \emph{Émilien} & 223 & 20 & 9 & 0.06 \small\it ns  \\
+        \emph{Mathieu} & 131 & 10 & 3 & 0.23 \small\tt **  \\
+        \emph{Nicolas} & 115 & 16 & 4 & 0.31 \small\tt *** \\
+        \emph{Johan}   & 98 & 10 & 5  & 0.10 \small\it ns  \\
+        \emph{Antoine} & 83 & 10 & 3  & 0.20 \small\tt *   \\
+        \emph{Adèle}   & 35 & 4 & 2   & 0.28 \small\tt *   \\
+        \emph{Marie}   & 40 & 5 & 1   & -0.02 \small\it ns \\ \midrule
+        Total          & 852 & 99 & 33  &   & \\ \bottomrule
+    \end{tabular}
+    \caption{\small Feedback from the children to the robot. \emph{\#demos}
+        denotes the total number of demonstrations drawn by the children;
+        \emph{\#positive} and \emph{\#negative} the number of positive \resp
+        negative feedbacks they provided. $r_{robot}$ reflects the correlation
+        between the feedback provided by the children and the performance of the
+        robot. Results taken from~\cite{jacq2016building}.}
+
+    \label{table:scores}
+\end{table}
+
 All children maintained their engagement during the entire sessions. They
@@
 
-\begin{table}
-    \centering
-    \begin{tabular}{cccccc}
-        \toprule
-        Child          & \# demos & \# positive & \# negative & $p_{robot}$ & $p_{child}$\\ \midrule
-        \emph{Valérie} & 127 & 24 & 6 & $< .003$ \tt ** & $< .055$ \tt .\\ 
-        \emph{Émilien} & 223 & 20 & 9 & $.17$ \it ns & $.35$ \it ns \\
-        \emph{Mathieu} & 131 & 10 & 3 & $< .004$ \tt ** & $< .008$ \tt **\\
-        \emph{Johan}   & 98 & 10 & 5 &  $.15$ \it ns & $.38$ \it ns  \\
-        \emph{Nicolas} & 115 & 16 & 4 & $< .0006$ \tt *** & $< .003$ \tt ** \\
-        \emph{Antoine} & 83 & 10 & 3 & $< .031$ \tt * & $.6$ \it ns \\
-        \emph{Adèle}   & 35 & 4 & 2 & $< .05$ \tt * & $< .04$ \tt * \\
-        \emph{Marie}   & 40 & 5 & 1 &  $.54$ \it ns & $.2$ \it ns \\ \midrule
-        Total          & 852 & 99 & 33 &   & \\ \bottomrule
-    \end{tabular}
-    \caption{\small results of evaluations. demo: number of demonstrations provided by
-        the child over all session. rew: number of rewards accorded by
-        the child. pun: number of punishments. p(robot): are the evaluations
-        significantly corresponding to the progression of the robot ? p(child): are the
-        evaluations significantly corresponding to the child's own
-progression ?}
-    \label{table:scores}
-\end{table}
-
-To study a possible correlation between the children's feedback and the robot's
-improvements, estimated the robot's progress as the difference between an
+To study the correlation between the children's feedback and the robot's
+actual improvements, we estimate the robot's progress as the difference between an
 initial score (average euclidian distance between the shape drawn by the robot
 at its first attempt and the shape of the reference letter) and the robot's
-score at the current round of demonstrations.
-
-Let $P_i$ be the estimated progression of the robot at time $ i$. Of course, if
-the child chose to switch to a new word at time $ j$ , we have $ P_{j}=0$ and
-clearly $ P_0=0$.  To measure whether the fact that the child was rewarding the
-robot when it was progressing was significant or not, we generated 10000 times
-the same number of reward/punishment but accorded at random times. Let $ R_i^n$
-be the $n^{th}$ generated evaluation at time $ i$ ($ R_i^n = 0$ if no evaluation
-occurred at time $ i$ , $ R_i^n=1$ if a reward occurred at time $i$ and
-$R_i^n=-1$ if a punishment occurred at time $i$), and $\overline{R}_i$ be the
-actual evaluation at time $i$. For each $n^{th}$ generated sequence of
-evaluation, we compute a score of evaluation: $$ S^n = \sum\limits_i{R_i^n
-P_i}$$ Then we can estimate the p-value $p$ of the actual score: $$ \overline{S}
-= \sum\limits_i{\overline{R}_i P_i}$$ given the distribution of the generated
-scores $\left(S^n\right)_n$, which is assumed to be Gaussian:
-
-$$
-p(\overline{S})= \mathbb{P}{\left[X>\overline{S}\right]} = 1-\phi{\left(\frac{\overline{S}-\mu}{\sigma}\right)}
-$$
-
-where $\phi$ denotes the cumulative distribution function of the standard normal
-distribution, $\mu$ the mean and $\sigma$ the deviation of the generated scores
-$\left(S^n\right)_n$.  As a result, we found that 5 of the 8 children obtained a
-score of evaluation significantly hight ($p(\overline{S})<0.05$). Score of
-evaluation p-values of each child is reported in the second-to-last column of
-Table~\ref{table:scores}.
-
-We also studied correlation between children's evaluations and their own
-progression. The analysis was conducted in the same way, using distances between
-children demonstrations and reference allographs to compute children
-progressions.  3 of 5 children that played ``seriously" obtained score of
-evaluation of their own progression significantly high (last column of
-Table~\ref{table:scores}). For those children, it seems that the robot was
-reflecting their own performances, and while they were judging the robot
-positively (three times more rewards than punishments) they were actually
-evaluating themselves.
+score at the current round of demonstrations. We then correlate this progress
+to the positive or negative feedback provided by the children, following a
+statistical method presented in~\cite{jacq2016building}.
+
+We find (Table~\ref{table:scores}) that the feedback of five out of the eight
+children do significantly correlate with the actual performance of the robot:
+this indicates that these children are effectively taking on the teacher's
+clothes and are seriously providing feedback to the robot. \fixme{what can we
+say about the 3 others? they did not play seriously, or they were unable to
+actually recognize the robot mistakes?}
 

### 2015-12-17 main.tex
Large proof-reading

@@
 
-\author{Séverin Lemaignan$^1$, Alexis Jacq$^{1,2}$, Fernando Garcia$^1$,
-    Deanna Hood$^1$, \\Aude
-    Billard$^1$, Ana Paiva$^2$, Pierre Dillenbourg$^1$ \\
+\author{Séverin Lemaignan$^1$, Alexis Jacq$^{1,3}$, Deanna Hood$^1$, Fernando Garcia$^1$, \\
+    Aude Billard$^2$, Ana Paiva$^3$, Pierre Dillenbourg$^1$ \\
 $^1$CHILI Lab, École Polytechnique Fédérale de Lausanne, Suisse,\\
-$^2$Instituto Superior Técnico, University of Lisbon, Portugal}
+$^2$LASA, École Polytechnique Fédérale de Lausanne, Suisse,\\
+$^3$Instituto Superior Técnico, University of Lisbon, Portugal}
 
@@
 and struggles at school with his poor handwriting and even poorer
-self-confidence.\footnote{Children's names have been changed.}
+self-confidence.\footnote{All children's names have been changed.}
 
@@
 \emph{motor} deficits.  Improvements in self-confidence and anxiety occur (at
-best) as a side-effect of the child improving his/her handwriting skills and,
-consequently, improving his/her performance at school.
+best) as a side-effect of the child improving their handwriting skills and,
+consequently, improving their performance at school.
 
@@
 child. By building on the \emph{learning by teaching} paradigm, not only the
-child practises handwriting but, as (s)he takes on the role of the teacher,
-(s)he also positively reinforce his/her self-esteem and motivation: his/her
-social role shifts from the ``underperformer'' to ``the one who knows and
+child practises handwriting but, as they take on the role of the teacher,
+they also positively reinforce their self-esteem and motivation: their
+social role shift from the ``underperformer'' to ``the one who knows and
 teaches''. And by relying on a robot, we can tailor the exercises and the
-learning rate to each children' needs, as we show in this article.
-
-\subsection{Learning by Teaching}
+learning curve to each children' needs, as we will show in this article.
+
+\subsection{Learning by Teaching Handwriting}
 
@@
 specific social role. The child is the teacher in this relationship and the
-robot is the learner: the child must engage in a (meta-) cognitive relationship
+robot is the learner: the child is to engage in a (meta-) cognitive relationship
 with the robot to try to understand why the robot fails and how to help it best.
@@
 
-Besides the commitment of the child into the interaction build on a
+The commitment of the child into the interaction build on a
 psychological effect known as the
@@
 educational robotic. Most of these do not consider the agency induced by the
-robot beyond its motivational aspect: playing with an interactive, partially autonomous
+robot beyond its motivational aspect (playing with an interactive, partially autonomous
 device naturally induces some form of anthropomorphizing, which leads to some
-level of projected agency, and participate the overall excitement (at least, on
+level of projected agency, and participate the overall excitement -- at least, on
 the short-term, before the novelty effect vanishes).
-
 In our case, the role of agency is stronger: it induces meta-cognition (``I am
@@
 
-We then present and report on the field experiments that we conducted over the
+We then present and report on the field experiments that we have conducted over the
 last two years, including four studies at school, one longer experiment
@@
 one-month long case studies. While the focus of the school
-experiments was mostly technical validation and data acquisition, the three
+experiments was mostly the technical validation and data acquisition, the three
 other experiments involved children with actual deficits, and gave us initial insights on the
@@
 A tactile tablet (with a custom application) is used for both the robot and the
-child to write: during a typical round, the child requests the robot to write
+child to write. During a typical round, the child requests the robot to write
 something (a single letter, a number or a full word), and push the tablet
-towards the robot. The robot writes on the tablet by gesturing the writing -- the
-letters being actually drawn by the tablet application itself. The child then pull back the
+towards the robot. The robot writes on the tablet by gesturing the writing in
+the air,  the letters being actually drawn by the tablet application itself. The child then pull back the
 tablet, corrects the robot's attempt by writing him/herself on top or next to
@@
 small robot icon on the tablet once the child has finished correcting. In our
-experiments, both were easy to grasp for children.
+experiments, and once introduced by the experimenter, both were easy to grasp
+for all the children.
 
@@
 letters, with distortions that are actually plausible handwriting errors: they
-are actually exaggerations of variance of writing styles that naturally
+are exaggerations of variance of writing styles that naturally
 occurs amongst adult writers.  Figure~\ref{fig:sampleLetters} shows examples of
@@
     \item we can compute an euclidian distance to the reference letter to assess
-        the visual proximity of the demonstration with the expected letter, thus
+        the topological proximity of the demonstration with the expected letter, thus
         providing a quantitative metric of writing performance.
@@
 Figure~\ref{learning_6_demos} (p.~\pageref{learning_6_demos}) illustrates the
-process with the complete learning cycle of the number ``6''.
+process with a complete learning cycle of the number ``6''.
 
@@
 robot's handwriting gesture generation (based on the NaoQI inverse kinematics
-library) and high level control of the activity.
+library) and the high-level control of the activity (relying on {\sc
+pyRobots}~\cite{lemaignan2015pyrobots} and a custom finite state machine).
 
@@
 project is available online\footnote{The primary repository is
-\url{https://github.com/chili-epfl/cowriter_letter_learning}.}. Further details
-on the technical implementation are available in~\cite{hood2015when}.
+\url{https://github.com/chili-epfl/cowriter_letter_learning}.}. The details
+of the technical implementation are available in~\cite{hood2015when}.
 
@@
 The system has been deployed and tested in several situations: in three
-different schools (relatively short child-robot interactions with
-more than 70 children, aged 5 to 8), in the surgery of an occupational
-therapist (with 8 children, each interacting several hours with the system),
-and in two case studies that each lasted several weeks. Table~\ref{studies}
+different schools (more than 70 children aged 5 to 8; relatively short duration interactions), in the surgery of an occupational
+therapist (8 children; each interacting several hours with the system),
+and during two case studies that each lasted several weeks. Table~\ref{studies}
 gives an overview of these studies.
@@
 \toprule
-{\bf Study}        & {\bf Type}                               & {\bf Avg. duration}    & {\bf Children \#}                     & {\bf Ages} \\ \midrule
+{\bf Study}        & {\bf Type}                               & {\bf Avg. duration}    & {\bf \# Children}                     & {\bf Ages} \\ \midrule
 {\it School A (1)} & Unstructured group interaction at school & 16 min/group           & 4 $\times$ 8 children                 & 6-7        \\
@@
 {\it Surgery}      & Individual interaction at surgery        & 3 sessions $\times$ 1h & 8                                     & 6-8        \\ \midrule 
-{\it Vincent}      & Case-study, spanning over 4 weeks        & 4 weeks $\times$ 1.5h  & 1                                     & 6          \\
-{\it Thomas}       & Case-study, spanning over 4 weeks        & 4 weeks $\times$ 1h    & 1                                     & 5          \\ \bottomrule
+{\it Vincent}      & Case-study                               & 4 weeks $\times$ 1.5h  & 1                                     & 6          \\
+{\it Thomas}       & Case-study                               & 4 weeks $\times$ 1h    & 1                                     & 5          \\ \bottomrule
 \end{tabular}
@@
     \caption{\small The first field studies were focused on technical validation, with
-    more than 70 pupils interacting with the robot over short periods (around 15
-minutes), either alone or in small groups.}
+    more than 70 pupils interacting with the robot over short periods (between
+    10 and 25 minutes), either alone or in small groups.}
     \label{fig:schools}
@@
 not to select specifically underperforming children as having more children (73
-in total) was beneficial for these preliminary studies, and this would have
-besides required careful organization with the school due to ethical concerns.
+in total) was beneficial for these preliminary studies. Besides, due to ethical concerns, this would have
+required complex organization with the school that we wanted to avoid at the validation stage.
 
@@
 instances that the robot fell during that day.  Otherwise, the technical
-components of the system operated autonomously and as expected over the
-sessions.
-
-Due to the modular software architecture (about 10 ROS nodes), the occasional
-crashes occurring during others sessions were usually quickly resolved by
+components of the system operated autonomously and as expected with all the groups of children.
+
+Due to the modular software architecture (mostly independent ROS nodes), the occasional
+crashes occurring during others studies were usually quickly resolved by
 re-launching the faulty node alone, and did not significantly impact the
@@
 style (most notably, the ones requiring multiple strokes per letter) not being
-adequately processed by the learning algorithm. Support for such letters has
-been added as a follow-up of the studies.
+adequately processed by the learning algorithm. Support for such letters was
+added as a follow-up of the validation studies.
 
@@
 
-More generally, the approach was well accepted and recognized as useful and
-promising by teachers and parents; this was \textit{a posteriori} confirmed by
-the numerous spontaneous contacts made by parents and therapists who were
+During two focus groups organized with teachers, the welcomed the approach and
+recognized it as useful and
+promising; this was \textit{a posteriori} confirmed by
+multiple spontaneous contacts made by parents and therapists who were
 looking forward to use the system with their children.
@@
 
-As indicated in the introduction, the literature suggests that 20 handwriting
+The literature suggests that 20 handwriting
 practice sessions is found to be the minimum to demonstrate effective results in
@@
         \toprule
-        Child & 1 & 2 & 3 & 4 & 5 & 6 & {\bf M} & {\it SD} \\
+        \bf Child & 1 & 2 & 3 & 4 & 5 & 6 & {\bf M} & {\it SD} \\
         \midrule
@@
 \emph{With-me-ness} is a quantifiable precursor of engagement: it measures the
-percentage of time spent by the child focusing onto the task at end. It was
+percentage of time spent by the child focusing onto the task at end. This metric was
 first devised in the context of computer-supported
@@
 allows real-time computation by the robot itself: while we did not yet make
-use of it, it would allow the robot to detect possible disengagements, and
+use of it, it does in principle allow the robot to detect on-line possible disengagements, and
 eventually address them by adapting its physical behaviour, switching to
@@
 case studies that have been conducted so far show that our system can sustain
-children' engagement over durations (5 hours) that are closer to what is
+children' engagement over durations (5 hours) that are getting closer to what is
 expected to have an impact in a real therapeutic context.
@@
 This experiment was conducted to study \item how easily children with actual
-deficit take on the role of a teacher, and \item if they engage \emph{seriously}
-in this role. As such, measuring actual handwriting improvements was not a primary goal
+deficits take on the role of a teacher, and \item if they engage \emph{seriously}
+in this role. Measuring actual handwriting improvements was not a primary goal
 of this experiment.
@@
 Normandy, France. Eight children participated, selected by the occupational
-therapist based on their age and type of deficit (related to handwriting).
+therapist based on their age and type of deficit (all related to handwriting).
 Valérie (female, 7 years old), Antoine (male, 6.5) and Johan (male, 7) are under
-the care of the occupational therapist. Émilien (male, 8) and Mathieu (male, 7)
+the direct care of the occupational therapist. Émilien (male, 8) and Mathieu (male, 7)
 are repeating their school year because of writing difficulties. Marie (female,
@@
 neurologist, and has been diagnosed with specific language impairment. Given
-their school level, all those children would be expected to know how to
+their age and school year, all those children would be expected to know how to
 correctly shape cursive letters.
@@
 session (except for Marie and Adèle who only attended one session). The
-experimenter role was limited to the explanation of the rules of the game and
+experimenter role was limited to the explanation of the task and
 the basic tablet usage. For this experiment, the children were provided with two
 tablets: one to choose the words (or letters) to teach to the robot, the other one
-to write, like in the previous studies. We also provided templates of the
-cursive letters, where the children to ask for it.
+to write, like in the validation studies. We also provided paper-based templates of the
+cursive letters, where the children to ask for them.
 
@@
 have to help the robot to improve its writing style), so as to assess if and how
-children would naturally take on the role of the teacher, and how seriously they
-engage into helping the robot.
-
-This assessment is carried through two additional buttons on the tablet: a
-green ``thumbs up'' and a red ``thumbs down''. The children are told to use
-them to evaluate the robot's improvements (``thumb up'' being a way to give
-positive feedback while ``thumb down'' would convey negative feedback). Our
-assessment builds on the hypothesis that the more the children provides feedback to
-the robot with the green/red buttons, the more they assume the role of a
+children would naturally take on the role of the teacher. We besides assessed how seriously they
+engage into helping the robot through two additional buttons on the tablet: a
+green ``thumbs up'' and a red ``thumbs down''. The children were told to freely use
+them to evaluate the robot's improvements (``thumb up'' to give
+positive feedback, ``thumb down'' to convey negative feedback). Our
+assessment builds on the hypothesis that the more the child provides feedback to
+the robot, the more they assume the role of a
 teacher. Then, by correlating the feedback with the actual performance of the
 robot, we can measure to what extend the children are seriously assuming their
-role: since the child knows what a correct letter should look like, if their
-feedback does not correlate with the actual performance of the robot, the child
-is likely engaged into a \emph{playful} instead of \emph{serious} role-playing.
+role: if their feedback does correlate with the actual performance of the robot, the child
+is likely engaged into a \emph{serious} teaching role-playing.
 
@@
     \centering
-    \begin{tabular}{ccccll}
+    \begin{tabular}{ccccl}
         \toprule
-        Child          & \#demos & \#positive & \#negative & $r_{robot}$ & \\ \midrule
-        \emph{Valérie} & 127 & 24 & 6 & 0.25 \small\tt **  \\ 
-        \emph{Émilien} & 223 & 20 & 9 & 0.06 \small\it ns  \\
-        \emph{Mathieu} & 131 & 10 & 3 & 0.23 \small\tt **  \\
-        \emph{Nicolas} & 115 & 16 & 4 & 0.31 \small\tt *** \\
-        \emph{Johan}   & 98 & 10 & 5  & 0.10 \small\it ns  \\
-        \emph{Antoine} & 83 & 10 & 3  & 0.20 \small\tt *   \\
+        \bf Child      & \bf \# Demos/hour & \bf \# Positive & \bf \# Negative & $r$ \\ \midrule
+        \emph{Valérie} & 42           & 24              & 6               & 0.25 \small\tt **  \\ 
+        \emph{Émilien} & 74           & 20              & 9               & 0.06 \small\it ns  \\
+        \emph{Mathieu} & 43           & 10              & 3               & 0.23 \small\tt **  \\
+        \emph{Nicolas} & 38           & 16              & 4               & 0.31 \small\tt *** \\
+        \emph{Johan}   & 32           & 10              & 5               & 0.10 \small\it ns  \\
+        \emph{Antoine} & 27           & 10              & 3               & 0.20 \small\tt *   \\
         \emph{Adèle}   & 35           & 4               & 2               & 0.28 \small\tt *   \\
-        \emph{Marie}   & 40 & 5 & 1   & -0.02 \small\it ns \\ \midrule
-        Total          & 852 & 99 & 33  &   & \\ \bottomrule
+        \emph{Marie}   & 40           & 5               & 1               & -0.02 \small\it ns \\ \bottomrule
     \end{tabular}
-    \caption{\small Feedback from the children to the robot. \emph{\#demos}
-        denotes the total number of demonstrations drawn by the children;
-        \emph{\#positive} and \emph{\#negative} the number of positive \resp
-        negative feedbacks they provided. $r_{robot}$ reflects the correlation
+    \caption{\small Feedback from the children to the robot. \emph{\#Demos}
+        denotes the average number of demonstrations per hour provided by the children;
+        \emph{\#Positive} and \emph{\#Negative} the total number of positive \resp
+        negative feedbacks they provided. $r$ is the correlation coefficient
         between the feedback provided by the children and the performance of the
@@
 feedback buttons (in total, 99 ``thumb up'' and 33 ``thumb down'' were recorded,
-see Table~\ref{table:scores}).
-
-To study the correlation between the children's feedback and the robot's
-actual improvements, we estimate the robot's progress as the difference between an
+see Table~\ref{table:scores}). This indicates that they are all able to engage into playing
+the role of a teacher.
+
+To study the correlation between the children' feedback and the
+actual improvements of the robot, we estimate the robot's progress as the difference between an
 initial score (average euclidian distance between the shape drawn by the robot
@@
 score at the current round of demonstrations. We then correlate this progress
-to the positive or negative feedback provided by the children, following a
-statistical method presented in~\cite{jacq2016building}.
-
-We find (Table~\ref{table:scores}) that the feedback of five out of the eight
-children do significantly correlate with the actual performance of the robot:
-this indicates that these children are effectively taking on the teacher's
-clothes and are seriously providing feedback to the robot. \fixme{what can we
-say about the 3 others? they did not play seriously, or they were unable to
-actually recognize the robot mistakes?}
+to the positive or negative feedback provided by the children (details of the
+method are presented in~\cite{jacq2016building}).
+
+We find that the feedback of five out of the eight
+children do significantly correlate with the actual performance of the robot (Table~\ref{table:scores}).
+This indicates that these children are effectively taking on the teacher's
+clothes and are seriously providing feedback to the robot. The
+observation of the three remaining children reveal a variety of behaviours (for
+instance, one was actually rating how ``nice'' the robot is, and another one
+had rather ``artistic'' writing style preferences that were independent of the
+actual legibility of the robot), and during this specific experiment, we eventually
+found that they were all seriously engaged into teaching.
+
+To summarize, this experiment shows that children with actual handwriting impairments
+do accept well the robot, can commit themselves into long interactions, do take
+on the role of the teacher (as evidenced by both the number of writing demonstrations
+they provided and their self-inclination to give feedback to the robot), and
+finally take their teacher's role seriously.
 
@@
         whole word. Empty boxes are drawn on the screen to serve as template for the child
-        and to make word segmentation more robust.}
+        and to make letter segmentation more robust.}
     \label{fig:diego}
@@
 
-To complement the intrinsic motivation of helping a robot to communicate with another one, we
+To supplement the intrinsic motivation of helping a robot to communicate with another one, we
 gradually increased the complexity of Vincent's task to keep it challenging and
@@
 
-    \caption{\small (French) text generated by the robot, before and after a one
+    \caption{\small Text (in French) generated by the robot, before and after a one
         hour long interaction session with the child. The red box
@@
 news. This represents an initial validation of our hypothesis: our system can effectively keep a
-child engaged with the robot for a relatively long periods of time (about 5
-hours spread over a month).
-
-No hard conclusion can be drawn in terms of actual handwriting remediation: we did
+child engaged with the robot for a relatively long period of time (about 5
+hours spread over a month), and we can build a tutor/protégé relationship.
+
+No hard conclusion can be drawn in terms of actual handwriting remediation as we did
 not design this study to formally assess possible improvements. However, as
 visible on Figure~\ref{fig:stimuli}, Vincent was able to significantly improve
-the robot's skill, and post-hoc interviews revealed that he acknowledged that he
-was the one helping the robot. In that respect, Vincent realized that he was
-``good enough'' at writing to help someone else. Quoting the mother a week
-after the end of the experiment: ``Vincent's handwriting has changed over the
+the robot's skill, and  he acknowledged that he
+was the one helping the robot during post-hoc interviews. In that respect, Vincent realized that he was
+``good enough'' at writing to help someone else. The fact that Vincent seriously
+committed to the role of a teacher is further supported by feedback sent by
+Vincent's parents a week after the end of the experiment: ``Vincent's handwriting has changed over the
 last weeks, going from a mix of standalone and cursive letters to full words in
@@
 
-The second long-term study was designed with the help of an occupational therapist in
-Geneva and aimed at deploying the system for the first time on a real
+The second long-term study was designed in collaboration with an occupational therapist in
+Geneva, and aimed at deploying the system on the longer-term with a real
 therapeutic case.
 
-Thomas is a 5.5 years old child, under the care of this occupational therapist. He
+Thomas is a 5.5 years old child. He
 has been diagnosed with visuo-constructive deficits, which translate into
@@
 interested in observing if the robot would induce a strong enough motivation for
-Thomas to focus on producing regular, consistent letters, thus overcoming his
+Thomas to focus on producing many regular, consistent letters, thus overcoming his
 deficit.
 
-The experiment took place at the therapist's surgery (four sessions spanning
+The experiment took place at the therapist's surgery (four sessions spreading
 over 5 weeks). Contrary to Vincent's experiment, we chose not to introduce any
-storyline beyond a simple prompt (``the robot wants to participate to a robotic
-handwriting contest, will you help him to prepare?''), only provided during the
+backstory beyond a simple prompt (``the robot wants to participate to a robotic
+handwriting contest, will you help him to prepare?'') only provided during the
 first session. Hence, we also tested during this case-study if the robot (and
@@
     \caption{\small Demonstrations provided by Thomas for the number ``6'' (top row) and
-        corresponding shapes generated by the robot. The plot beneath shows the distance
-        (in the eigenspace of the shape) between Thomas demonstrations, the
-        robot's attempts and our reference shape. After eight demonstrations,
+        corresponding shapes generated by the robot. The plot beneath shows the
+        distance to the reference shape (in the eigenspace of the shape) to
+        Thomas' demonstrations, and to the
+        robot's attempts. After eight demonstrations,
         Thomas decided that the robot's ``6'' was good enough, and switched to
@@
 Since Thomas would frequently draw mirrored numbers, or hard-to-recognize
-shapes, the learning algorithm of the robot initially tended to converge towards
+shapes (see Figure~\ref{Thomas_progress}), the learning algorithm of the robot initially tended to converge towards
 meaningless scribblings. We addressed this issue by having the robot to
-\emph{refuse} allographs that were too far from the reference shape (the robt
+\emph{refuse} allographs that were too far from the reference shape (the robot
 would instead say \emph{``I'm
@@
 Despite his attention deficit, Thomas was able to remain engaged in the activity during more than
-forty minutes in each session. In total, 55 allographs out of 82 
+forty minutes in each session (a long time for a five years old). In total, 55 allographs out of 82 
 demonstrated by the child were acceptable considering our threshold (with a
@@
 clear their drawing and try again before sending it). According to
-the therapist, it was the first time that Thomas would correct himself in such a
+the therapist, it was the first time that Thomas was seen to correct himself in such a
 way, explicitly having to reflect on how \emph{another agent} (the robot) would
 interpret and understand his writing. Figure~\ref{Thomas_progress} shows how
-he gradually improved his demonstrations for two different numbers, according to the
-metric we used to make the robot accept/refuse samples.
+he gradually improved his demonstrations for two different numbers.
 
 Since the robot's handwriting started from a simple primitive (a stroke), each
-time Thomas succeeded to have his demonstration accepted by it, the robot's
+time Thomas succeeded in having his demonstration accepted by it, the robot's
 improvement was clearly visible (as shown in Figure~\ref{learning_6_demos}).
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
+\section{Summary of the Findings}
 
@@
 \begin{itemize}
-    \item robots in an educative context are \textbf{relevant and effective
+    \item robots in an educative context are certainly \textbf{relevant and effective
         beyond STEM} (Science, Technology, Engineering and Mathematics)
@@
     \item this proves essential to \textbf{sustain a long-term interaction} (several
-        hours) around a fundamentally routine -- yet challenging -- educational
-        task (handwriting learning),
+        hours) although the task would be typically considered as repetitive yet
+        challenging by the children,
 
     \item besides, the social commitment induces a \textbf{cognitive engagement}
-        of the child with the robot which is \textbf{a key learning lever} as
+        of the child with the robot, which is \textbf{a key learning lever} as
         its elicits \textbf{reflective, meta-cognitive mechanisms on the
@@
 the interest and attention of children on otherwise boring writing exercises,
-and we evidence clear handwriting improvements.
+and we evidence handwriting improvements.
 
@@
 handwriting is a complex cognitive skill, that builds on many individual and
-social factors. Self-confidence is one of them. We expect that our
-approach that endows the child with the role of a teacher who can help a robot,
-may help some children to recover self-esteem and self-confidence. However the
-experiments that we have conducted so do not allow us to confirm this
+social factors. Self-confidence is one of them. Our
+approach that endows the child with the role of a teacher who can help a robot:
+we expect it may help some children to recover self-esteem and self-confidence. However the
+experiments that we have conducted so far do not allow us to confirm this
 hypothesis, and more research need to be conducted in this direction.
@@
 \paragraph{Possible Ethical Concerns} Two aspects of this research ought to be
-discussed in term of their possible ethical implications: the perceived role of
-the robot \textit{vis-à-vis} the teacher, and the implications of the
+discussed in terms of their possible ethical implications: the perceived role of
+the robot \textit{vis-à-vis} the real teachers, and the implications of the
 mentor-protégé relationship for children, especially vulnerable ones.
@@
 alone (or even semi-alone). We can therefore reassure teachers: robots are not
-going to replace them any time soon.
+going to replace them any time soon, and the initial feedback that we received
+during the focus groups shows that, once explained, our approach
+make sense and is indeed welcomed by the teachers themselves.
 
 The implication of the mentor-protégé relationship on the children is less
-clearly understood. The children can establish strong affective bonds with the robot (as
-witnessed by the letter sent by Vincent several months after he interacted with
+clearly understood. We have certainly seen that the children can establish strong affective bonds with the robot (as
+witnessed for instance by the letter sent by Vincent several months after he interacted with
 the robot), but we are not yet able to precisely characterize these bonds. The
-ethical implication of the mentor-protégé relationship have been explored before
+ethical implications of the mentor-protégé relationship have been explored before
 in the context of human teaching~\cite{brad1999mentor,wendelyn2008context}, but
-they mostly look at the question from the perspective of the protégé, whereas in our
+they mostly looked at the question from the perspective of the protégé, whereas in our
 case, the child is the mentor. As such, relatively little is known on the
@@
 but its potential to positively impact the child's motivation and self-esteem.
-We hope that this article shows that this is a very relevant context of use for
+While we need to carefully clear up the possible ethical concerns, we hope that this article shows that this is a very relevant context of use for
 robots: when facing a child with school difficulties, robots can play the role
-of a naïve learner which neither adults nor peers -- because of the social
-effects it would induce -- can convincingly play.
+of a naïve learner which neither adults nor peers can convincingly play.
 
@@
 of Geneva, the Institut Florimont and the Haut-Lac school for their
-participation and support. We express our great gratitude to Anne Gossin and
+participation and support. We express our gratitude to Anne Gossin and
 Fanny Lebel for inviting us to their surgery, for their involvement in the

### 2015-12-17 main.tex
[minor] Lines re-wrapping

@@
 facing however the same difficulty to write in a legible manner. And, hidden
-beyond this impaired skill, psycho-social difficulties arise: they
-underperform at school, Thomas has to go for follow-up visits every week,
-they both live under the label ``requires special care''. This is a source of
-anxiety for the children and for their parents alike.
+beyond this impaired skill, psycho-social difficulties arise: they underperform
+at school, Thomas has to go for follow-up visits every week, they both live
+under the label ``requires special care''. This is a source of anxiety for the
+children and for their parents alike.
 
@@
 child. By building on the \emph{learning by teaching} paradigm, not only the
-child practises handwriting but, as they take on the role of the teacher,
-they also positively reinforce their self-esteem and motivation: their
-social role shift from the ``underperformer'' to ``the one who knows and
-teaches''. And by relying on a robot, we can tailor the exercises and the
-learning curve to each children' needs, as we will show in this article.
+child practises handwriting but, as they take on the role of the teacher, they
+also positively reinforce their self-esteem and motivation: their social role
+shift from the ``underperformer'' to ``the one who knows and teaches''. And by
+relying on a robot, we can tailor the exercises and the learning curve to each
+children' needs, as we will show in this article.
 
@@
 
-We also investigate here a particular role for a robot in the
-education of handwriting: not only is the robot actively performing the activity
-by drawing letters, but it does so in a way that engages the child in a very
-specific social role. The child is the teacher in this relationship and the
-robot is the learner: the child is to engage in a (meta-) cognitive relationship
-with the robot to try to understand why the robot fails and how to help it best.
-Here, the robot is more than just an activity facilitator or orchestrator -- its
+We also investigate here a particular role for a robot in the education of
+handwriting: not only is the robot actively performing the activity by drawing
+letters, but it does so in a way that engages the child in a very specific
+social role. The child is the teacher in this relationship and the robot is the
+learner: the child is to engage in a (meta-) cognitive relationship with the
+robot to try to understand why the robot fails and how to help it best.  Here,
+the robot is more than just an activity facilitator or orchestrator -- its
 physical presence and embodiment induce agency and anthropomorphising, and
@@
 
-The commitment of the child into the interaction build on a
-psychological effect known as the
-``protégé effect''~\cite{Chase2009}: the teacher feels responsible for his student, commits
-to the student's success and possibly experiences student's failure as his own
-failure to teach. Teachable computer-based agents have previously been used to
-encourage this ``protégé effect'', wherein students invest more effort into
-learning when it is for a teachable agent than for themselves~\cite{Chase2009}.
-We rely on this cognitive mechanism to reinforce the child's commitment into the
-robot-mediated handwriting activity, and we indeed show sustained child-robot
-engagement over extended periods of time (several hours spread of a month).
-
-For these two reasons, our approach is to be distinguished from previous works in
-educational robotic. Most of these do not consider the agency induced by the
-robot beyond its motivational aspect (playing with an interactive, partially autonomous
-device naturally induces some form of anthropomorphizing, which leads to some
-level of projected agency, and participate the overall excitement -- at least, on
-the short-term, before the novelty effect vanishes).
-In our case, the role of agency is stronger: it induces meta-cognition (``I am
-interacting with an agent, so I need to reflect on how to best teach him'')
-which is beneficial for the learning process; it also induces a protégé
-effect (``I want my robot-agent to succeed!'') which support the commitment of
-child into the interaction, also for longer period of time.
+The commitment of the child into the interaction build on a psychological effect
+known as the ``protégé effect''~\cite{Chase2009}: the teacher feels responsible
+for his student, commits to the student's success and possibly experiences
+student's failure as his own failure to teach. Teachable computer-based agents
+have previously been used to encourage this ``protégé effect'', wherein students
+invest more effort into learning when it is for a teachable agent than for
+themselves~\cite{Chase2009}.  We rely on this cognitive mechanism to reinforce
+the child's commitment into the robot-mediated handwriting activity, and we
+indeed show sustained child-robot engagement over extended periods of time
+(several hours spread of a month).
+
+For these two reasons, our approach is to be distinguished from previous works
+in educational robotic. Most of these do not consider the agency induced by the
+robot beyond its motivational aspect (playing with an interactive, partially
+autonomous device naturally induces some form of anthropomorphizing, which leads
+to some level of projected agency, and participate the overall excitement -- at
+least, on the short-term, before the novelty effect vanishes).  In our case, the
+role of agency is stronger: it induces meta-cognition (``I am interacting with
+an agent, so I need to reflect on how to best teach him'') which is beneficial
+for the learning process; it also induces a protégé effect (``I want my
+robot-agent to succeed!'') which support the commitment of child into the
+interaction, also for longer period of time.
 
@@
 
-The following sections of the paper go into details. We first provides a
-brief overview of the robotic system and the interaction it induces. We also
-present the machine-learning techniques that allow the robot to learn
-from the children, followed by the implementation on a {\sc nao} robot.
-
-We then present and report on the field experiments that we have conducted over the
-last two years, including four studies at school, one longer experiment
-with eight children at the surgery of an occupational therapist, and two
-one-month long case studies. While the focus of the school
-experiments was mostly the technical validation and data acquisition, the three
-other experiments involved children with actual deficits, and gave us initial insights on the
-relevance and effectiveness of our approach.
+The following sections of the paper go into details. We first provides a brief
+overview of the robotic system and the interaction it induces. We also present
+the machine-learning techniques that allow the robot to learn from the children,
+followed by the implementation on a {\sc nao} robot.
+
+We then present and report on the field experiments that we have conducted over
+the last two years, including four studies at school, one longer experiment with
+eight children at the surgery of an occupational therapist, and two one-month
+long case studies. While the focus of the school experiments was mostly the
+technical validation and data acquisition, the three other experiments involved
+children with actual deficits, and gave us initial insights on the relevance and
+effectiveness of our approach.
 
@@
 towards the robot. The robot writes on the tablet by gesturing the writing in
-the air,  the letters being actually drawn by the tablet application itself. The child then pull back the
-tablet, corrects the robot's attempt by writing him/herself on top or next to
-the robot's writing (see Figure~\ref{fig:diego}), and ``send'' his/her
-demonstration to the robot by pressing a small button on the tablet. The robot
-``learns'' from this demonstration and tries again.
-
-Since the children are assumed to take on the role of the teachers, we had to ensure
-they would be able to manage by themselves the turn-taking and the overall
-progression of the activity (moving forwards to the next letter or word). In our design,
-the turn-taking relies on the robot prompting for feedback once it is done with
-its writing (through simple sentences like ``What do you think?''), and pressing on a
-small robot icon on the tablet once the child has finished correcting. In our
-experiments, and once introduced by the experimenter, both were easy to grasp
-for all the children.
+the air,  the letters being actually drawn by the tablet application itself. The
+child then pull back the tablet, corrects the robot's attempt by writing
+him/herself on top or next to the robot's writing (see Figure~\ref{fig:diego}),
+and ``send'' his/her demonstration to the robot by pressing a small button on
+the tablet. The robot ``learns'' from this demonstration and tries again.
+
+Since the children are assumed to take on the role of the teachers, we had to
+ensure they would be able to manage by themselves the turn-taking and the
+overall progression of the activity (moving forwards to the next letter or
+word). In our design, the turn-taking relies on the robot prompting for feedback
+once it is done with its writing (through simple sentences like ``What do you
+think?''), and pressing on a small robot icon on the tablet once the child has
+finished correcting. In our experiments, and once introduced by the
+experimenter, both were easy to grasp for all the children.
 
@@
 Implementing such a system raises several challenges: first, the acquisition,
-analyse and learning from hand-written demonstration, which lays at the core of the
-our approach, necessitates the development of several algorithms for the robot to generate
-initial bad writing and to respond in an adequate manner, showing visible (but
-not too quick) writing improvements.
+analyse and learning from hand-written demonstration, which lays at the core of
+the our approach, necessitates the development of several algorithms for the
+robot to generate initial bad writing and to respond in an adequate manner,
+showing visible (but not too quick) writing improvements.
 
 Then, the actual implementation on the robot requires the coordination of
-several modules (from performing gestures and acquiring the user's input to
-the state machine implementing the high-level behaviour), spread over several devices (the robot itself,
-one laptop and up to four tactile tablets for some of the studies that we conducted). We
-relied on ROS to ensure the synchronization and communication between these
-modules.
+several modules (from performing gestures and acquiring the user's input to the
+state machine implementing the high-level behaviour), spread over several
+devices (the robot itself, one laptop and up to four tactile tablets for some of
+the studies that we conducted). We relied on ROS to ensure the synchronization
+and communication between these modules.
 
@@
 letters, with distortions that are actually plausible handwriting errors: they
-are exaggerations of variance of writing styles that naturally
-occurs amongst adult writers.  Figure~\ref{fig:sampleLetters} shows examples of
-deformed ``g'' generated with such a technique.
+are exaggerations of variance of writing styles that naturally occurs amongst
+adult writers.  Figure~\ref{fig:sampleLetters} shows examples of deformed ``g''
+generated with such a technique.
 
@@
 \emph{assess} their quality and \emph{learn} from them. Figure~\ref{fig:h} shows
-for example nine allographs of ``h'' written by a 6 years old child, along with the
-mean letter from the dataset (reference letter). By projecting each of the
+for example nine allographs of ``h'' written by a 6 years old child, along with
+the mean letter from the dataset (reference letter). By projecting each of the
 demonstrations onto the eigenspace of ``h'' (Figure~\ref{fig:h}b), we observe
@@
         mean-shift algorithm) by their visual styles,
+
     \item we can compute an euclidian distance to the reference letter to assess
-        the topological proximity of the demonstration with the expected letter, thus
-        providing a quantitative metric of writing performance.
+        the topological proximity of the demonstration with the expected letter,
+        thus providing a quantitative metric of writing performance.
+
 \end{itemize}
@@
 both a ROS API\footnote{The ROS documentation for {\sc nao} is available at
-\url{http://wiki.ros.org/nao}.} and the Aldebaran-provided NaoQI API, one
-to four Android tablets (the main tablet is used to draw the robot's letter and
-to acquire the children's demonstrations; more tablets have been used in some
+\url{http://wiki.ros.org/nao}.} and the Aldebaran-provided NaoQI API, one to
+four Android tablets (the main tablet is used to draw the robot's letter and to
+acquire the children's demonstrations; more tablets have been used in some
 studies, either to let the child input words to be written, or for the
@@
 project is available online\footnote{The primary repository is
-\url{https://github.com/chili-epfl/cowriter_letter_learning}.}. The details
-of the technical implementation are available in~\cite{hood2015when}.
+\url{https://github.com/chili-epfl/cowriter_letter_learning}.}. The details of
+the technical implementation are available in~\cite{hood2015when}.
 
@@
 The system has been deployed and tested in several situations: in three
-different schools (more than 70 children aged 5 to 8; relatively short duration interactions), in the surgery of an occupational
-therapist (8 children; each interacting several hours with the system),
-and during two case studies that each lasted several weeks. Table~\ref{studies}
-gives an overview of these studies.
+different schools (more than 70 children aged 5 to 8; relatively short duration
+interactions), in the surgery of an occupational therapist (8 children; each
+interacting several hours with the system), and during two case studies that
+each lasted several weeks. Table~\ref{studies} gives an overview of these
+studies.
 
@@
 not to select specifically underperforming children as having more children (73
-in total) was beneficial for these preliminary studies. Besides, due to ethical concerns, this would have
-required complex organization with the school that we wanted to avoid at the validation stage.
+in total) was beneficial for these preliminary studies. Besides, due to ethical
+concerns, this would have required complex organization with the school that we
+wanted to avoid at the validation stage.
 
@@
 instances that the robot fell during that day.  Otherwise, the technical
-components of the system operated autonomously and as expected with all the groups of children.
-
-Due to the modular software architecture (mostly independent ROS nodes), the occasional
-crashes occurring during others studies were usually quickly resolved by
-re-launching the faulty node alone, and did not significantly impact the
+components of the system operated autonomously and as expected with all the
+groups of children.
+
+Due to the modular software architecture (mostly independent ROS nodes), the
+occasional crashes occurring during others studies were usually quickly resolved
+by re-launching the faulty node alone, and did not significantly impact the
 interaction.
@@
 During two focus groups organized with teachers, the welcomed the approach and
-recognized it as useful and
-promising; this was \textit{a posteriori} confirmed by
-multiple spontaneous contacts made by parents and therapists who were
-looking forward to use the system with their children.
+recognized it as useful and promising; this was \textit{a posteriori} confirmed
+by multiple spontaneous contacts made by parents and therapists who were looking
+forward to use the system with their children.
 
@@
 
-The literature suggests that 20 handwriting
-practice sessions is found to be the minimum to demonstrate effective results in
-handwriting remediation~\cite{Hoy2011}. This highlights the necessity to sustain
-a child-robot engagement over the long-term if we want to achieve measurable learning gains.
+The literature suggests that 20 handwriting practice sessions is found to be the
+minimum to demonstrate effective results in handwriting
+remediation~\cite{Hoy2011}. This highlights the necessity to sustain a
+child-robot engagement over the long-term if we want to achieve measurable
+learning gains.
 
 Factually, the children engaged into the teaching activity: in the
-\textit{School B} study, for instance, they demonstrated an average of
-10.9 demonstration letters (SD = 4.4) for an average session duration of 11 minutes.
-In 9 out of the 14 sessions (64\%), the robot
-received demonstration letters even \emph{after} reaching the final stage of the
-interaction, suggesting an intrinsic motivation to further engage in the
-interaction.
+\textit{School B} study, for instance, they demonstrated an average of 10.9
+demonstration letters (SD = 4.4) for an average session duration of 11 minutes.
+In 9 out of the 14 sessions (64\%), the robot received demonstration letters
+even \emph{after} reaching the final stage of the interaction, suggesting an
+intrinsic motivation to further engage in the interaction.
 
@@
 \emph{With-me-ness} is a quantifiable precursor of engagement: it measures the
-percentage of time spent by the child focusing onto the task at end. This metric was
-first devised in the context of computer-supported
-learning~\cite{sharma2014me}, and we have previously studied its applicability to
-human-robot interaction and formalized the exact methodology
+percentage of time spent by the child focusing onto the task at end. This metric
+was first devised in the context of computer-supported
+learning~\cite{sharma2014me}, and we have previously studied its applicability
+to human-robot interaction and formalized the exact methodology
 in~\cite{lemaignan2016realtime}\footnote{Note that with-me-ness is a metric that
 allows real-time computation by the robot itself: while we did not yet make
-use of it, it does in principle allow the robot to detect on-line possible disengagements, and
-eventually address them by adapting its physical behaviour, switching to
-different activities, etc.}.
+use of it, it does in principle allow the robot to detect on-line possible
+disengagements, and eventually address them by adapting its physical
+behaviour, switching to different activities, etc.}.
 
@@
 This experiment was conducted to study \item how easily children with actual
-deficits take on the role of a teacher, and \item if they engage \emph{seriously}
-in this role. Measuring actual handwriting improvements was not a primary goal
-of this experiment.
+deficits take on the role of a teacher, and \item if they engage
+\emph{seriously} in this role. Measuring actual handwriting improvements was
+not a primary goal of this experiment.
 \end{inparaenum}
@@
 Valérie (female, 7 years old), Antoine (male, 6.5) and Johan (male, 7) are under
-the direct care of the occupational therapist. Émilien (male, 8) and Mathieu (male, 7)
-are repeating their school year because of writing difficulties. Marie (female,
-6) and Adèle (female, 8) are both ranked at the bottom of their respective
-classes in writing activities. Nicolas (male, 7) is under the care of a
-neurologist, and has been diagnosed with specific language impairment. Given
+the direct care of the occupational therapist. Émilien (male, 8) and Mathieu
+(male, 7) are repeating their school year because of writing difficulties. Marie
+(female, 6) and Adèle (female, 8) are both ranked at the bottom of their
+respective classes in writing activities. Nicolas (male, 7) is under the care of
+a neurologist, and has been diagnosed with specific language impairment. Given
 their age and school year, all those children would be expected to know how to
@@
 session (except for Marie and Adèle who only attended one session). The
-experimenter role was limited to the explanation of the task and
-the basic tablet usage. For this experiment, the children were provided with two
-tablets: one to choose the words (or letters) to teach to the robot, the other one
-to write, like in the validation studies. We also provided paper-based templates of the
-cursive letters, where the children to ask for them.
+experimenter role was limited to the explanation of the task and the basic
+tablet usage. For this experiment, the children were provided with two tablets:
+one to choose the words (or letters) to teach to the robot, the other one to
+write, like in the validation studies. We also provided paper-based templates of
+the cursive letters, where the children to ask for them.
 
@@
 have to help the robot to improve its writing style), so as to assess if and how
-children would naturally take on the role of the teacher. We besides assessed how seriously they
-engage into helping the robot through two additional buttons on the tablet: a
-green ``thumbs up'' and a red ``thumbs down''. The children were told to freely use
-them to evaluate the robot's improvements (``thumb up'' to give
-positive feedback, ``thumb down'' to convey negative feedback). Our
+children would naturally take on the role of the teacher. We besides assessed
+how seriously they engage into helping the robot through two additional buttons
+on the tablet: a green ``thumbs up'' and a red ``thumbs down''. The children
+were told to freely use them to evaluate the robot's improvements (``thumb up''
+to give positive feedback, ``thumb down'' to convey negative feedback). Our
 assessment builds on the hypothesis that the more the child provides feedback to
-the robot, the more they assume the role of a
-teacher. Then, by correlating the feedback with the actual performance of the
-robot, we can measure to what extend the children are seriously assuming their
-role: if their feedback does correlate with the actual performance of the robot, the child
-is likely engaged into a \emph{serious} teaching role-playing.
+the robot, the more they assume the role of a teacher. Then, by correlating the
+feedback with the actual performance of the robot, we can measure to what extend
+the children are seriously assuming their role: if their feedback does correlate
+with the actual performance of the robot, the child is likely engaged into a
+\emph{serious} teaching role-playing.
 
@@
 feedback buttons (in total, 99 ``thumb up'' and 33 ``thumb down'' were recorded,
-see Table~\ref{table:scores}). This indicates that they are all able to engage into playing
-the role of a teacher.
-
-To study the correlation between the children' feedback and the
-actual improvements of the robot, we estimate the robot's progress as the difference between an
-initial score (average euclidian distance between the shape drawn by the robot
-at its first attempt and the shape of the reference letter) and the robot's
-score at the current round of demonstrations. We then correlate this progress
-to the positive or negative feedback provided by the children (details of the
-method are presented in~\cite{jacq2016building}).
-
-We find that the feedback of five out of the eight
-children do significantly correlate with the actual performance of the robot (Table~\ref{table:scores}).
+see Table~\ref{table:scores}). This indicates that they are all able to engage
+into playing the role of a teacher.
+
+To study the correlation between the children' feedback and the actual
+improvements of the robot, we estimate the robot's progress as the difference
+between an initial score (average euclidian distance between the shape drawn by
+the robot at its first attempt and the shape of the reference letter) and the
+robot's score at the current round of demonstrations. We then correlate this
+progress to the positive or negative feedback provided by the children (details
+of the method are presented in~\cite{jacq2016building}).
+
+We find that the feedback of five out of the eight children do significantly
+correlate with the actual performance of the robot (Table~\ref{table:scores}).
 This indicates that these children are effectively taking on the teacher's
-clothes and are seriously providing feedback to the robot. The
-observation of the three remaining children reveal a variety of behaviours (for
-instance, one was actually rating how ``nice'' the robot is, and another one
-had rather ``artistic'' writing style preferences that were independent of the
-actual legibility of the robot), and during this specific experiment, we eventually
+clothes and are seriously providing feedback to the robot. The observation of
+the three remaining children reveal a variety of behaviours (for instance, one
+was actually rating how ``nice'' the robot is, and another one had rather
+``artistic'' writing style preferences that were independent of the actual
+legibility of the robot), and during this specific experiment, we eventually
 found that they were all seriously engaged into teaching.
 
-To summarize, this experiment shows that children with actual handwriting impairments
-do accept well the robot, can commit themselves into long interactions, do take
-on the role of the teacher (as evidenced by both the number of writing demonstrations
-they provided and their self-inclination to give feedback to the robot), and
-finally take their teacher's role seriously.
+To summarize, this experiment shows that children with actual handwriting
+impairments do accept well the robot, can commit themselves into long
+interactions, do take on the role of the teacher (as evidenced by both the
+number of writing demonstrations they provided and their self-inclination to
+give feedback to the robot), and finally take their teacher's role seriously.
 
@@
 For the first case study, we invited and followed Vincent, a six years old
-child, once per week over a period of a month. Our primary aim was to address the
-question of whether we can sustain Vincent's engagement and commitment to the
-writing activity over such a period.
+child, once per week over a period of a month. Our primary aim was to address
+the question of whether we can sustain Vincent's engagement and commitment to
+the writing activity over such a period.
 
@@
 boy: one of our {\sc nao} was away for a mysterious scientific mission, and it
-needed the support of another one -- which would remain at the lab -- to interpret
-curious pictures that were sent every week. Vincent had to help the second robot
-understanding the pictures, and since the two robots had somehow beforehand
-agreed to communicate with ``letters written like humans'' (\ie handwritten),
-Vincent also had to help the robot to write good looking letters (because, well,
-this robot was terrible at writing!) The experimental setup was similar to
-Figure~\ref{experimental_setup}, except that Vincent had to tell the robot what
-to write with small plastic letters (visible behind the robot on Figure~\ref{fig:diego}).
-
-To supplement the intrinsic motivation of helping a robot to communicate with another one, we
-gradually increased the complexity of Vincent's task to keep it challenging and
-interesting (the first week: demonstration of single letters; the second week:
-short words; the third week: a full letter -- Figure~\ref{fig:stimuli}).
-
-The last session was set as a test: the ``explorer'' robot
-had come back from its mission and it actually challenged the other robot in
-front of Vincent: \emph{``I don't believe you wrote yourself these nice letters that I
-received! Prove it to me by writing something in front of me!''} This situation
-was meant to evidence the Protégé effect: by judging the other robot's
-handwriting, the ``explorer'' robot would implicitly judge Vincent's skills as
-teacher, and in turn, Vincent's handwriting.
+needed the support of another one -- which would remain at the lab -- to
+interpret curious pictures that were sent every week. Vincent had to help the
+second robot understanding the pictures, and since the two robots had somehow
+beforehand agreed to communicate with ``letters written like humans'' (\ie
+handwritten), Vincent also had to help the robot to write good looking letters
+(because, well, this robot was terrible at writing!) The experimental setup was
+similar to Figure~\ref{experimental_setup}, except that Vincent had to tell the
+robot what to write with small plastic letters (visible behind the robot on
+Figure~\ref{fig:diego}).
+
+To supplement the intrinsic motivation of helping a robot to communicate with
+another one, we gradually increased the complexity of Vincent's task to keep it
+challenging and interesting (the first week: demonstration of single letters;
+the second week: short words; the third week: a full letter --
+Figure~\ref{fig:stimuli}).
+
+The last session was set as a test: the ``explorer'' robot had come back from
+its mission and it actually challenged the other robot in front of Vincent:
+\emph{``I don't believe you wrote yourself these nice letters that I received!
+Prove it to me by writing something in front of me!''} This situation was meant
+to evidence the Protégé effect: by judging the other robot's handwriting, the
+``explorer'' robot would implicitly judge Vincent's skills as teacher, and in
+turn, Vincent's handwriting.
 
@@
 
-Over the whole duration of the study, Vincent had provided 154 demonstrations to the robot, and he remained
-actively engaged over the four weeks. The story was well accepted by the child and
-he seriously engaged into the game. After the first week, he showed good
-confidence to play with the robot and by the end of the study he had built
-affective bonds with the robot, as evidenced by several letters he did send to the
-robot \emph{after} the end of the study (one of them four months later) to get
-news. This represents an initial validation of our hypothesis: our system can effectively keep a
-child engaged with the robot for a relatively long period of time (about 5
-hours spread over a month), and we can build a tutor/protégé relationship.
-
-No hard conclusion can be drawn in terms of actual handwriting remediation as we did
-not design this study to formally assess possible improvements. However, as
+Over the whole duration of the study, Vincent had provided 154 demonstrations to
+the robot, and he remained actively engaged over the four weeks. The story was
+well accepted by the child and he seriously engaged into the game. After the
+first week, he showed good confidence to play with the robot and by the end of
+the study he had built affective bonds with the robot, as evidenced by several
+letters he did send to the robot \emph{after} the end of the study (one of them
+four months later) to get news. This represents an initial validation of our
+hypothesis: our system can effectively keep a child engaged with the robot for a
+relatively long period of time (about 5 hours spread over a month), and we can
+build a tutor/protégé relationship.
+
+No hard conclusion can be drawn in terms of actual handwriting remediation as we
+did not design this study to formally assess possible improvements. However, as
 visible on Figure~\ref{fig:stimuli}, Vincent was able to significantly improve
-the robot's skill, and  he acknowledged that he
-was the one helping the robot during post-hoc interviews. In that respect, Vincent realized that he was
-``good enough'' at writing to help someone else. The fact that Vincent seriously
+the robot's skill, and  he acknowledged that he was the one helping the robot
+during post-hoc interviews. In that respect, Vincent realized that he was ``good
+enough'' at writing to help someone else. The fact that Vincent seriously
 committed to the role of a teacher is further supported by feedback sent by
-Vincent's parents a week after the end of the experiment: ``Vincent's handwriting has changed over the
-last weeks, going from a mix of standalone and cursive letters to full words in
-cursive. This requires a lot of efforts and concentration from him, but he did
-succeed during the sessions with the robot as he knew he had to show a
-consistent writing to the robot''.
+Vincent's parents a week after the end of the experiment: ``Vincent's
+handwriting has changed over the last weeks, going from a mix of standalone and
+cursive letters to full words in cursive. This requires a lot of efforts and
+concentration from him, but he did succeed during the sessions with the robot as
+he knew he had to show a consistent writing to the robot''.
 
@@
 
-The second long-term study was designed in collaboration with an occupational therapist in
-Geneva, and aimed at deploying the system on the longer-term with a real
-therapeutic case.
-
-Thomas is a 5.5 years old child. He
-has been diagnosed with visuo-constructive deficits, which translate into
-difficulties for him to consistently draw letters. Besides, focusing on a task
-is difficult for Thomas, who tends to rapidly shift his attention to other
-things. Since the robot's learning algorithm requires repeated demonstrations of
-similarly shaped letters to converge, the occupational therapist was especially
-interested in observing if the robot would induce a strong enough motivation for
-Thomas to focus on producing many regular, consistent letters, thus overcoming his
-deficit.
+The second long-term study was designed in collaboration with an occupational
+therapist in Geneva, and aimed at deploying the system on the longer-term with a
+real therapeutic case.
+
+Thomas is a 5.5 years old child. He has been diagnosed with visuo-constructive
+deficits, which translate into difficulties for him to consistently draw
+letters. Besides, focusing on a task is difficult for Thomas, who tends to
+rapidly shift his attention to other things. Since the robot's learning
+algorithm requires repeated demonstrations of similarly shaped letters to
+converge, the occupational therapist was especially interested in observing if
+the robot would induce a strong enough motivation for Thomas to focus on
+producing many regular, consistent letters, thus overcoming his deficit.
 
@@
 first session. Hence, we also tested during this case-study if the robot (and
-the Protégé effect) would induce by itself a strong enough intrinsic motivation to
-keep the child engaged over the five weeks.
+the Protégé effect) would induce by itself a strong enough intrinsic motivation
+to keep the child engaged over the five weeks.
 
@@
     \includegraphics[width=0.7\linewidth]{learning_6_distances}
-    \caption{\small Demonstrations provided by Thomas for the number ``6'' (top row) and
-        corresponding shapes generated by the robot. The plot beneath shows the
-        distance to the reference shape (in the eigenspace of the shape) to
-        Thomas' demonstrations, and to the
-        robot's attempts. After eight demonstrations,
-        Thomas decided that the robot's ``6'' was good enough, and switched to
-    another character. In that respect, he was the one leading the learning
-process of the robot.}
+    \caption{\small Demonstrations provided by Thomas for the number ``6'' (top
+        row) and corresponding shapes generated by the robot. The plot beneath
+        shows the distance to the reference shape (in the eigenspace of the
+        shape) to Thomas' demonstrations, and to the robot's attempts. After
+        eight demonstrations, Thomas decided that the robot's ``6'' was good
+        enough, and switched to another character. In that respect, he was the
+        one leading the learning process of the robot.}
+
     \label{learning_6_demos}
@@
 
-Since Thomas would frequently draw mirrored numbers, or hard-to-recognize
-shapes (see Figure~\ref{Thomas_progress}), the learning algorithm of the robot initially tended to converge towards
-meaningless scribblings. We addressed this issue by having the robot to
-\emph{refuse} allographs that were too far from the reference shape (the robot
-would instead say \emph{``I'm
-not sure to understand what you are drawing...''}), so that the child had to
-pay good attention to what he would demonstrate to the robot. Also, to make
-the robot's progress evident, we modified the initialization step of the
-learning algorithm to start with a roughly vertical stroke instead of a
-deformed number (see the initial state on Figure~\ref{learning_6_demos}).
+Since Thomas would frequently draw mirrored numbers, or hard-to-recognize shapes
+(see Figure~\ref{Thomas_progress}), the learning algorithm of the robot
+initially tended to converge towards meaningless scribblings. We addressed this
+issue by having the robot to \emph{refuse} allographs that were too far from the
+reference shape (the robot would instead say \emph{``I'm not sure to understand
+what you are drawing...''}), so that the child had to pay good attention to what
+he would demonstrate to the robot. Also, to make the robot's progress evident,
+we modified the initialization step of the learning algorithm to start with a
+roughly vertical stroke instead of a deformed number (see the initial state on
+Figure~\ref{learning_6_demos}).
 
@@
 
-Despite his attention deficit, Thomas was able to remain engaged in the activity during more than
-forty minutes in each session (a long time for a five years old). In total, 55 allographs out of 82 
-demonstrated by the child were acceptable considering our threshold (with a
-progressive improvement from 13 out of 28 in the first session up to 26 out
-of 29 in the last session).
+Despite his attention deficit, Thomas was able to remain engaged in the activity
+during more than forty minutes in each session (a long time for a five years
+old). In total, 55 allographs out of 82 demonstrated by the child were
+acceptable considering our threshold (with a progressive improvement from 13 out
+of 28 in the first session up to 26 out of 29 in the last session).
 
@@
 allographs, he started to focus on it and he would typically draw 5 or 6 times
-the number before actually sending to the robot (the tablet lets children
-clear their drawing and try again before sending it). According to
-the therapist, it was the first time that Thomas was seen to correct himself in such a
-way, explicitly having to reflect on how \emph{another agent} (the robot) would
-interpret and understand his writing. Figure~\ref{Thomas_progress} shows how
-he gradually improved his demonstrations for two different numbers.
+the number before actually sending to the robot (the tablet lets children clear
+their drawing and try again before sending it). According to the therapist, it
+was the first time that Thomas was seen to correct himself in such a way,
+explicitly having to reflect on how \emph{another agent} (the robot) would
+interpret and understand his writing. Figure~\ref{Thomas_progress} shows how he
+gradually improved his demonstrations for two different numbers.
 
@@
 The four school studies, the experiment at the surgery and the two case-studies
-provide a first broad picture of how a robot-based remediation to handwriting
-is accepted by the children and practitioners, and what are the outcomes that
-can be expected.
+provide a first broad picture of how a robot-based remediation to handwriting is
+accepted by the children and practitioners, and what are the outcomes that can
+be expected.
 
@@
 
-We believe that this research provides a novel perspective on educative
-robots at several levels. We show that:
+We believe that this research provides a novel perspective on educative robots
+at several levels. We show that:
 
 \begin{itemize}
-    \item robots in an educative context are certainly \textbf{relevant and effective
-        beyond STEM} (Science, Technology, Engineering and Mathematics)
-    \textbf{teaching},
+    \item robots in an educative context are certainly \textbf{relevant and
+        effective beyond STEM} (Science, Technology, Engineering and
+        Mathematics) \textbf{teaching},
 
@@
 
-    \item this proves essential to \textbf{sustain a long-term interaction} (several
-        hours) although the task would be typically considered as repetitive yet
-        challenging by the children,
+    \item this proves essential to \textbf{sustain a long-term interaction}
+        (several hours) although the task would be typically considered as
+        repetitive yet challenging by the children,
 
     \item besides, the social commitment induces a \textbf{cognitive engagement}
-        of the child with the robot, which is \textbf{a key learning lever} as
-        its elicits \textbf{reflective, meta-cognitive mechanisms on the
-        learning task}.
+        of the child with the robot, which is \textbf{a key learning lever} as its
+        elicits \textbf{reflective, meta-cognitive mechanisms on the learning task}.
+
 \end{itemize}
@@
 handwriting is a complex cognitive skill, that builds on many individual and
-social factors. Self-confidence is one of them. Our
-approach that endows the child with the role of a teacher who can help a robot:
-we expect it may help some children to recover self-esteem and self-confidence. However the
+social factors. Self-confidence is one of them. Our approach that endows the
+child with the role of a teacher who can help a robot: we expect it may help
+some children to recover self-esteem and self-confidence. However the
 experiments that we have conducted so far do not allow us to confirm this
@@
 going to replace them any time soon, and the initial feedback that we received
-during the focus groups shows that, once explained, our approach
-make sense and is indeed welcomed by the teachers themselves.
+during the focus groups shows that, once explained, our approach make sense and
+is indeed welcomed by the teachers themselves.
 
 The implication of the mentor-protégé relationship on the children is less
-clearly understood. We have certainly seen that the children can establish strong affective bonds with the robot (as
-witnessed for instance by the letter sent by Vincent several months after he interacted with
-the robot), but we are not yet able to precisely characterize these bonds. The
-ethical implications of the mentor-protégé relationship have been explored before
-in the context of human teaching~\cite{brad1999mentor,wendelyn2008context}, but
-they mostly looked at the question from the perspective of the protégé, whereas in our
-case, the child is the mentor. As such, relatively little is known on the
-psychological implications for a child to commit in helping a robot, and as
-advocated by Belpaeme and Morse~\cite{belpaeme2010time}, we likely need to first
-gain more field experience before being able to draw conclusions.
+clearly understood. We have certainly seen that the children can establish
+strong affective bonds with the robot (as witnessed for instance by the letter
+sent by Vincent several months after he interacted with the robot), but we are
+not yet able to precisely characterize these bonds. The ethical implications of
+the mentor-protégé relationship have been explored before in the context of
+human teaching~\cite{brad1999mentor,wendelyn2008context}, but they mostly looked
+at the question from the perspective of the protégé, whereas in our case, the
+child is the mentor. As such, relatively little is known on the psychological
+implications for a child to commit in helping a robot, and as advocated by
+Belpaeme and Morse~\cite{belpaeme2010time}, we likely need to first gain more
+field experience before being able to draw conclusions.
 
@@
 but its potential to positively impact the child's motivation and self-esteem.
-While we need to carefully clear up the possible ethical concerns, we hope that this article shows that this is a very relevant context of use for
-robots: when facing a child with school difficulties, robots can play the role
-of a naïve learner which neither adults nor peers can convincingly play.
+While we need to carefully clear up the possible ethical concerns, we hope that
+this article shows that this is a very relevant context of use for robots: when
+facing a child with school difficulties, robots can play the role of a naïve
+learner which neither adults nor peers can convincingly play.
 
@@
 of Geneva, the Institut Florimont and the Haut-Lac school for their
-participation and support. We express our gratitude to Anne Gossin and
-Fanny Lebel for inviting us to their surgery, for their involvement in the
-experiments and for bearing with our disruptive robots. And finally, we want to
-say a very special \emph{Thank you!} to Vincent, Thomas, Valérie, Émilien,
-Mathieu, Johan, Nicolas, Antoine, Adèle and Marie: by spending several hours
-being patient teachers for our robots, you have paved the way for what may
-become tomorrow a new way of learning how to deal with all these letters and
-words!
+participation and support. We express our gratitude to Anne Gossin and Fanny
+Lebel for inviting us to their surgery, for their involvement in the experiments
+and for bearing with our disruptive robots. And finally, we want to say a very
+special \emph{Thank you!} to Vincent, Thomas, Valérie, Émilien, Mathieu, Johan,
+Nicolas, Antoine, Adèle and Marie: by spending several hours being patient
+teachers for our robots, you have paved the way for what may become tomorrow a
+new way of learning how to deal with all these letters and words!
 

### 2015-12-17 main.tex
Discussion of the findings

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Summary of the Findings}
+\section{Summary of the Findings and Discussion}
 
@@
 provide a first broad picture of how a robot-based remediation to handwriting is
-accepted by the children and practitioners, and what are the outcomes that can
-be expected.
+accepted by the children and practitioners, and what outcomes can be expected.
 
 In total, more than 80 children have interacted with the system, for a total
-duration of more than 38 hours in multiple experimental configurations
+duration of more than 38 hours, in multiple experimental configurations
 (individual interaction \vs pairs \vs groups; at school, in the lab, at a
 surgery; short interactions \vs long, repeated interactions; 5 to 8 years old
-children).
+children). We summarize hereafter the main findings from these experiments and
+discuss some of the critical points of our approach.
+
+\paragraph{Technical Assessment} We set ourselves the challenge of developing an
+autonomous robotic system able to perform handwriting tasks with children.
+Over the course of the different studies and experiments, thousands of
+letters have been indeed generated and drawn on the tactile tablet by the robot, in
+response to thousands of demonstrations from children. Most of the experiments have been
+carried out in the wild, at six different locations in total. As reported in the
+technical studies section, the system performed generally well: the children had no
+issues interacting with the robot, the robot experienced relatively few crashes,
+and those crashes did not significantly impaired the interactions. The system
+has furthermore proved robust enough to conduct several hour-long experiments.
+However, we also found that the level of expertise required to deploy and
+operate our system still makes the presence of an experimenter mandatory at all
+time. Non-trivial technical developments may be required to reach a level of
+usability suitable for a broader, non-expert, audience.
+
+\paragraph{Acceptance} When we started the project, the mere acceptance of the
+robot in the educative and/or medical context of handwriting remediation was not
+obvious. Expectedly, the attitude of practitioners toward the robot indeed
+varies a lot at individual levels. The process that led us to select the schools
+and surgeries for the experiments is interesting in that regard. The very first
+school we contacted was a local public school. The head teacher was positive
+about the study, but did ask for official agreement, which never happened. As we
+learnt several months later, the request was still laying on the desk of a
+regional education officer who happens not to be too keen on novel technologies.
+Conversely, the three schools where we did conduct the first studies are all
+private schools, generally open to novel learning methods. The teachers we met
+during these experiments certainly had questions about the rationale and
+end-purpose of the project, but were confident about its educative and social
+value.
+
+Similarly with the occupational therapists: while Thomas' therapist did
+voluntarily contacted us (after hearing about the project at the radio) and
+readily involved herself in the design of the experiment with Thomas, we had
+less positive feedback from other practitioners: in one instance, we contacted
+ourselves a local group of school psychologists and occupational therapists to
+present the project: after the meeting, the opinions were definitively mixed,
+with some therapists willing to conduct actual experiments with their children,
+and others not quite as enthusiastic.
+
+Lastly (and as expected), the reaction of the children to the robot was good:
+they enjoyed interacting with the system, and, as we have shown, they actually
+committed to their teacher role. The teaching situation was well accepted: while
+we had initial doubts about how believable a learner the robot would be, the
+children did not appear to show any specific issues with it. In particular,
+technical choices like having the robot to only gesture writing on the tablet
+instead of actually \emph{physically} writing with a pen, were not raised as
+issues.
+
+\paragraph{Engagement} We measured the engagement of the children by three
+different means: the number of demonstrations they provided to the robot, the
+amount of qualitative feedback they gave to the robot (in the \emph{Surgery}
+experiment), and to what extend they were focusing on the task (measure of the
+with-me-ness). Independently of the experimental setting, it appears that the
+children engage easily into the interaction. More interestingly, we show in the
+\emph{Surgery} experiment that they generally take on the role of a teacher
+easily, that they act this role seriously (and not only playfully), and that
+they can assume this role over an extended period of time. Designing a system
+that keeps children engaged over several hours is especially important for
+handwriting remediation, and we show in the case-studies that even children as
+young as 5.5 years old like Thomas were able to do so. This seems to indicate
+that our approach could be relevant for a broad range of ages.
+
+Another finding relates to the Protégé effect: as seen with Vincent's case-study
+(with the parents emphasizing how Vincent was aware he had to be consistent in
+his writing to help the robot, or with him sending us a mail several months
+later to know how the robot was doing with its writing) or with Thomas's
+case-study (when Thomas realizes that the robot ``does not understand'' when his
+demonstrations are not legible enough, and consequently quickly improves his
+own writing to better help the robot), our system does seem to elicit a Protégé
+effect that not only help the children to remain engaged into their teaching
+over several weeks, but also positively impact their learning process (Vincent
+strives to write in a consistent manner, as does Thomas). This supports {\it a
+posteriori} our choice to build an interaction situation based on the
+\emph{learning by teaching} paradigm.
+
+\paragraph{Remediation Efficacy} Vincent's case study did provide us with
+initial material to evidence handwriting improvements (Figure~\ref{fig:stimuli})
+and the study with Thomas provide further data, both quantitative
+(Figure~\ref{Thomas_progress}) and qualitative (feedback from the therapist that
+points how Thomas is much better at drawing consistent shapes in a repeated manner,
+as well as reflecting on his own performance by training several times before
+actually sending a demonstration to the robot).
+
+We must however remain cautious here as for the actual role of the system: while
+the children were indeed the one deciding what to teach to the robot and the
+robot was autonomously learning and responding, the role of the adults (the
+experimenters or the occupational therapists) should not be underestimated.
+Beyond the normal explanations on how to operate the tablet and how to interact
+with the robot, the adults played the role of a facilitator in every of the
+studies by prompting the children to comment on the robot's performance,
+suggesting possible corrections, or proposing to try another letter/number/word.
+This facilitation not only compensates for the possible shortcomings of the interaction,
+but is also a fundamental part of the learning process itself. In that respect,
+our robot is essentially a tool that creates a favourable learning situation for
+the child, and where the adult (be it a teacher or a therapist) keeps its entire
+educative role.
 
@@
 We believe that this research provides a novel perspective on educative robots
-at several levels. We show that:
+at several levels. We have shown that:
 
@@
 actual handwriting impairments. Our results are promising in terms of sustaining
-the interest and attention of children on otherwise boring writing exercises,
+the interest and attention of children on otherwise repetitive writing exercises,
 and we evidence handwriting improvements.
@@
 handwriting is a complex cognitive skill, that builds on many individual and
-social factors. Self-confidence is one of them. Our approach that endows the
-child with the role of a teacher who can help a robot: we expect it may help
-some children to recover self-esteem and self-confidence. However the
-experiments that we have conducted so far do not allow us to confirm this
-hypothesis, and more research need to be conducted in this direction.
+social factors. Self-confidence is one of them. Our approach endows the child
+with the role of a teacher who can help a robot: we expect it may as well help
+some children to recover self-esteem and self-confidence by putting them in a
+positive, gratifying role. The experiments that we have conducted so far do not
+allow us to confirm this hypothesis yet, and more research will have to be
+conducted in this direction.
 

### 2015-12-17 main.tex
Minor modification in the 'ethical concerns' section

@@
 The importance of the adult is further supported by our experiments: even with
-an autonomous, nominally performing robot, to put the teacher's clothes on
-remains (expectedly) difficult for 5-6 years old children, and during the
-experiments we conducted, the adult always played a key role at prompting the
-child to give feedback to the robot or to move to the next letter or word.  At a
-higher orchestration level (and as reported in the two case studies with Vincent
-and Thomas), the educational scenarios were also always designed and monitored
-by the adults.
+an autonomous, nominally performing robot, to put the teacher's clothes on is
+not obvious for 5-6 years old children, and during the experiments we conducted,
+the adult always played a key role at prompting the child to give feedback to
+the robot or to move to the next letter or word.  At a higher orchestration
+level (and as reported in the two case studies with Vincent and Thomas), the
+educational scenarios were also always designed and monitored by the adults.
 

### 2015-12-27 main.tex
Remove extraneous info re eigenvalues

@@
 
-By changing the eigenvalues associated to these eigenvectors and regenerating
-letters with the reverse procedure, it becomes then easy to generate new
-letters, with distortions that are actually plausible handwriting errors: they
-are exaggerations of variance of writing styles that naturally occurs amongst
+New letters can be generated by varying the weightings in linear combinations of
+these eigenvectors, with distortions that are actually plausible handwriting errors: 
+they are exaggerations of variations of writing styles that naturally occur amongst
 adult writers.  Figure~\ref{fig:sampleLetters} shows examples of deformed ``g''
@@
     \caption{\small \label{fig:sampleLetters} \textbf{Generating bad letters}:
-        effect of varying the first five eigenvalues (rows) of the shape model
-        of ``g'' by different factors (columns). The percentage of the total
-        variance in the dataset explained by each eigenvalue is shown      on
-        the appropriate row. Examples noted A, B, C, D illustrate how the
-        PCA-based approach allows to \emph{automatically} generate letters whose
-        errors can be \emph{semantically} interpreted, \eg A has a too large bottom
+        effect of independently varying the weightings (columns) of the first five eigenvectors 
+        (rows) of the shape model of ``g''. Examples A, B, C, D illustrate how the
+        PCA-based approach allows for the \emph{automatic} generation of letters whose
+        errors can be \emph{semantically} interpreted, \eg A has too large a bottom
         loop, B has a wide top loop, the bottom loop of C is not correctly
@@
 
-The same technique can also be applied to \emph{classify} demonstrations,
+The same shape model can also be used to \emph{classify} demonstrations,
 \emph{assess} their quality and \emph{learn} from them. Figure~\ref{fig:h} shows
 for example nine allographs of ``h'' written by a 6 years old child, along with
-the mean letter from the dataset (reference letter). By projecting each of the
+a reference letter. By projecting each of the
 demonstrations onto the eigenspace of ``h'' (Figure~\ref{fig:h}b), we observe

### 2015-12-27 main.tex
Copyedit

@@
     Aude Billard$^2$, Ana Paiva$^3$, Pierre Dillenbourg$^1$ \\
-$^1$CHILI Lab, École Polytechnique Fédérale de Lausanne, Suisse,\\
-$^2$LASA, École Polytechnique Fédérale de Lausanne, Suisse,\\
+$^1$CHILI Lab, École Polytechnique Fédérale de Lausanne, Switzerland,\\
+$^2$LASA, École Polytechnique Fédérale de Lausanne, Switzerland,\\
 $^3$Instituto Superior Técnico, University of Lisbon, Portugal}
@@
 deficits. He is under the care of an occupational therapist, and tries to
-workaround his inability to draw letters in a consistent manner. Vincent is six
+work around his inability to draw letters in a consistent manner. Vincent is six
 and struggles at school with his poor handwriting and even poorer
@@
     \includegraphics[width=0.9\linewidth]{henry}
-    \caption{\small Thomas teaching Nao how to write numbers, with the help of an
-    occupational therapist.}
+    \caption{\small Thomas teaching a \sc{Nao} robot how to write numbers, with 
+    the help of an occupational therapist.}
     \label{fig:henry}
@@
 Remediations for handwriting difficulties traditionally involve long
-interventions (at least 10 weeks~\cite{Hoy2011}), essentially consisting in
+interventions (at least 10 weeks~\cite{Hoy2011}), essentially consisting of
 handwriting training with occupational therapists, and primarily addressing the
@@
 We present in this article a new take on this educative challenge: a remediation
-procedure that involve a ``bad writer'' robot that is \emph{taught} by the
-child. By building on the \emph{learning by teaching} paradigm, not only the
-child practises handwriting but, as they take on the role of the teacher, they
+procedure that involves a ``bad writer'' robot that is \emph{taught} by the
+child. By building on the \emph{learning by teaching} paradigm, not only does the
+child practise handwriting but, as they take on the role of the teacher, they
 also positively reinforce their self-esteem and motivation: their social role
-shift from the ``underperformer'' to ``the one who knows and teaches''. And by
+shifts from the ``underperformer'' to ``the one who knows and teaches''. And by
 relying on a robot, we can tailor the exercises and the learning curve to each
-children' needs, as we will show in this article.
+child's needs individually, as we will show in this article.
 
@@
 ``learning from demonstration'' approaches to robots learning physical skills,
-as the agent we present is only simulating fine motor skills for interaction
+as the agent we present is only simulating poor fine motor skills for interaction
 purposes.
@@
 system, children learn vocabulary by teaching the {\sc nao} robot to act out
-verbs. The robot is tele-operated (Wizard-Of-Oz) and mimics the actions that the
+verbs. The robot is tele-operated (Wizard-of-Oz) and mimics the actions that the
 children teach it, but with no long-term memory or learning algorithm in place.
@@
 learner: the child is to engage in a (meta-) cognitive relationship with the
-robot to try to understand why the robot fails and how to help it best.  Here,
+robot to try to understand why the robot fails and how to best help it.  Here,
 the robot is more than just an activity facilitator or orchestrator -- its
-physical presence and embodiment induce agency and anthropomorphising, and
+physical presence and embodiment induce agency and anthropomorphizing, and
 cognitively engage the child into the learning activity (be it consciously or
@@
 
-The commitment of the child into the interaction build on a psychological effect
+The commitment of the child into the interaction builds on a psychological effect
 known as the ``protégé effect''~\cite{Chase2009}: the teacher feels responsible
@@
 indeed show sustained child-robot engagement over extended periods of time
-(several hours spread of a month).
+(several hours spread over a month).
 
 For these two reasons, our approach is to be distinguished from previous works
-in educational robotic. Most of these do not consider the agency induced by the
+in educational robotics. Most of these do not consider the agency induced by the
 robot beyond its motivational aspect (playing with an interactive, partially
 autonomous device naturally induces some form of anthropomorphizing, which leads
-to some level of projected agency, and participate the overall excitement -- at
+to some level of projected agency, and contributes to the overall excitement -- at
 least, on the short-term, before the novelty effect vanishes).  In our case, the
@@
 for the learning process; it also induces a protégé effect (``I want my
-robot-agent to succeed!'') which support the commitment of child into the
-interaction, also for longer period of time.
+robot-agent to succeed!'') which supports the commitment of the child into the
+interaction, also for longer periods of time.
 
@@
 a robotic system that would effectively support handwriting remediation in an
-original way: by getting children to teach a robotic agent how to write, those
+original way. By getting children to teach a robotic agent how to write, those
 children would both practise without knowing it and recover self-confidence and
-self-esteem by supporting a worst-than-themselves robotic ``student''.
-
-The following sections of the paper go into details. We first provides a brief
+self-esteem by supporting a worse-than-themselves robotic ``student''.
+
+The following sections of the paper go into the details. We first provide a brief
 overview of the robotic system and the interaction it induces. We also present
@@
 We then present and report on the field experiments that we have conducted over
-the last two years, including four studies at school, one longer experiment with
-eight children at the surgery of an occupational therapist, and two one-month
+the last two years, including four studies at schools, one longer experiment with
+eight children in an occupational therapy clinic, and two one-month
 long case studies. While the focus of the school experiments was mostly the
 technical validation and data acquisition, the three other experiments involved
-children with actual deficits, and gave us initial insights on the relevance and
+children with actual deficits, and gave us initial insights into the relevance and
 effectiveness of our approach.
@@
 child to write. During a typical round, the child requests the robot to write
-something (a single letter, a number or a full word), and push the tablet
+something (a single letter, a number or a full word), and pushes the tablet
 towards the robot. The robot writes on the tablet by gesturing the writing in
 the air,  the letters being actually drawn by the tablet application itself. The
-child then pull back the tablet, corrects the robot's attempt by writing
-him/herself on top or next to the robot's writing (see Figure~\ref{fig:diego}),
-and ``send'' his/her demonstration to the robot by pressing a small button on
+child then pulls back the tablet, corrects the robot's attempt by writing
+him/herself on top of or next to the robot's writing (see Figure~\ref{fig:diego}),
+and ``sends'' his/her demonstration to the robot by pressing a small button on
 the tablet. The robot ``learns'' from this demonstration and tries again.
@@
 think?''), and pressing on a small robot icon on the tablet once the child has
-finished correcting. In our experiments, and once introduced by the
-experimenter, both were easy to grasp for all the children.
+finished correcting. In our experiments, once introduced by the
+experimenter, both steps were easy to grasp for the children.
 
@@
 Implementing such a system raises several challenges: first, the acquisition,
-analyse and learning from hand-written demonstration, which lays at the core of
+analysis and learning from hand-written demonstration, which lays at the core of
 the our approach, necessitates the development of several algorithms for the
@@
 several modules (from performing gestures and acquiring the user's input to the
-state machine implementing the high-level behaviour), spread over several
+state machine implementing the high-level behavior), spread over several
 devices (the robot itself, one laptop and up to four tactile tablets for some of
@@
 
-    \item we can compute an euclidian distance to the reference letter to assess
+    \item we can compute a euclidian distance to the reference letter to assess
         the topological proximity of the demonstration with the expected letter,
@@
     }
-    \subfigure[Same samples, normalized and projected in the eigenspace spanned from the
-        3 first eigenvectors: clusters arise, that actually match writing styles.]{
+    \subfigure[The same samples, normalized and projected in the eigenspace spanned by the
+        first 3 eigenvectors: clusters arise, that actually match writing styles.]{
         \includegraphics[width=0.45\columnwidth]{eigenspace-uniformization}
@@
 The algorithm for machine-learning becomes then a simple matter of converging at
-a specific pace towards the child demonstration in the eigenspace.
+a specific pace towards the child's demonstration in the eigenspace.
 Figure~\ref{learning_6_demos} (p.~\pageref{learning_6_demos}) illustrates the
@@
 
-Our system is embodied in an Aldebaran's {\sc nao} (V4 or V5, depending on the
-studies) humanoid robot. This choice is motivated by its approachable
+Our system is embodied in an Aldebaran {\sc nao} (V4 or V5, depending on the
+study) humanoid robot. This choice is motivated by its approachable
 design~\cite{Gouaillier2008}, its size (58cm) and inherently safe structure
 (lightweight plastic) making it suitable for close interaction with children,
-its low price (making it closer to what school may afford in the coming years)
-and finally its ease of deployment on the field.
+its low price (making it closer to what schools may afford in the coming years)
+and finally its ease of deployment in the field.
 
@@
 for \emph{simulated handwriting}: the robot draws letters in the air, and the
-actual writing is displayed on a synchronised tablet.
+actual writing is displayed on a synchronized tablet.
 
@@
 
-Since the system does not actually require any CPU-intensive process, the laptop
+Since the system does not actually require any CPU-intensive processes, the laptop
 can be removed and the whole logic run on the robot. Due to the relative
@@
 different schools (more than 70 children aged 5 to 8; relatively short duration
-interactions), in the surgery of an occupational therapist (8 children; each
+interactions), in an occupational therapy clinic (8 children; each
 interacting several hours with the system), and during two case studies that
@@
 {\it School A (2)} & Individual interaction at school         & 20 min                 & 6                                     & 5-6        \\ \midrule
-{\it Surgery}      & Individual interaction at surgery        & 3 sessions $\times$ 1h & 8                                     & 6-8        \\ \midrule 
+{\it Clinic}       & Individual interaction at clinic         & 3 sessions $\times$ 1h & 8                                     & 6-8        \\ \midrule 
 {\it Vincent}      & Case-study                               & 4 weeks $\times$ 1.5h  & 1                                     & 6          \\
@@
 not to select specifically underperforming children as having more children (73
-in total) was beneficial for these preliminary studies. Besides, due to ethical
+in total) was beneficial for these preliminary studies. In addition, due to ethical
 concerns, this would have required complex organization with the school that we
@@
 
-\paragraph{System validation}
+\paragraph{System Validation}
 
 From a technical perspective, the system achieves an acceptable level of
-reliability and allow a technically sound autonomous interaction. For instance,
+reliability and allows a technically sound autonomous interaction. For instance,
 during the second school study (\textit{School B}), the robot withstood
@@
 The otherwise technical limitations were related to some letters or writing
-style (most notably, the ones requiring multiple strokes per letter) not being
+styles (most notably, the ones requiring multiple strokes per letter) not being
 adequately processed by the learning algorithm. Support for such letters was
-added as a follow-up of the validation studies.
+added as a follow-up to the validation studies.
 
@@
 approach to be effective. When asked, no child indicated that they did not
-believe that the robot was writing by itself. There were, at times, questions
-about the robot's writing method at the beginning of the interaction, but when
+believe that the robot was writing by itself: that it was simply miming the 
+actions in the air as the tablet displayed the letters. There were, at times, questions
+about the robot's writing without a pen at the beginning of the interaction, but when
 advised that the robot ``tells the tablet what it wants to write,'' this was
-accepted by the children.  Besides, teachers interviewed for their feedback on
+accepted by the children.  In addition, teachers interviewed for their feedback on
 the system advised that children are asked to draw letters in the air in a
-similar manner as part of their handwriting education. The behaviour is hence
+similar manner as part of their handwriting education. The behavior is hence
 not unfamiliar to children.
 
-During two focus groups organized with teachers, the welcomed the approach and
+During two focus groups organized with teachers, they welcomed the approach and
 recognized it as useful and promising; this was \textit{a posteriori} confirmed
 by multiple spontaneous contacts made by parents and therapists who were looking
-forward to use the system with their children.
-
-\paragraph{Sustained engagement}
+forward to using the system with their children.
+
+\paragraph{Sustained Engagement}
 
@@
 
+We also conducted a quantitative assessment of the engagement levels of the
+children. Table~\ref{tab:results-with-me-ness} reports the levels of
+\emph{with-me-ness} of the children during the second study at school A.
+\emph{With-me-ness} is a quantifiable precursor of engagement: it measures the
+percentage of time spent by the child focusing on the task at hand. This metric
+was first devised in the context of computer-supported
+learning~\cite{sharma2014me}, and we have previously studied its applicability
+to human-robot interaction and formalized the exact methodology
+in~\cite{lemaignan2016realtime}\footnote{Note that with-me-ness is a metric that
+allows real-time computation by the robot itself: while we did not yet make
+use of it, it does in principle allow the robot to detect on-line possible
+disengagements, and eventually address them by adapting its physical
+behavior, switching to different activities, etc.}.
+
 \begin{table}[h!]
@@
 
-We also conducted a quantitative assessment of the engagement levels of the
-children. Table~\ref{tab:results-with-me-ness} reports the levels of
-\emph{with-me-ness} of the children during the second study at school A.
-\emph{With-me-ness} is a quantifiable precursor of engagement: it measures the
-percentage of time spent by the child focusing onto the task at end. This metric
-was first devised in the context of computer-supported
-learning~\cite{sharma2014me}, and we have previously studied its applicability
-to human-robot interaction and formalized the exact methodology
-in~\cite{lemaignan2016realtime}\footnote{Note that with-me-ness is a metric that
-allows real-time computation by the robot itself: while we did not yet make
-use of it, it does in principle allow the robot to detect on-line possible
-disengagements, and eventually address them by adapting its physical
-behaviour, switching to different activities, etc.}.
-
 The average with-me-ness is well above 80\%, and confirms that the children
-where very much engaged into these 20 minutes of interaction with the robot,
+were very much engaged into these 20 minutes of interaction with the robot,
 paying close attention to the task.
 
-The other three experiments (the group study at the surgery and the two
+The other three experiments (the group study at the clinic and the two
 case-studies) provide further qualitative evidence of engagement over longer
-interaction periods. In particular, as reported hereafter, the two 4-weeks long
+interaction periods. In particular, as reported hereafter, the two 4-week long
 case studies that have been conducted so far show that our system can sustain
-children' engagement over durations (5 hours) that are getting closer to what is
+children's engagement over durations (5 hours) that are getting closer to what is
 expected to have an impact in a real therapeutic context.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Surgery Study: How Children Take on the Role of a Teacher}\label{normandie}
+\subsection{Clinic Study: How Children Take on the Role of a Teacher}\label{normandie}
 
@@
 
-The experiment took place at the surgery of an occupational therapist in
+The experiment took place at an occupational therapy clinic in
 Normandy, France. Eight children participated, selected by the occupational
@@
 a neurologist, and has been diagnosed with specific language impairment. Given
-their age and school year, all those children would be expected to know how to
+their age and school year, all of these children would be expected to know how to
 correctly shape cursive letters.
 
-Over a period of two weeks, each child attended three time a one hour long
+Over a period of two weeks, each child attended three times a one hour long
 session (except for Marie and Adèle who only attended one session). The
-experimenter role was limited to the explanation of the task and the basic
+experimenter's role was limited to the explanation of the task and the basic
 tablet usage. For this experiment, the children were provided with two tablets:
 one to choose the words (or letters) to teach to the robot, the other one to
-write, like in the validation studies. We also provided paper-based templates of
-the cursive letters, where the children to ask for them.
+write, as in the validation studies. We also provided paper-based templates of
+the cursive letters, were the children to ask for them.
 
@@
 have to help the robot to improve its writing style), so as to assess if and how
-children would naturally take on the role of the teacher. We besides assessed
+children would naturally take on the role of the teacher. We additionally assessed
 how seriously they engage into helping the robot through two additional buttons
@@
 the robot, the more they assume the role of a teacher. Then, by correlating the
-feedback with the actual performance of the robot, we can measure to what extend
+feedback with the actual performance of the robot, we can measure to what extent
 the children are seriously assuming their role: if their feedback does correlate
@@
 
-All children maintained their engagement during the entire sessions. They
-provided on average 42 demonstrations per session. All children did use the
-feedback buttons (in total, 99 ``thumb up'' and 33 ``thumb down'' were recorded,
+All children maintained their engagement during the entirety of the sessions. They
+provided on average 42 demonstrations per session. All children used the
+feedback buttons (in total, 99 ``thumbs up'' and 33 ``thumbs down'' were recorded,
 see Table~\ref{table:scores}). This indicates that they are all able to engage
@@
 
-To study the correlation between the children' feedback and the actual
+To study the correlation between the children's feedback and the actual
 improvements of the robot, we estimate the robot's progress as the difference
-between an initial score (average euclidian distance between the shape drawn by
+between an initial score (euclidian distance between the shape drawn by
 the robot at its first attempt and the shape of the reference letter) and the
@@
 
-We find that the feedback of five out of the eight children do significantly
+We find that the feedback of five out of the eight children does significantly
 correlate with the actual performance of the robot (Table~\ref{table:scores}).
 This indicates that these children are effectively taking on the teacher's
-clothes and are seriously providing feedback to the robot. The observation of
-the three remaining children reveal a variety of behaviours (for instance, one
+role and are seriously providing feedback to the robot. The observation of
+the three remaining children reveal a variety of behaviors (for instance, one
 was actually rating how ``nice'' the robot is, and another one had rather
@@
 
-For the first case study, we invited and followed Vincent, a six years old
+For the first case study, we invited and followed Vincent, a six year old
 child, once per week over a period of a month. Our primary aim was to address
@@
     \caption{\small Vincent correcting {\sc nao}'s attempt by rewriting the
-        whole word. Empty boxes are drawn on the screen to serve as template for the child
+        whole word. Empty boxes are drawn on the screen to serve as a template for the child
         and to make letter segmentation more robust.}
@@
 The study took place at our laboratory (Figure~\ref{fig:diego}), and we chose to
-design the activity around a storyline meant to be attractive for a 6 years old
-boy: one of our {\sc nao} was away for a mysterious scientific mission, and it
+design the activity around a storyline meant to be attractive for a 6 year old
+boy: one of our {\sc nao}s was away for a mysterious scientific mission, and it
 needed the support of another one -- which would remain at the lab -- to
@@
 beforehand agreed to communicate with ``letters written like humans'' (\ie
-handwritten), Vincent also had to help the robot to write good looking letters
-(because, well, this robot was terrible at writing!) The experimental setup was
+handwritten), Vincent also had to help the robot to write good-looking letters
+(because, well, this robot was terrible at writing!). The experimental setup was
 similar to Figure~\ref{experimental_setup}, except that Vincent had to tell the
@@
         highlights one instance of striking improvement of the robot's
-        handwriting legibility.}
+        handwriting legibility \emph{`envoyer'}.}
 
@@
 
-Over the whole duration of the study, Vincent had provided 154 demonstrations to
+Over the whole duration of the study, Vincent provided 154 demonstrations to
 the robot, and he remained actively engaged over the four weeks. The story was
@@
 handwriting has changed over the last weeks, going from a mix of standalone and
-cursive letters to full words in cursive. This requires a lot of efforts and
+cursive letters to full words in cursive. This requires a lot of effort and
 concentration from him, but he did succeed during the sessions with the robot as
-he knew he had to show a consistent writing to the robot''.
+he knew he had to show a consistent style of writing to the robot''.
 
@@
 
-Thomas is a 5.5 years old child. He has been diagnosed with visuo-constructive
+Thomas is a 5.5 year old child. He has been diagnosed with visuo-constructive
 deficits, which translate into difficulties for him to consistently draw
-letters. Besides, focusing on a task is difficult for Thomas, who tends to
+letters. In addition, focusing on a task is difficult for Thomas, who tends to
 rapidly shift his attention to other things. Since the robot's learning
@@
 
-The experiment took place at the therapist's surgery (four sessions spreading
+The experiment took place at the therapist's clinic (four sessions spread
 over 5 weeks). Contrary to Vincent's experiment, we chose not to introduce any
-backstory beyond a simple prompt (``the robot wants to participate to a robotic
-handwriting contest, will you help him to prepare?'') only provided during the
+backstory beyond a simple prompt (``the robot wants to participate in a robotic
+handwriting contest, will you help him prepare?'') only provided during the
 first session. Hence, we also tested during this case-study if the robot (and
@@
 
-The occupational therapist had recently carried activities with Thomas about
+The occupational therapist had recently carried out activities with Thomas on
 writing numbers, so we decided with her to focus on these as well: Thomas would
@@
 initially tended to converge towards meaningless scribblings. We addressed this
-issue by having the robot to \emph{refuse} allographs that were too far from the
-reference shape (the robot would instead say \emph{``I'm not sure to understand
+issue by having the robot \emph{refuse} allographs that were too far from the
+reference shape (the robot would instead say \emph{``I'm not sure I understand
 what you are drawing...''}), so that the child had to pay good attention to what
@@
 Despite his attention deficit, Thomas was able to remain engaged in the activity
-during more than forty minutes in each session (a long time for a five years
+during more than forty minutes in each session (a long time for a five year
 old). In total, 55 allographs out of 82 demonstrated by the child were
@@
 Since the robot's handwriting started from a simple primitive (a stroke), each
-time Thomas succeeded in having his demonstration accepted by it, the robot's
+time Thomas succeeded in having his demonstration accepted by the robot, the
 improvement was clearly visible (as shown in Figure~\ref{learning_6_demos}).
@@
 
-The four school studies, the experiment at the surgery and the two case-studies
+The four school studies, the experiment at the clinic and the two case-studies
 provide a first broad picture of how a robot-based remediation to handwriting is
@@
 (individual interaction \vs pairs \vs groups; at school, in the lab, at a
-surgery; short interactions \vs long, repeated interactions; 5 to 8 years old
+occupational therapy clinic; short interactions \vs long, repeated interactions; 5 to 8 year old
 children). We summarize hereafter the main findings from these experiments and
@@
 issues interacting with the robot, the robot experienced relatively few crashes,
-and those crashes did not significantly impaired the interactions. The system
+and those crashes did not significantly impair the interactions. The system
 has furthermore proved robust enough to conduct several hour-long experiments.
@@
 robot in the educative and/or medical context of handwriting remediation was not
-obvious. Expectedly, the attitude of practitioners toward the robot indeed
+obvious. As expected, the attitude of practitioners toward the robot indeed
 varies a lot at individual levels. The process that led us to select the schools
-and surgeries for the experiments is interesting in that regard. The very first
+and clinics for the experiments is interesting in that regard. The very first
 school we contacted was a local public school. The head teacher was positive
-about the study, but did ask for official agreement, which never happened. As we
+about the study, but did ask for official approval, which never happened. As we
 learnt several months later, the request was still laying on the desk of a
-regional education officer who happens not to be too keen on novel technologies.
+regional education officer who happens to not be too keen on novel technologies.
 Conversely, the three schools where we did conduct the first studies are all
@@
 Similarly with the occupational therapists: while Thomas' therapist did
-voluntarily contacted us (after hearing about the project at the radio) and
+voluntarily contact us (after hearing about the project on the radio) and
 readily involved herself in the design of the experiment with Thomas, we had
@@
 different means: the number of demonstrations they provided to the robot, the
-amount of qualitative feedback they gave to the robot (in the \emph{Surgery}
-experiment), and to what extend they were focusing on the task (measure of the
+amount of qualitative feedback they gave to the robot (in the \emph{Clinic}
+experiment), and to what extent they were focusing on the task (measure of the
 with-me-ness). Independently of the experimental setting, it appears that the
 children engage easily into the interaction. More interestingly, we show in the
-\emph{Surgery} experiment that they generally take on the role of a teacher
+\emph{Clinic} experiment that they generally take on the role of a teacher
 easily, that they act this role seriously (and not only playfully), and that
@@
 own writing to better help the robot), our system does seem to elicit a Protégé
-effect that not only help the children to remain engaged into their teaching
-over several weeks, but also positively impact their learning process (Vincent
+effect that not only helps the children to remain engaged in their teaching
+over several weeks, but also positively impacts their learning process (Vincent
 strives to write in a consistent manner, as does Thomas). This supports {\it a
@@
 initial material to evidence handwriting improvements (Figure~\ref{fig:stimuli})
-and the study with Thomas provide further data, both quantitative
+and the study with Thomas provides further data, both quantitative
 (Figure~\ref{Thomas_progress}) and qualitative (feedback from the therapist that
@@
 We must however remain cautious here as for the actual role of the system: while
-the children were indeed the one deciding what to teach to the robot and the
+the children were indeed the ones deciding what to teach to the robot and the
 robot was autonomously learning and responding, the role of the adults (the
 experimenters or the occupational therapists) should not be underestimated.
-Beyond the normal explanations on how to operate the tablet and how to interact
-with the robot, the adults played the role of a facilitator in every of the
+Beyond the normal explanations of how to operate the tablet and how to interact
+with the robot, the adults played the role of a facilitator in each of the
 studies by prompting the children to comment on the robot's performance,
@@
 but is also a fundamental part of the learning process itself. In that respect,
-our robot is essentially a tool that creates a favourable learning situation for
-the child, and where the adult (be it a teacher or a therapist) keeps its entire
+our robot is essentially a tool that creates a favorable learning situation for
+the child, and where the adult (be it a teacher or a therapist) keeps their entire
 educative role.
@@
         \emph{learning by teaching} paradigm} from education sciences to
-        robotics, even in a \textbf{complex form}: handwriting is difficult
+        robotics, even in a \textbf{complex form}: handwriting is a difficult
         physical skill, the robot learns and interact autonomously, the child is
@@
 
-    \item this proves essential to \textbf{sustain a long-term interaction}
-        (several hours) although the task would be typically considered as
-        repetitive yet challenging by the children,
-
-    \item besides, the social commitment induces a \textbf{cognitive engagement}
-        of the child with the robot, which is \textbf{a key learning lever} as its
-        elicits \textbf{reflective, meta-cognitive mechanisms on the learning task}.
+    \item this social commitment induces \textbf{cognitive engagement}
+        of the child with the robot, which is \textbf{a key learning lever} as it
+        elicits \textbf{reflective, meta-cognitive mechanisms} on the learning task, and that
+
+    \item we have been able to \textbf{sustain a long-term interaction}
+        (several hours) involving a task that would typically be considered 
+        repetitive and challenging by the children.
 
@@
 
-As we show through extensive experimental validation, these claims are not just
-words: we have effectively deployed robots on the field, with children suffering
+As we have shown through extensive experimental validation, these claims are not just
+words: we have effectively deployed robots in the field, with children suffering
 actual handwriting impairments. Our results are promising in terms of sustaining
-the interest and attention of children on otherwise repetitive writing exercises,
+the interest and attention of children in otherwise repetitive writing exercises,
 and we evidence handwriting improvements.
@@
 
-\paragraph{Possible Ethical Concerns} Two aspects of this research ought to be
+\paragraph{Possible Ethical Concerns} There are two aspects of this research ought to be
 discussed in terms of their possible ethical implications: the perceived role of
@@
 questioned: as we see it, the role of the robot within the classroom (or at the
-therapist's surgery) does not infringe upon the role of the adult (teacher or
+therapist's clinic) does not infringe upon the role of the adult (teacher or
 therapist). The core of the \emph{learning by teaching} paradigm relies on the
@@
 perspective, the robot does not replace the teacher, on the contrary. It plays a
-different role in the classroom, which happen to be novel as well: the robot is
+different role in the classroom, which happens to be novel as well: the robot is
 the \emph{least} performing student, and still a very patient, always eager to
@@
 The importance of the adult is further supported by our experiments: even with
-an autonomous, nominally performing robot, to put the teacher's clothes on is
-not obvious for 5-6 years old children, and during the experiments we conducted,
+an autonomous, nominally performing robot, to play the teacher's role is
+not obvious for 5-6 year old children, and during the experiments we conducted,
 the adult always played a key role at prompting the child to give feedback to
@@
 child is the mentor. As such, relatively little is known on the psychological
-implications for a child to commit in helping a robot, and as advocated by
+implications for a child to commit to helping a robot, and as advocated by
 Belpaeme and Morse~\cite{belpaeme2010time}, we likely need to first gain more
@@
 
-Beyond handwriting, we however do believe that this work provides a novel
+Beyond handwriting, we do however believe that this work provides a novel
 perspective on the role for robots in the field of education. \emph{Learning by
@@
 participation and support. We express our gratitude to Anne Gossin and Fanny
-Lebel for inviting us to their surgery, for their involvement in the experiments
+Lebel for inviting us to their clinic, for their involvement in the experiments
 and for bearing with our disruptive robots. And finally, we want to say a very

### 2016-01-04 rebuttal.tex
minor change to rebuttal

@@
 experiment focused on measuring how \emph{serious} the children were at
-teaching the robot,
-
-Other specific points were raised:
-
-\begin{itemize}
-    \item inclusions criteria should now be clearer,
-    \item
-
-\end{itemize}
+teaching the robot.
 
@@
 turn-taking? If so, how did the system fare with those children?} In the case
-the children did not pro-actively engaged into turn-taking, the facilitator (be
-it one of the experimenter or the therapist) would suggest it.
+the children did not pro-actively engage into turn-taking, the facilitator (be
+it one of the experimenters or the therapist) would suggest it.
 
@@
 objective metric (with-me-ness) that relies on the focus of attention of the
-children to assess to which extend the child is into the task.
+children to assess to which extent the child is into the task.
 

### 2016-01-04 rebuttal.tex
Rephrasing in the rebuttal

@@
 We would like to thank both the reviewers and the editor for going through our
-manuscript. As a consequence of your comments, the manuscript underwent
+manuscript. As a follow-up of your comments, the manuscript underwent
 several significant changes, that we summarize hereafter.
@@
     \item The reporting of the studies is lacking
-    \item Possible ethical concerns are not emphasized
+    \item Possible ethical concerns are not sufficiently discussed
 \end{itemize}
@@
 
-We have added new quantitative data in the reporting of the case study 2, that
-show a quantifiable improvement of handwriting. We want however to remain
+We have added quantitative data in the reporting of the case study 2, that
+explicitly show a quantifiable improvement of handwriting. We want however to remain
 cautious in our claims: handwriting acquisition is a complex cognitive skill.
@@
 
-Finally, we have added the results of one more experiment (conducted during the
+We have also added the results of an additional experiment (conducted during the
 summer 2015), that involved 8 children with handwriting deficits. This
@@
 
+Finally, we have added a whole new section (section 4) that explicitly
+summarizes and discusses our findings.
+
 \section*{Ethical concerns}
 
-An explicit paragraph has been added to the discussion that states the possible
-ethical concerns raised by this work. While the psychological impact of the
+An explicit paragraph has been added to the conclusion that states the possible
+ethical concerns raised by this work. Note that, while the psychological impact of the
 mentor-protégé relationship has been studied in the literature, it has been
-done \emph{from the perspective of the protégé}. In our situation, the protégé
-is the robot, and the child is the mentor: we could not find existing
-literature discussing this situation, but would be nevertheless glad to further
-expand the discussion if necessary.
+discussed from the perspective of the protégé only. In our situation, the protégé
+is the robot, and the child is the mentor: to our best knowledge, the ethical
+implications for the child of such a situation are mostly an open question. We
+underline accordingly this point in the article.
 
@@
 \item \emph{The paper would benefit from further explanation of the adult's
-active role throughout the robot-child interaction} The manuscript hopefully
+active role throughout the robot-child interaction} The new version of the manuscript hopefully
 clarifies the role of the facilitator in the different studies.

### 2016-01-04 main.tex
Rephrasing of the conclusion of the section 3.2

@@
 ``artistic'' writing style preferences that were independent of the actual
-legibility of the robot), and during this specific experiment, we eventually
-found that they were all seriously engaged into teaching.
+legibility of the robot), but none of these children adopted a playful-only
+behavior toward the robot.
 
@@
 number of writing demonstrations they provided and their self-inclination to
-give feedback to the robot), and finally take their teacher's role seriously.
+give feedback to the robot), and finally do usually take their teacher's role seriously.
 

### 2016-01-12 main.tex
[minor] Minor changes suggested by Ferran

@@
 Since our application is about teaching a robot to write, generating (initially
-bad) letters and learning from demonstrations is a core aspect of the project.
+bad) letters and learning from writing demonstrations is a core aspect of the
+project.
 
@@
 reason about the shape of letters in their eigenspace, instead of the natural
-cartesian space. The eigenspace of each letters is spanned by the first $n$
+cartesian space. The eigenspace of each letter is spanned by the first $n$
 eigenvectors (in our experiments, $3 < n < 6$) of the covariance matrix
@@
 (Figure~\ref{fig:schools}). These experiments were meant to technically validate
-the system (is it actually able to autonomously write and learn from
+the system (is it actually able to autonomously write and learn from handwriting
 demonstrations?) and test the interaction (is the apparatus easy to grasp and to

### 2016-01-12 main.tex
List research questions in intro -- Pierre's suggestion

@@
 
-The following sections of the paper go into the details. We first provide a brief
-overview of the robotic system and the interaction it induces. We also present
+\subsection{Research Questions}
+
+These initial considerations can be turned into a set of research questions.
+
+\paragraph{Technical Feasibility} The technical feasibility of our system (and
+its actual implementation) is the first question -- actually a prerequisite --
+that we want to address. The robotic system needs to be able to produce
+handwriting in a believable manner; the legibility of its handwriting (\ie the
+apparent skill of the robot) needs to be easy to parametrise and must evolve
+over time in a controlled manner; the robot needs to acquire, recognize and
+learn from handwriting demonstrations that may be hardly legible (due to the
+impairments of our target children); the design and implementation of the
+robot's behaviors must effectively sustain and support a complex and long-term
+interaction with children; and finally, in contrast to previous work, we aim at
+building a fully autonomous agent that can be deployed and used by non-experts
+(teachers, therapists). Taken together, these technical objectives are
+ambitious and concern a range of different technical subcomponents.
+
+\paragraph{Acceptance} A second research question that we study here relates to
+the acceptance of such a system in educative and medical contexts. Remediation
+to cognitive deficits is a sensitive question for the caretakers, all the more
+so when working with children. Introducing a robot-based system in such
+situations is not self-evident, especially as the mentor-protégé relationship
+further relies on building strong affective and cognitive bonds between the
+child and the robot (\ie the robot is not simply experienced as a passive tool,
+as would be a computer program for instance). By establishing collaborations
+with teachers, parents and therapists, we aim at going to the practitioners'
+field level and explore the conditions for such a robot-based system to be
+accepted.
+
+\paragraph{Engagement} We have already mentioned that handwriting remediation
+requires exercising over long periods of time. In terms of human-robot
+interaction, this translates into a third research question: how to sustain
+several hours of interaction while performing a task (handwriting) which is
+essentially a boring routine? This is usually studied in term of
+\emph{engagement}: how to engage, and thereafter sustain the engagement, of a
+child into an interaction with a robot? As already hinted, we hypothesize that
+the \emph{learning by teaching} paradigm may prove an effective lever to get the
+child to commit to help the robot over an extended period of time. This needs to
+be evidenced, which incidentally calls for actually running interaction
+experiments over long durations (several hours spread over several weeks).
+
+\paragraph{Remediation Efficacy} Finally, the last, key research question is: is
+our approach effective in helping children to recover from handwriting deficits?
+Answering this question in a broad, general way is unlikely as {\it 1)} the
+observed deficits vary a lot between the children (no single metric) and have
+complex, compound causes, and, {\it 2)} ``learning how to write'' is not a
+monolithic, linear skill that could be easily rated from 0 to 1.  However, by
+crossing qualitative assessments made by the caretakers with quantitative
+measures of visual similarity between the letters drawn by the children and some
+reference templates, we aim at assessing the impact of our system on the
+children' handwriting, at least in terms of trend.
+
+\vspace{2em}
+The following sections explore and attempt to answer these
+questions. We first provide an
+overview of the robotic system and the interaction it induces
+in Section~\ref{impl}. We present there
 the machine-learning techniques that allow the robot to learn from the children,
-followed by the implementation on a {\sc nao} robot.
+as well as the actual implementation of the system on a {\sc nao} robot.
 
 We then present and report on the field experiments that we have conducted over
-the last two years, including four studies at schools, one longer experiment with
-eight children in an occupational therapy clinic, and two one-month
-long case studies. While the focus of the school experiments was mostly the
+the last two years, including four studies at schools
+(Section~\ref{studies_school}), one longer experiment with
+eight children in an occupational therapy clinic (Section~\ref{study_normandy}), and two one-month
+long case studies (Sections~\ref{vincent} and~\ref{thomas}). While the focus of the school experiments was mostly the
 technical validation and data acquisition, the three other experiments involved
@@
 \section{Implementation of the Interaction}
+\label{impl}
 
@@
 \subsection{System Validation at Schools}
+\label{studies_school}
 
@@
 case studies that have been conducted so far show that our system can sustain
-children's engagement over durations (5 hours) that are getting closer to what is
+children's engagement over durations (5 hours) that are closer to what is
 expected to have an impact in a real therapeutic context.
@@
 \subsection{Clinic Study: How Children Take on the Role of a Teacher}\label{normandie}
+\label{study_normandy}
 
@@
 \subsection{Case Study 1: Vincent}
+\label{vincent}
 
@@
 \subsection{Case Study 2: Thomas}
+\label{thomas}
 

### 2016-01-13 main.tex
Details how with-me-ness is computed -- Pierre's suggestion

@@
 to human-robot interaction and formalized the exact methodology
-in~\cite{lemaignan2016realtime}\footnote{Note that with-me-ness is a metric that
-allows real-time computation by the robot itself: while we did not yet make
-use of it, it does in principle allow the robot to detect on-line possible
-disengagements, and eventually address them by adapting its physical
-behavior, switching to different activities, etc.}.
+in~\cite{lemaignan2016realtime}. We compute it by first estimating the focus of
+attention of the child over the course of the interaction (using real-time 6D
+head pose estimation); then, by matching this measured focus of attention
+with the \emph{expected} attentional targets. For instance, when the child is supposed
+to write a demonstration, we expect him to look at the tablet if he is actually
+engaged into the task; when the robot is telling a story,
+we expect the engaged child to look at the robot, etc. The resulting values
+(Table~\ref{tab:results-with-me-ness}) represent therefore the percentage of
+interaction where each of the children were actually focusing on the task\footnote{Note that
+with-me-ness is a metric that allows real-time computation by the robot
+itself: while we did not yet make use of it, it does in principle allow the
+robot to detect on-line possible disengagements, and eventually address them
+by adapting its physical behavior, switching to different activities, etc.}.
 

### 2016-01-13 main.tex
Seriously engaging in role -> adopting role -- Pierre's suggestion

@@
 This experiment was conducted to study \item how easily children with actual
-deficits take on the role of a teacher, and \item if they engage
-\emph{seriously} in this role. Measuring actual handwriting improvements was
-not a primary goal of this experiment.
+deficits take on the role of a teacher, and \item if they adopt this role.
+Measuring actual handwriting improvements was not a primary goal of this
+experiment.
 \end{inparaenum}
@@
 feedback with the actual performance of the robot, we can measure to what extent
-the children are seriously assuming their role: if their feedback does correlate
-with the actual performance of the robot, the child is likely engaged into a
-\emph{serious} teaching role-playing.
+the children are adopting their teacher role: if their feedback does correlate
+with the actual performance of the robot, the child has likely successfully
+adopted the teacher role.
 
@@
 impairments do accept well the robot, can commit themselves into long
-interactions, do take on the role of the teacher (as evidenced by both the
+interactions, and do adopt the role of the teacher (as evidenced by both the
 number of writing demonstrations they provided and their self-inclination to
-give feedback to the robot), and finally do usually take their teacher's role seriously.
+give feedback to the robot).
 
@@
 during post-hoc interviews. In that respect, Vincent realized that he was ``good
-enough'' at writing to help someone else. The fact that Vincent seriously
-committed to the role of a teacher is further supported by feedback sent by
+enough'' at writing to help someone else. The fact that Vincent adopted
+the role of a teacher is further supported by feedback sent by
 Vincent's parents a week after the end of the experiment: ``Vincent's

### 2016-01-13 main.tex
Remove mention of public/private schools -- Pierre's suggestion

@@
 varies a lot at individual levels. The process that led us to select the schools
-and clinics for the experiments is interesting in that regard. The very first
-school we contacted was a local public school. The head teacher was positive
-about the study, but did ask for official approval, which never happened. As we
-learnt several months later, the request was still laying on the desk of a
-regional education officer who happens to not be too keen on novel technologies.
-Conversely, the three schools where we did conduct the first studies are all
-private schools, generally open to novel learning methods. The teachers we met
-during these experiments certainly had questions about the rationale and
-end-purpose of the project, but were confident about its educative and social
-value.
+and clinics for the experiments is interesting in that regard.  The head teacher
+of the very first school that we contacted was positive about the study, but he
+could not obtain official approval from his hierarchy due to concerns with
+technologies at school. The three schools where we did conduct the first studies
+did not have such initial reservations. The teachers we met during these
+experiments certainly had questions about the rationale and end-purpose of the
+project, but were confident about its educative and social value.
 

### 2016-01-13 main.tex
Rephrase paragraph on ethics -- Pierre's suggestion

@@
 not otherwise tractable for a teacher with 20 pupils.  This is however unlikely
-to happen soon. In our experience, the teacher keeps such an important role that
-the interaction (and the learning!) would hardly occur if the child is left
-alone (or even semi-alone). We can therefore reassure teachers: robots are not
-going to replace them any time soon, and the initial feedback that we received
-during the focus groups shows that, once explained, our approach make sense and
-is indeed welcomed by the teachers themselves.
+to happen soon. In our experience, the teacher/facilitator keeps an instrumental
+role during the interaction, and the learning would hardly occur if the child is
+left alone (or even semi-alone). The initial feedback that we received from the
+teachers during the focus groups confirms this perception: once explained, our
+approach make sense to them and is indeed welcomed as a relevant pedagogical
+innovation.
 

### 2016-01-13 rebuttal.tex
Minor update to the rebuttal

@@
 summer 2015), that involved 8 children with handwriting deficits. This
-experiment focused on measuring how \emph{serious} the children were at
+experiment focused on assessing (teacher) role adoption by the children when 
 teaching the robot.
 
-Finally, we have added a whole new section (section 4) that explicitly
-summarizes and discusses our findings.
+Finally, we have added new sections (sections 1.3 and 4) that respectively
+explicitly state our research questions, and summarizes and discusses our findings.
 

### 2016-01-13 main.tex
Proof-reading intro

@@
     \includegraphics[width=0.9\linewidth]{henry}
-    \caption{\small Thomas teaching a \sc{Nao} robot how to write numbers, with 
+    \caption{\small Thomas teaching a {\sc Nao} robot how to write numbers, with 
     the help of an occupational therapist.}
@@
 the acceptance of such a system in educative and medical contexts. Remediation
-to cognitive deficits is a sensitive question for the caretakers, all the more
-so when working with children. Introducing a robot-based system in such
-situations is not self-evident, especially as the mentor-protégé relationship
-further relies on building strong affective and cognitive bonds between the
-child and the robot (\ie the robot is not simply experienced as a passive tool,
-as would be a computer program for instance). By establishing collaborations
-with teachers, parents and therapists, we aim at going to the practitioners'
-field level and explore the conditions for such a robot-based system to be
-accepted.
+to cognitive deficits, even mild ones, is a sensitive question for the
+caretakers, all the more so when working with children. Introducing a
+robot-based system in such situations is not self-evident, especially as the
+mentor-protégé relationship further relies on building strong affective and
+cognitive bonds between the child and a machine (here, the robot is not simply
+experienced as a passive tool, as would be an educational computer program for
+instance: it instead plays a potentially disruptive role). By establishing
+collaborations with teachers, parents and therapists, we aim however at going
+down to the practitioners' field level and explore the conditions for such a
+robot-based system to be accepted.
 
@@
 several hours of interaction while performing a task (handwriting) which is
-essentially a boring routine? This is usually studied in term of
-\emph{engagement}: how to engage, and thereafter sustain the engagement, of a
-child into an interaction with a robot? As already hinted, we hypothesize that
-the \emph{learning by teaching} paradigm may prove an effective lever to get the
-child to commit to help the robot over an extended period of time. This needs to
-be evidenced, which incidentally calls for actually running interaction
-experiments over long durations (several hours spread over several weeks).
-
-\paragraph{Remediation Efficacy} Finally, the last, key research question is: is
+essentially perceived as a boring routine? Sustained interaction is usually
+studied in term of \emph{engagement}: how to engage, and thereafter sustain the
+engagement, of a child into an interaction with a robot? We hypothesize that the
+\emph{learning by teaching} paradigm may be an effective lever to get the child
+to commit in helping the robot over an extended period of time. This needs to be
+evidenced -- which incidentally calls for setting up and carrying long-term
+child-robot experiments.
+
+\paragraph{Remediation Efficacy} Finally, the last, key, research question is: is
 our approach effective in helping children to recover from handwriting deficits?
@@
 reference templates, we aim at assessing the impact of our system on the
-children' handwriting, at least in terms of trend.
+children' handwriting, if only in terms of trends.
 
@@
 the last two years, including four studies at schools
-(Section~\ref{studies_school}), one longer experiment with
-eight children in an occupational therapy clinic (Section~\ref{study_normandy}), and two one-month
-long case studies (Sections~\ref{vincent} and~\ref{thomas}). While the focus of the school experiments was mostly the
-technical validation and data acquisition, the three other experiments involved
-children with actual deficits, and gave us initial insights into the relevance and
-effectiveness of our approach.
+(Section~\ref{studies_school}), one longer experiment with eight children in an
+occupational therapy clinic (Section~\ref{study_normandy}), and two one-month
+long case studies (Sections~\ref{vincent} and~\ref{thomas}). While the focus of
+the school experiments was mostly the technical validation of the system and
+data acquisition, the three other experiments involved children with actual
+deficits, and gave us initial insights into the relevance and effectiveness of
+our approach.
 

### 2016-01-18 main.tex
Removed Aude from co-author, at her request

@@
 
-\author{Séverin Lemaignan$^1$, Alexis Jacq$^{1,3}$, Deanna Hood$^1$, Fernando Garcia$^1$, \\
-    Aude Billard$^2$, Ana Paiva$^3$, Pierre Dillenbourg$^1$ \\
+\author{Séverin Lemaignan$^1$, Alexis Jacq$^{1,2}$, Deanna Hood$^1$,\\Fernando
+    Garcia$^1$, Ana Paiva$^2$, Pierre Dillenbourg$^1$ \\
 $^1$CHILI Lab, École Polytechnique Fédérale de Lausanne, Switzerland,\\
-$^2$LASA, École Polytechnique Fédérale de Lausanne, Switzerland,\\
-$^3$Instituto Superior Técnico, University of Lisbon, Portugal}
+$^2$Instituto Superior Técnico, University of Lisbon, Portugal}
 

### 2016-03-16 main.tex
Double column layout
Make the paper shorter

@@
-%\documentclass[twocolumn]{article}
-\documentclass{article}
+\documentclass[twocolumn]{article}
+%\documentclass{article}
 
@@
 
-\subsection{Learning by Teaching Handwriting}
+\subsection*{Learning by Teaching Handwriting}
 
@@
 aspect: by demonstrating handwriting to a robot, we aim at improving the
-\emph{child's} performance. Note that our work must be distinguished from
-``learning from demonstration'' approaches to robots learning physical skills,
-as the agent we present is only simulating poor fine motor skills for interaction
-purposes.
+\emph{child's} performance.
 
@@
 
-\subsection{Agency and Commitment}
+\subsection*{Agency and Commitment}
 
@@
 robot beyond its motivational aspect (playing with an interactive, partially
-autonomous device naturally induces some form of anthropomorphizing, which leads
-to some level of projected agency, and contributes to the overall excitement -- at
-least, on the short-term, before the novelty effect vanishes).  In our case, the
+autonomous device induces excitement -- at
+least, on the short-term).  In our case, the
 role of agency is stronger: it induces meta-cognition (``I am interacting with
@@
 
-Building on these socio-cognitive mechanisms, our intent is therefore to design
-a robotic system that would effectively support handwriting remediation in an
-original way. By getting children to teach a robotic agent how to write, those
-children would both practise without knowing it and recover self-confidence and
-self-esteem by supporting a worse-than-themselves robotic ``student''.
-
-\subsection{Research Questions}
-
-These initial considerations can be turned into a set of research questions.
-
-\paragraph{Technical Feasibility} The technical feasibility of our system (and
-its actual implementation) is the first question -- actually a prerequisite --
-that we want to address. The robotic system needs to be able to produce
-handwriting in a believable manner; the legibility of its handwriting (\ie the
-apparent skill of the robot) needs to be easy to parametrise and must evolve
-over time in a controlled manner; the robot needs to acquire, recognize and
-learn from handwriting demonstrations that may be hardly legible (due to the
-impairments of our target children); the design and implementation of the
-robot's behaviors must effectively sustain and support a complex and long-term
-interaction with children; and finally, in contrast to previous work, we aim at
-building a fully autonomous agent that can be deployed and used by non-experts
-(teachers, therapists). Taken together, these technical objectives are
-ambitious and concern a range of different technical subcomponents.
-
-\paragraph{Acceptance} A second research question that we study here relates to
-the acceptance of such a system in educative and medical contexts. Remediation
-to cognitive deficits, even mild ones, is a sensitive question for the
-caretakers, all the more so when working with children. Introducing a
-robot-based system in such situations is not self-evident, especially as the
-mentor-protégé relationship further relies on building strong affective and
-cognitive bonds between the child and a machine (here, the robot is not simply
-experienced as a passive tool, as would be an educational computer program for
-instance: it instead plays a potentially disruptive role). By establishing
-collaborations with teachers, parents and therapists, we aim however at going
-down to the practitioners' field level and explore the conditions for such a
-robot-based system to be accepted.
-
-\paragraph{Engagement} We have already mentioned that handwriting remediation
-requires exercising over long periods of time. In terms of human-robot
-interaction, this translates into a third research question: how to sustain
-several hours of interaction while performing a task (handwriting) which is
-essentially perceived as a boring routine? Sustained interaction is usually
-studied in term of \emph{engagement}: how to engage, and thereafter sustain the
-engagement, of a child into an interaction with a robot? We hypothesize that the
-\emph{learning by teaching} paradigm may be an effective lever to get the child
-to commit in helping the robot over an extended period of time. This needs to be
-evidenced -- which incidentally calls for setting up and carrying long-term
-child-robot experiments.
-
-\paragraph{Remediation Efficacy} Finally, the last, key, research question is: is
-our approach effective in helping children to recover from handwriting deficits?
-Answering this question in a broad, general way is unlikely as {\it 1)} the
-observed deficits vary a lot between the children (no single metric) and have
-complex, compound causes, and, {\it 2)} ``learning how to write'' is not a
-monolithic, linear skill that could be easily rated from 0 to 1.  However, by
-crossing qualitative assessments made by the caretakers with quantitative
-measures of visual similarity between the letters drawn by the children and some
-reference templates, we aim at assessing the impact of our system on the
-children' handwriting, if only in terms of trends.
-
 \vspace{2em}
-The following sections explore and attempt to answer these
-questions. We first provide an
-overview of the robotic system and the interaction it induces
-in Section~\ref{impl}. We present there
-the machine-learning techniques that allow the robot to learn from the children,
-as well as the actual implementation of the system on a {\sc nao} robot.
+These initial considerations can be turned into a set of research questions: how
+to develop a fully autonomous robotic system that produces believable
+handwriting, foster interaction and adequately learn from children's demonstration?
+Can such a system be accepted by children and practitioners alike, and
+eventually be integrated in existing educative and therapeutic practises?
+Can we build and evidence interactions that sustains children's engagement over extended
+periods of time? And finally, is this novel approach effective? Do we evidence
+actual handwriting improvements by children facing difficulties?
+
+We attempt to provide initial insights on these questions in the remaining
+sections of the article. We first provide an overview of the robotic system and
+the interaction it induces in Section~\ref{impl}. We present there the
+machine-learning techniques that allow the robot to learn from the children, as
+well as the actual implementation of the system on a {\sc nao} robot.
 
@@
 face-to-face child-robot interaction with an autonomous Aldebran's {\sc nao}
-robot.
-
-A tactile tablet (with a custom application) is used for both the robot and the
+robot. A tactile tablet is used for both the robot and the
 child to write. During a typical round, the child requests the robot to write
@@
 child then pulls back the tablet, corrects the robot's attempt by writing
-him/herself on top of or next to the robot's writing (see Figure~\ref{fig:diego}),
-and ``sends'' his/her demonstration to the robot by pressing a small button on
-the tablet. The robot ``learns'' from this demonstration and tries again.
+herself on top of or next to the robot's writing (see Figure~\ref{fig:diego}),
+and sends her demonstration to the robot by pressing a small button on
+the tablet. The robot learns from this demonstration and tries again.
 
@@
     \centering
-    \includegraphics[width=0.6\columnwidth]{experimental_setup}
-    \caption{\small Our experimental setup: face-to-face interaction with a {\sc
-        nao} robot.  The robot writes on the tactile tablet, the child then
-        corrects the robot by directly overwriting its letters on the tablet
-        with a stylus. An adult (either a therapist or an experimenter,
-        depending on the studies), remains next to the child to guide the work
+    \includegraphics[width=0.9\columnwidth]{experimental_setup}
+    \caption{\small Our experimental setup: the robot writes on a tactile tablet, the child then
+        corrects the robot by overwriting its letters with a stylus. An adult
+        (either a therapist or an experimenter) guides the work
         (prompting, turn taking, etc.). For some studies, a second tablet and an
@@
 
-We detail each of these in the following sections.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 generated from a standard dataset of adult letters (the UJI Pen Characters 2
-dataset~\cite{Llorens2008}).  This procedure, based on a Principle Component
-Analysis (PCA), is explained in details in a previously published
-article~\cite{hood2015when}.
-
-New letters can be generated by varying the weightings in linear combinations of
+dataset~\cite{Llorens2008}).  Our method, based on a Principle Component
+Analysis (PCA), is presented in~\cite{hood2015when}.
+
+New letters are generated by varying the weightings in linear combinations of
 these eigenvectors, with distortions that are actually plausible handwriting errors: 
@@
         PCA-based approach allows for the \emph{automatic} generation of letters whose
-        errors can be \emph{semantically} interpreted, \eg A has too large a bottom
+        errors are typical of children handwriting: A has too large a bottom
         loop, B has a wide top loop, the bottom loop of C is not correctly
@@
 for example nine allographs of ``h'' written by a 6 years old child, along with
-a reference letter. By projecting each of the
+the reference shape. By projecting each of the
 demonstrations onto the eigenspace of ``h'' (Figure~\ref{fig:h}b), we observe
@@
 
-\begin{figure}[ht!]
+\begin{figure*}[ht!]
     \centering
     \subfigure[Nine allographs of the cursive ``h'', next to the reference]{
-        \includegraphics[width=0.45\columnwidth]{h}
+        \includegraphics[width=0.45\linewidth]{h}
     }
@@
         first 3 eigenvectors: clusters arise, that actually match writing styles.]{
-        \includegraphics[width=0.45\columnwidth]{eigenspace-uniformization}
+        \includegraphics[width=0.45\linewidth]{eigenspace-uniformization}
     }
@@
     \label{fig:h}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[ht!]
-\centering
-
-\resizebox{0.7\linewidth}{!}{%
-
-\begin{tikzpicture}[
-    >=latex,
-    node distance=2cm,
-    every edge/.style={draw, very thick},
-    redarrow/.style={draw,red, text=black},
-    greenarrow/.style={draw,GreenYellow,text=black},
-    yellowarrow/.style={draw,BurntOrange,text=black},
-    cmpt/.style={draw, align=center, rounded corners, inner sep=5pt, font=\sf, fill=black!20},
-    label/.style={midway, align=left, font=\scriptsize\sf, fill=white, above,opacity=0,text opacity=1}]
-
-    \node at (0,0) (laptop) {\includegraphics[width=2cm]{laptop}};
-    \node[below right=2 of laptop] (nao) {\includegraphics[width=2cm]{nao}};
-    \node[below left=2 of laptop] (tablet) {\includegraphics[width=2cm]{tablet+stylus}};
-    \node[above=2 of laptop] (selection) {\includegraphics[width=2cm]{selection_tablet}};
-
-    \node[draw,above right=2 of laptop,anchor=north west,text width=4cm] (processes)
-    {\sf\scriptsize machine-learning, \\letters/gestures
-    generation, \\interaction supervision};
-    \path (laptop) edge[dashed] (processes);
-
-    \path (nao) edge [->,redarrow, bend left] node[label, auto] {robot state} (laptop);
-    \path (laptop) edge [->,greenarrow, bend left] node[label, auto] {writing gestures} (nao);
-
-    \path (tablet) edge [->,redarrow, bend left] node[label, auto] {demonstrations,\\turn taking} (laptop);
-    \path (laptop) edge [->,redarrow, bend left] node[label,
-    auto=right,align=right]
-    {path of\\ letters to display} (tablet);
-
-    \path (selection) edge [->,redarrow] node[label, auto=right] {letter/word to write} (laptop);
-
-    \path (-5, 2) edge [->, redarrow] node[label] {ROS} ++(1, 0);
-    \path (-5, 2.6) edge [->, greenarrow] node[label] {NaoQI} ++(1, 0);
-
-\end{tikzpicture}
-}
-
-\caption{\small \textbf{Overview of the system}. In total, the system runs about 10 ROS nodes,
-    distributed over the robot itself, a central laptop and Android tablets.}
-
-    \label{fig:archi}
-\end{figure}
-
-The overall architecture of the system (Figure~\ref{fig:archi}) is therefore
+The overall architecture of the system is therefore
 spread over several devices: the {\sc nao} robot itself, that we address via
-both a ROS API\footnote{The ROS documentation for {\sc nao} is available at
-\url{http://wiki.ros.org/nao}.} and the Aldebaran-provided NaoQI API, one to
+both a ROS API and the Aldebaran-provided NaoQI API, one to
 four Android tablets (the main tablet is used to draw the robot's letter and to
@@
 different schools (more than 70 children aged 5 to 8; relatively short duration
-interactions), in an occupational therapy clinic (8 children; each
+interactions), in an occupational therapy clinic (8 children, each
 interacting several hours with the system), and during two case studies that
-each lasted several weeks. Table~\ref{studies} gives an overview of these
-studies.
-
-\begin{table}[ht!]
-\centering
-\caption{\small Field studies conducted within the project}
-\label{studies}
-\footnotesize
-\begin{tabular}{@{}lp{4cm}p{2.2cm}p{2cm}l@{}}
-\toprule
-{\bf Study}        & {\bf Type}                               & {\bf Avg. duration}    & {\bf \# Children}                     & {\bf Ages} \\ \midrule
-{\it School A (1)} & Unstructured group interaction at school & 16 min/group           & 4 $\times$ 8 children                 & 6-7        \\
-{\it School B}     & Individual/Pair interaction at school    & 11 min/group           & 7 (individual) + 7 $\times$ 2 (pairs) & 7-8        \\
-{\it School C}     & Pair interaction at school               & 26 min/group           & 7 $\times$ 2                          & 5-6        \\
-{\it School A (2)} & Individual interaction at school         & 20 min                 & 6                                     & 5-6        \\ \midrule
-{\it Clinic}       & Individual interaction at clinic         & 3 sessions $\times$ 1h & 8                                     & 6-8        \\ \midrule 
-{\it Vincent}      & Case-study                               & 4 weeks $\times$ 1.5h  & 1                                     & 6          \\
-{\it Thomas}       & Case-study                               & 4 weeks $\times$ 1h    & 1                                     & 5          \\ \bottomrule
-\end{tabular}
-\end{table}
-
+each lasted several weeks.
 We report hereafter the main design choices and results for each of these
@@
     \centering
-    \includegraphics[height=3.9cm]{schools}
-    \includegraphics[height=3.9cm]{schools2}
+    \includegraphics[height=5cm]{schools2}
     \caption{\small The first field studies were focused on technical validation, with
@@
 
-\paragraph{System Validation}
+\paragraph{System validation}
 
@@
 approach to be effective. When asked, no child indicated that they did not
-believe that the robot was writing by itself: that it was simply miming the 
-actions in the air as the tablet displayed the letters. There were, at times, questions
+believe that the robot was writing by itself. There were, at times, questions
 about the robot's writing without a pen at the beginning of the interaction, but when
 advised that the robot ``tells the tablet what it wants to write,'' this was
-accepted by the children.  In addition, teachers interviewed for their feedback on
-the system advised that children are asked to draw letters in the air in a
-similar manner as part of their handwriting education. The behavior is hence
-not unfamiliar to children.
+accepted by the children.
 
@@
 
-The literature suggests that 20 handwriting practice sessions is found to be the
-minimum to demonstrate effective results in handwriting
-remediation~\cite{Hoy2011}. This highlights the necessity to sustain a
-child-robot engagement over the long-term if we want to achieve measurable
-learning gains.
-
 Factually, the children engaged into the teaching activity: in the
-\textit{School B} study, for instance, they demonstrated an average of 10.9
+study at school B, for instance, they demonstrated an average of 10.9
 demonstration letters (SD = 4.4) for an average session duration of 11 minutes.
@@
 \emph{With-me-ness} is a quantifiable precursor of engagement: it measures the
-percentage of time spent by the child focusing on the task at hand. This metric
-was first devised in the context of computer-supported
-learning~\cite{sharma2014me}, and we have previously studied its applicability
-to human-robot interaction and formalized the exact methodology
-in~\cite{lemaignan2016realtime}. We compute it by first estimating the focus of
+percentage of time spent by the child focusing on the task at hand~\cite{lemaignan2016realtime}. We compute it by first estimating the focus of
 attention of the child over the course of the interaction (using real-time 6D
@@
 with the \emph{expected} attentional targets. For instance, when the child is supposed
-to write a demonstration, we expect him to look at the tablet if he is actually
+to write a demonstration, we expect her to look at the tablet if she is actually
 engaged into the task; when the robot is telling a story,
-we expect the engaged child to look at the robot, etc. The resulting values
-(Table~\ref{tab:results-with-me-ness}) represent therefore the percentage of
-interaction where each of the children were actually focusing on the task\footnote{Note that
-with-me-ness is a metric that allows real-time computation by the robot
-itself: while we did not yet make use of it, it does in principle allow the
-robot to detect on-line possible disengagements, and eventually address them
-by adapting its physical behavior, switching to different activities, etc.}.
-
-\begin{table}[h!]
+we expect the engaged child to look at the robot, etc.
+
+\begin{table*}[h!]
     \centering
     \caption{\textbf{Levels of with-me-ness}. Percentage of interaction time
-        during which the child was effectively focusing his/her attention on the
+        during which the child was effectively focusing her attention on the
         task. The six children are those from the second study at School A.
@@
     \label{tab:results-with-me-ness}
-\end{table}
+\end{table*}
 
@@
 
-The other three experiments (the group study at the clinic and the two
-case-studies) provide further qualitative evidence of engagement over longer
-interaction periods. In particular, as reported hereafter, the two 4-week long
-case studies that have been conducted so far show that our system can sustain
-children's engagement over durations (5 hours) that are closer to what is
-expected to have an impact in a real therapeutic context.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{inparaenum}[\itshape 1\upshape)]
-This experiment was conducted to study \item how easily children with actual
-deficits take on the role of a teacher, and \item if they adopt this role.
-Measuring actual handwriting improvements was not a primary goal of this
-experiment.
-\end{inparaenum}
-
-The experiment took place at an occupational therapy clinic in
-Normandy, France. Eight children participated, selected by the occupational
-therapist based on their age and type of deficit (all related to handwriting).
-Valérie (female, 7 years old), Antoine (male, 6.5) and Johan (male, 7) are under
-the direct care of the occupational therapist. Émilien (male, 8) and Mathieu
-(male, 7) are repeating their school year because of writing difficulties. Marie
-(female, 6) and Adèle (female, 8) are both ranked at the bottom of their
-respective classes in writing activities. Nicolas (male, 7) is under the care of
-a neurologist, and has been diagnosed with specific language impairment. Given
-their age and school year, all of these children would be expected to know how to
-correctly shape cursive letters.
-
-Over a period of two weeks, each child attended three times a one hour long
-session (except for Marie and Adèle who only attended one session). The
-experimenter's role was limited to the explanation of the task and the basic
-tablet usage. For this experiment, the children were provided with two tablets:
-one to choose the words (or letters) to teach to the robot, the other one to
-write, as in the validation studies. We also provided paper-based templates of
-the cursive letters, were the children to ask for them.
+This experiment was conducted to study how easily children with actual
+deficits take on the role of a teacher, and to what extent they adopt this role
+(measuring actual handwriting improvements was not a primary goal of this
+experiment).
+
+The experiment took place at an occupational therapy clinic in Normandy, France.
+Eight children participated, selected by the occupational therapist based on
+their age and type of deficit (all related to handwriting).  Given their age (6
+to 8 years old) and school year, all of these children would be expected to know
+how to correctly shape cursive letters. Each child attended three times a one
+hour long session spread over two weeks (except for two of them who only
+attended one session). The experimenter's role was limited to the explanation of
+the task and the basic tablet usage. The children were provided with two
+tablets: one to choose the words (or letters) to teach to the robot, one to
+write.
 
@@
 have to help the robot to improve its writing style), so as to assess if and how
-children would naturally take on the role of the teacher. We additionally assessed
-how seriously they engage into helping the robot through two additional buttons
-on the tablet: a green ``thumbs up'' and a red ``thumbs down''. The children
-were told to freely use them to evaluate the robot's improvements (``thumb up''
-to give positive feedback, ``thumb down'' to convey negative feedback). Our
-assessment builds on the hypothesis that the more the child provides feedback to
-the robot, the more they assume the role of a teacher. Then, by correlating the
-feedback with the actual performance of the robot, we can measure to what extent
-the children are adopting their teacher role: if their feedback does correlate
-with the actual performance of the robot, the child has likely successfully
-adopted the teacher role.
+children would naturally take on the role of the teacher. We additionally
+assessed how seriously they engage into helping the robot through two additional
+buttons on the tablet: a green ``thumbs up'' and a red ``thumbs down''. The
+children were told to freely use them to evaluate the robot's improvements
+(``thumb up'' to give positive feedback, ``thumb down'' to give negative
+feedback). Our assessment builds on the hypothesis that the more the child
+provides feedback to the robot, the more they assume the role of a teacher.
+Then, by correlating the feedback with the actual performance of the robot, we
+can measure to what extent the children are adopting their teacher role: if
+their feedback does correlate with the actual performance of the robot, the
+child has likely successfully adopted the teacher role.
 
@@
 
-\begin{table}
-    \centering
-    \begin{tabular}{ccccl}
-        \toprule
-        \bf Child      & \bf \# Demos/hour & \bf \# Positive & \bf \# Negative & $r$ \\ \midrule
-        \emph{Valérie} & 42           & 24              & 6               & 0.25 \small\tt **  \\ 
-        \emph{Émilien} & 74           & 20              & 9               & 0.06 \small\it ns  \\
-        \emph{Mathieu} & 43           & 10              & 3               & 0.23 \small\tt **  \\
-        \emph{Nicolas} & 38           & 16              & 4               & 0.31 \small\tt *** \\
-        \emph{Johan}   & 32           & 10              & 5               & 0.10 \small\it ns  \\
-        \emph{Antoine} & 27           & 10              & 3               & 0.20 \small\tt *   \\
-        \emph{Adèle}   & 35           & 4               & 2               & 0.28 \small\tt *   \\
-        \emph{Marie}   & 40           & 5               & 1               & -0.02 \small\it ns \\ \bottomrule
-    \end{tabular}
-    \caption{\small Feedback from the children to the robot. \emph{\#Demos}
-        denotes the average number of demonstrations per hour provided by the children;
-        \emph{\#Positive} and \emph{\#Negative} the total number of positive \resp
-        negative feedbacks they provided. $r$ is the correlation coefficient
-        between the feedback provided by the children and the performance of the
-        robot. Results taken from~\cite{jacq2016building}.}
-
-    \label{table:scores}
-\end{table}
-
-All children maintained their engagement during the entirety of the sessions. They
+All children maintained their engagement during the whole sessions. They
 provided on average 42 demonstrations per session. All children used the
-feedback buttons (in total, 99 ``thumbs up'' and 33 ``thumbs down'' were recorded,
-see Table~\ref{table:scores}). This indicates that they are all able to engage
-into playing the role of a teacher.
+feedback buttons (in total, 99 ``thumbs up'' and 33 ``thumbs down'' were
+recorded). This indicates that they are all able to engage into playing the role
+of a teacher.
 
@@
 improvements of the robot, we estimate the robot's progress as the difference
-between an initial score (euclidian distance between the shape drawn by
-the robot at its first attempt and the shape of the reference letter) and the
+between an initial score (euclidian distance between the shape drawn by the
+robot at its first attempt and the shape of the reference letter) and the
 robot's score at the current round of demonstrations. We then correlate this
 progress to the positive or negative feedback provided by the children (details
-of the method are presented in~\cite{jacq2016building}).
+of the method, as well as the complete results, are presented
+in~\cite{jacq2016building}).
 
 We find that the feedback of five out of the eight children does significantly
-correlate with the actual performance of the robot (Table~\ref{table:scores}).
-This indicates that these children are effectively taking on the teacher's
-role and are seriously providing feedback to the robot. The observation of
-the three remaining children reveal a variety of behaviors (for instance, one
-was actually rating how ``nice'' the robot is, and another one had rather
-``artistic'' writing style preferences that were independent of the actual
-legibility of the robot), but none of these children adopted a playful-only
-behavior toward the robot.
+correlate with the actual performance of the robot.  This indicates that these
+children are effectively taking on the teacher's role and are seriously
+providing feedback to the robot. The observation of the three remaining children
+reveal a variety of behaviors (for instance, one was actually rating how
+``nice'' the robot is, and another one had rather ``artistic'' writing style
+preferences that were independent of the actual legibility of the robot), but
+none of these children adopted a playful-only behavior toward the robot.
 
@@
 number of writing demonstrations they provided and their self-inclination to
-give feedback to the robot).
+give adequate feedback to the robot).
 
@@
 \label{vincent}
-
-\paragraph{Context, Study Design}
 
@@
     \centering
-    \includegraphics[width=0.5\linewidth]{diego}
+    \includegraphics[width=0.8\linewidth]{diego}
     \caption{\small Vincent correcting {\sc nao}'s attempt by rewriting the
@@
 
-\paragraph{Context, Study Design}
-
 The second long-term study was designed in collaboration with an occupational
@@
 Despite his attention deficit, Thomas was able to remain engaged in the activity
-during more than forty minutes in each session (a long time for a five year
+during more than forty minutes in each session (a long time for a five years
 old). In total, 55 allographs out of 82 demonstrated by the child were
@@
 duration of more than 38 hours, in multiple experimental configurations
-(individual interaction \vs pairs \vs groups; at school, in the lab, at a
-occupational therapy clinic; short interactions \vs long, repeated interactions; 5 to 8 year old
+(individual interaction \vs pairs \vs groups; at school, in the lab, at an
+occupational therapy clinic; short interactions \vs long, repeated interactions;
+5 to 8 years old
 children). We summarize hereafter the main findings from these experiments and
@@
 \paragraph{Technical Assessment} We set ourselves the challenge of developing an
-autonomous robotic system able to perform handwriting tasks with children.
-Over the course of the different studies and experiments, thousands of
-letters have been indeed generated and drawn on the tactile tablet by the robot, in
-response to thousands of demonstrations from children. Most of the experiments have been
-carried out in the wild, at six different locations in total. As reported in the
-technical studies section, the system performed generally well: the children had no
+autonomous robotic system able to perform handwriting tasks with children in
+real world environments. The system performed generally well: the children had no
 issues interacting with the robot, the robot experienced relatively few crashes,
@@
 
-\paragraph{Acceptance} When we started the project, the mere acceptance of the
-robot in the educative and/or medical context of handwriting remediation was not
-obvious. As expected, the attitude of practitioners toward the robot indeed
-varies a lot at individual levels. The process that led us to select the schools
-and clinics for the experiments is interesting in that regard.  The head teacher
-of the very first school that we contacted was positive about the study, but he
-could not obtain official approval from his hierarchy due to concerns with
-technologies at school. The three schools where we did conduct the first studies
-did not have such initial reservations. The teachers we met during these
-experiments certainly had questions about the rationale and end-purpose of the
-project, but were confident about its educative and social value.
-
-Similarly with the occupational therapists: while Thomas' therapist did
+\paragraph{Acceptance} The acceptance of the
+robot in a educative and/or medical context was not initially
+obvious, and indeed, the attitude of practitioners toward the robot varies.
+For instance, while Thomas' therapist did
 voluntarily contact us (after hearing about the project on the radio) and
-readily involved herself in the design of the experiment with Thomas, we had
+readily involved herself in the design of the experiment, we had
 less positive feedback from other practitioners: in one instance, we contacted
-ourselves a local group of school psychologists and occupational therapists to
+ourselves a local group of school psychologists to
 present the project: after the meeting, the opinions were definitively mixed,
@@
 
-Lastly (and as expected), the reaction of the children to the robot was good:
-they enjoyed interacting with the system, and, as we have shown, they actually
-committed to their teacher role. The teaching situation was well accepted: while
-we had initial doubts about how believable a learner the robot would be, the
-children did not appear to show any specific issues with it. In particular,
-technical choices like having the robot to only gesture writing on the tablet
-instead of actually \emph{physically} writing with a pen, were not raised as
-issues.
+Expectedly the reaction of the children to the robot was good: they enjoyed
+interacting with the system, and, as we have shown, they actually committed to
+their teacher role over extended periods of time. The teaching situation was
+well accepted and technical constraints like having the robot to only gesture
+writing on the tablet instead of actually \emph{physically} writing with a pen,
+were not actually raised as issues.
 
@@
 
-Another finding relates to the Protégé effect: as seen with Vincent's case-study
-(with the parents emphasizing how Vincent was aware he had to be consistent in
-his writing to help the robot, or with him sending us a mail several months
-later to know how the robot was doing with its writing) or with Thomas's
-case-study (when Thomas realizes that the robot ``does not understand'' when his
-demonstrations are not legible enough, and consequently quickly improves his
-own writing to better help the robot), our system does seem to elicit a Protégé
-effect that not only helps the children to remain engaged in their teaching
-over several weeks, but also positively impacts their learning process (Vincent
-strives to write in a consistent manner, as does Thomas). This supports {\it a
-posteriori} our choice to build an interaction situation based on the
-\emph{learning by teaching} paradigm.
+Another finding relates to the Protégé effect: as seen with Vincent (when he
+sends us a mail several months later to know how the robot is doing with its
+writing) or with Thomas (when he realizes that the robot ``does not
+understand'', and consequently quickly improves his own writing to better help
+the robot), our system does elicit a Protégé effect that not only helps the
+children to remain engaged in their teaching over several weeks, but also
+positively impacts their learning process (Vincent strives to write in a
+consistent manner, as does Thomas). This supports {\it a posteriori} our choice
+to build an interaction situation based on the \emph{learning by teaching}
+paradigm.
 
@@
 We believe that this research provides a novel perspective on educative robots
-at several levels. We have shown that:
+at several levels. Namely, we show that:
 
@@
 
-As we have shown through extensive experimental validation, these claims are not just
-words: we have effectively deployed robots in the field, with children suffering
-actual handwriting impairments. Our results are promising in terms of sustaining
-the interest and attention of children in otherwise repetitive writing exercises,
-and we evidence handwriting improvements.
-
-It is however still early to quantify the lasting effects of this remediation:
-handwriting is a complex cognitive skill, that builds on many individual and
-social factors. Self-confidence is one of them. Our approach endows the child
-with the role of a teacher who can help a robot: we expect it may as well help
-some children to recover self-esteem and self-confidence by putting them in a
-positive, gratifying role. The experiments that we have conducted so far do not
-allow us to confirm this hypothesis yet, and more research will have to be
-conducted in this direction.
+Supported by several field experiments, our initial results are promising as we
+ultimately evidence handwriting improvements. It is however still early to quantify the
+lasting effects of this remediation: handwriting is a complex cognitive skill,
+that builds on many individual and social factors. Self-confidence is one of
+them. Our approach endows the child with the role of a teacher who can help a
+robot: we expect it may as well help some children to recover self-esteem and
+self-confidence by putting them in a positive, gratifying role. The experiments
+that we have conducted so far do not allow us to confirm this hypothesis yet,
+and more research will have to be conducted in this direction.
 
@@
 the \emph{least} performing student, and still a very patient, always eager to
-improve, one.
-
-The importance of the adult is further supported by our experiments: even with
-an autonomous, nominally performing robot, to play the teacher's role is
-not obvious for 5-6 year old children, and during the experiments we conducted,
-the adult always played a key role at prompting the child to give feedback to
-the robot or to move to the next letter or word.  At a higher orchestration
-level (and as reported in the two case studies with Vincent and Thomas), the
-educational scenarios were also always designed and monitored by the adults.
-
-We initially envisioned our system to be run in the back of a classroom with one
-child: this would have allowed an individual, face-to-face remediation approach,
-not otherwise tractable for a teacher with 20 pupils.  This is however unlikely
-to happen soon. In our experience, the teacher/facilitator keeps an instrumental
+improve, one. In our experience, the teacher/facilitator keeps an instrumental
 role during the interaction, and the learning would hardly occur if the child is
@@
 
-We warmly thank all the children, teachers and staff of the International School
-of Geneva, the Institut Florimont and the Haut-Lac school for their
-participation and support. We express our gratitude to Anne Gossin and Fanny
-Lebel for inviting us to their clinic, for their involvement in the experiments
-and for bearing with our disruptive robots. And finally, we want to say a very
-special \emph{Thank you!} to Vincent, Thomas, Valérie, Émilien, Mathieu, Johan,
-Nicolas, Antoine, Adèle and Marie: by spending several hours being patient
-teachers for our robots, you have paved the way for what may become tomorrow a
-new way of learning how to deal with all these letters and words!
+We warmly thank all the children who participated to the studies, and the
+teachers and staff of the International School of Geneva, the Institut Florimont
+and the Haut-Lac school for their support. We express our particular gratitude
+to Anne Gossin and Fanny Lebel for inviting us to their clinics.
 
